---
title: "analysis_pipeline"
author: "Zilong Wang"
date: "2026-03-16"
output:
  pdf_document: default
  html_document: default
---
# Download and install all the required R packages
```{r}
# Install CRAN packages
install.packages(c(
  "ggplot2", "patchwork", "ggpubr", "dplyr", "scales",
  "stringr", "ape", "data.table", "flextable", "vegan",
  "parallelDist", "reshape2", "forcats"
))

# Install Bioconductor packages
if (!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}
BiocManager::install(c("Biostrings", "ComplexHeatmap"))

# itol.toolkit (GitHub)
install.packages("devtools")
devtools::install_github("TongZhou2017/itol.toolkit")
```



#The proportion of annotable reads in the visualized sequence to the total number of reads
```{r,fig.height = 5,fig.with = 12}
rm(list = ls())
library(ggplot2)
library(patchwork)
sample_taxid_abundance_16s = read.delim("raw/sample_taxid_abundance_16s.txt", row.names = 1)
sample_taxid_abundance_its = read.delim("raw/sample_taxid_abundance_its.txt", row.names = 1)
# Calculate the annotation ratio ((1 - Unassigned / Total sum of each column))
unassigned_values <- as.numeric(sample_taxid_abundance_16s[rownames(sample_taxid_abundance_16s) == "Unassigned", ])
col_totals <- colSums(sample_taxid_abundance_16s)
annotated_prop <- 1 - (unassigned_values / col_totals)
plot_data <- data.frame(annotated_proportion = annotated_prop)

# Calculate the basic statistics (mean and standard deviation)
mean_val <- mean(plot_data$annotated_proportion, na.rm = TRUE)
sd_val <- sd(plot_data$annotated_proportion, na.rm = TRUE)


hist_plot1 <- ggplot(plot_data, aes(x = annotated_proportion)) +
  geom_histogram(fill = "#56B394", color = "black", bins = 30, alpha = 0.8) + 
  xlab("The annotated proportion") +
  ylab("Counts")+
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.1,
           label = paste0("Mean: ", round(mean_val, 3), "\nSD: ", round(sd_val, 3)),
           size = 5) +
  theme_classic() + 
  theme(
    legend.position = "top" ,
    panel.background = element_blank(), 
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2), 
    plot.title = element_text(
      size = 20,
      face = "bold",
      color = "black",
      hjust = 0.5,  
      margin = margin(b = 10)  
    ),
    plot.subtitle = element_text(
      size = 12,
      color = "black",
      hjust = 0.5,  
      lineheight = 1.2,  
      margin = margin(b = 15) 
    ),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y = element_text(size = 14, color = "black", angle = 0), 
    axis.title.y = element_text(size = 14, color = "black", face = "bold"), 
    legend.title = element_text(size = 14, color = "black"), 
    legend.text = element_text(size = 14, color = "black"), 
    plot.margin = margin(t = 10,  
                         r = 30,  
                         b = 10,  
                         l = 10,  
                         unit = "pt") 
  )

#its
unassigned_values <- as.numeric(sample_taxid_abundance_its[rownames(sample_taxid_abundance_its) == "Unassigned", ])
col_totals <- colSums(sample_taxid_abundance_its)
annotated_prop <- 1 - (unassigned_values / col_totals)
plot_data <- data.frame(annotated_proportion = annotated_prop)
mean_val <- mean(plot_data$annotated_proportion, na.rm = TRUE)
sd_val <- sd(plot_data$annotated_proportion, na.rm = TRUE)
hist_plot2 <- ggplot(plot_data, aes(x = annotated_proportion)) +
  geom_histogram(fill = "#56B394", color = "black", bins = 30, alpha = 0.8) +
  xlab("The annotated proportion") +
  ylab("Counts")+
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.1,
           label = paste0("Mean: ", round(mean_val, 3), "\nSD: ", round(sd_val, 3)),
           size = 5) +
  theme_classic() + 
  theme(
    legend.position = "top" ,
    panel.background = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2), 
    plot.title = element_text(
      size = 20,
      face = "bold",
      color = "black",
      hjust = 0.5,
      margin = margin(b = 10) 
    ),
    plot.subtitle = element_text(
      size = 12,
      color = "black",
      hjust = 0.5,  
      lineheight = 1.2, 
      margin = margin(b = 15) 
    ),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y = element_text(size = 14, color = "black", angle = 0),
    axis.title.y = element_text(size = 14, color = "black", face = "bold"), 
    legend.title = element_text(size = 14, color = "black"),
    legend.text = element_text(size = 14, color = "black"), 
    plot.margin = margin(t = 10,  
                         r = 30,  
                         b = 10,  
                         l = 10, 
                         unit = "pt")
  )
library(ggpubr)
combined_plot <- ggarrange(hist_plot1, hist_plot2,  
                           ncol = 2 
)
combined_plot
dir.create("result/Sample_Overview")
ggsave("result/Sample_Overview/Annotated_Proportions.pdf",combined_plot,width = 11,height = 3.2)
```

#Visualization extraction methods and the proportion of sequencing regions
```{r,fig.show='animate'}
rm(list = ls())
source("Tools.R")
library(ggplot2)
library(dplyr)
library(scales)
metadata_16s <- read.delim("raw/metadata_16s.txt")
metadata_its <- read.delim("raw/metadata_its.txt")

# 16s extract_way
p1 = draw_pie_chart(metadata_16s,"extract_way", label = TRUE, pre = TRUE,"Extract Way")
p1
ggsave("result/Sample_Overview/extact_way_16s.pdf",p1,width = 6,height = 6)
# 16s Sequencing Region
p2 = draw_pie_chart(metadata_16s,"region", label = F, pre = TRUE,"Sequencing Region",color_list = c("#1F4A7A", "#E0F2FF","#8CA9D9", "#2A7777", "#A3BFD4","#7ED4D4", "#B5E8D5"))
p2
ggsave("result/Sample_Overview/Sequencing_Region_16s.pdf",p2,width = 6,height = 4)
# its extract_way
p3 = draw_pie_chart(metadata_its,"extract_way", label = TRUE, pre = TRUE,"Extract Way")
p3
ggsave("result/Sample_Overview/extact_way_its.pdf",p3,width = 6,height = 4)
# its Sequencing Region
p4 = draw_pie_chart(metadata_its,"region", label = TRUE, pre = TRUE,"Region",color_list = c("#1F4A7A", "#E0F2FF"))
p4
ggsave("result/Sample_Overview/Sequencing_Region_its.pdf",p4,width = 6,height = 4)
```


# Construct a species-level phylogenetic tree based on TimeTree5
```{r}
rm(list = ls())
library(dplyr)
library(stringr)
dir.create("planttree")
metadata_16s = read.delim("raw/metadata_16s.txt")
metadata_its = read.delim("raw/metadata_its.txt")
all_plant_tax = unique(rbind(metadata_16s[,c("Genus","Species","timetree")],metadata_its[,c("Genus","Species","timetree")]))
all_plant_timetree = unique(c(metadata_16s$timetree,metadata_its$timetree))
# Create the input file for timetree5
writeLines(all_plant_timetree,"planttree/species_all.txt")
#！！！You need to obtain the tree file by yourself from the TimeTree5 platform！！！
# The obtained tree file (species_all.nwk)  has been placed in the "raw" folder.

if (file.exists("raw/species_all.nwk")) {
  message("Find the target file and continue running...")
  library(ape)
  species_all_tree = read.tree("raw/species_all.nwk")
  #修改叶节点标签
  species_all_tree$tip.label <- all_plant_tax$Species[match(species_all_tree$tip.label, all_plant_tax$timetree)]
  write.tree(species_all_tree, file = "planttree/plant_species_tree_all.nwk")
  #裁剪出16s树
  tree_16s <- keep.tip(species_all_tree, tip = unique(metadata_16s$Species))
  write.tree(tree_16s, file = "planttree/plant_species_tree_16s.nwk")
  #裁剪出its树
  tree_its <- keep.tip(species_all_tree, tip = unique(metadata_its$Species))
  write.tree(tree_its, file = "planttree/plant_species_tree_its.nwk")
  #从Species树中裁剪出一颗Genus树
  #从每个Genus中挑选一个species
  replace_vec_SP <- all_plant_tax[!duplicated(all_plant_tax[, "Genus"]), ]
  plant_Genus_tree_all = keep.tip(species_all_tree, tip = replace_vec_SP$Species)
  plant_Genus_tree_all$tip.label <- replace_vec_SP$Genus[match(plant_Genus_tree_all$tip.label, replace_vec_SP$Species)]
  write.tree(plant_Genus_tree_all, file = "planttree/plant_Genus_tree_all.nwk")
} else {
  stop("请将species_all.txt文件在timetree5平台中注释，导出树文件到planttree文件夹")
}

```






#制作植物分类树和对应总结表
```{r}
rm(list = ls())
library(dplyr)
library(data.table)
library(flextable)
library(ape)
library(itol.toolkit)
library(stringr)
library(tidyr)
library(tibble)
source("Tools.R")
plant_Genus_tree_all = read.tree("planttree/plant_Genus_tree_all.nwk")
tree_label <- plant_Genus_tree_all$tip.label[plant_Genus_tree_all$edge[,2][plant_Genus_tree_all$edge[,2] <= length(plant_Genus_tree_all$tip.label)]]
name_df = data.frame(index = c(1:length(tree_label)),
                     Genus = tree_label)
metadata_16s <- read.delim("raw/metadata_16s.txt")
metadata_16s$type = "16s"
metadata_its <- read.delim("raw/metadata_its.txt")
metadata_its$type = "its"
all_metadata = rbind(metadata_16s,metadata_its)
write.table(all_metadata,file = "result/Sample_Overview/all_metadata.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
country_df = as.data.frame(table(all_metadata$country))
country_df = country_df[order(country_df$Freq, decreasing = TRUE), ]
country_df = country_df[country_df$Var1!="unknown",]
writeLines(paste0('"', country_df$Var1, '"', collapse = ","),"result/Sample_Overview/country.txt")
writeLines(paste0(country_df$Freq, collapse = ","),"result/Sample_Overview/country_Freq.txt")
# 生成两个十六进制颜色之间的渐变色彩
# 参数：color1 起始色（#xxxxxx），color2 结束色（#xxxxxx），n 渐变颜色数量
# 返回：n个十六进制颜色字符向量
p = draw_pie_chart(all_metadata,"country", label = F, pre = TRUE,"Country",limit_n = 15,
                   color_list = c("#3A5A40", "#F2CC8F", "#81B29A", "#3D405B", "#F4A261",
                                  "#099396" ,"#E7D6C2" ,"#547FB6" ,"#C5D8E7" ,"#BF6173",
                                  "#E07A5F", "#90DBF4", "#641E16", "#A7C957", "#D35400"))
p
ggsave("result/Sample_Overview/country_freq_pie.pdf",p,width = 8,height = 6)



result <- all_metadata %>%
  group_by(Order, Genus) %>%
  summarize(
    samples_16s = sum(`type` == "16s", na.rm = TRUE),
    samples_its = sum(`type` == "its", na.rm = TRUE),
    ratio_samples = paste(samples_16s, samples_its, sep = "|"),
    studies_16s = n_distinct(Cite3[`type` == "16s"]),
    studies_its = n_distinct(Cite3[`type` == "its"]),
    ratio_studies = paste(studies_16s, studies_its, sep = "|"),
    references = paste(unique(Cite3), collapse = "; ")
  ) %>%
  ungroup() %>%
  arrange(Order, Genus)

result = merge(name_df,result,by="Genus")
result = result %>% arrange(index)
write.table(result,file = "result/Sample_Overview/sample_summary_table.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
colnames(result)
ft <- flextable(result[ ,c("Genus","Order","ratio_samples","ratio_studies","references")])
all_metadata <- rbind(metadata_16s, metadata_its)

# Set column headers
ft <- set_header_labels(ft,
                        Order = "Order",
                        Genus = "Genus",
                        ratio_samples = "Samples 16S/ITS",
                        ratio_studies = "Studies 16S/ITS",
                        references = "References"
)
# Merge vertically identical Order cells
ft <- merge_v(ft, j = "Order")
# Add thin inner borders for all cells (corrected with officer::fp_border)
ft <- border_inner(ft, border = officer::fp_border(width = 0.5))
# Add outer borders
ft <- border_outer(ft, border = officer::fp_border(width = 1))
ft

#制作树注释文件
#计算每个植物属的richness
sample_taxid_01_16s = fread("standard/sample_taxid_01_16s.txt")
sample_taxid_01_16s = as.data.frame(column_to_rownames(sample_taxid_01_16s,"taxid"))

sample_taxid_01_its = fread("standard/sample_taxid_01_its.txt")
sample_taxid_01_its = as.data.frame(column_to_rownames(sample_taxid_01_its,"taxid"))

long_richness_16s = as.data.frame(colSums(sample_taxid_01_16s))
PG_mean_richness_16s = merge_datafame(long_richness_16s,row_group = metadata_16s[ ,c("SampleID","Genus")],row_method = "mean")
long_richness_16s = merge(long_richness_16s,metadata_16s[ ,c("SampleID","Genus")],by.x = "row.names",by.y = "SampleID",all.x = TRUE)[ ,c(3,2)]
colnames(long_richness_16s)[2] = "richness_16s_log10"
long_richness_16s$richness_16s_log10 = log10(long_richness_16s$richness_16s_log10)

long_richness_its = as.data.frame(colSums(sample_taxid_01_its))
PG_mean_richness_its = merge_datafame(long_richness_its,row_group = metadata_its[ ,c("SampleID","Genus")],row_method = "mean")
long_richness_its = merge(long_richness_its,metadata_its[ ,c("SampleID","Genus")],by.x = "row.names",by.y = "SampleID",all.x = TRUE)[ ,c(3,2)]
colnames(long_richness_its)[2] = "richness_its_log10"
long_richness_its$richness_its_log10 = log10(long_richness_its$richness_its_log10)

dir.create("planttree/all_tree")
annotation = unique(all_metadata[ ,c("Kingdom","Phylum","Class","Order","Family","Genus")])
annotation = merge(annotation,result[ ,-3],by = "Genus")
annotation$lable = paste(annotation$Genus,annotation$ratio_samples)
annotation = annotation %>% arrange(index)
write.table(annotation,file = "planttree/all_tree/all_plant_annotation.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
colorlist <- c(
  "#855B67" ,"#B6BCCA" ,"#C1B5BF" ,"#FDE0E4" ,
  "#78A495" ,"#8A926B" ,"#5D6C57" ,"#6E736D" ,"#E4E2C9",
  "#CFE1D3" ,"#995B70" ,"#846992" ,"#D4C0C2" ,"#F7E09B",
  "#099396" ,"#E7D6C2" ,"#547FB6" ,"#C5D8E7" ,"#BF6173",
  "#3A5A40", "#90DBF4", "#641E16", "#BFEFEE", "#D35400", 
  "#1D3557", "#F1C40F", "#5B2C6F", "#A7C957", "#E63946", 
  "#355070", "#F4A261", "#1E5959", "#CB997E", 
  "#2A9D8F", "#FFE8D6", "#6D597A", "#F9C74F", "#2E4053", 
  "#A8DADC", "#BA4A00", "#7DCEA0", "#424949", "#E5989B",
  "#E07A5F", "#F2CC8F", "#81B29A", "#3D405B", "#F4A261",
  "#E76F51", "#2A9D8F", "#264653", "#F9C74F", "#A8DADC",
  "#E63946", "#F1FAEE", "#457B9D", "#E0A458", "#6D597A",
  "#B5E8D5", "#FFBF69", "#2EC4B6", "#1D3557", "#F6BD60",
  "#8AC6C3", "#E56B6F", "#4A6FA5", "#F5E8C7", "#233D4D"
)
tree <- read.tree("planttree/plant_Genus_tree_all.nwk")
annotation = read.delim("planttree/all_tree/all_plant_annotation.txt")
hub <- create_hub(tree = tree)
#制作Order色环
#配置颜色
unique_Order = unique(annotation$Order)
colorlist_Order = colorlist[1:length(unique_Order)]
Order_color_map = data.frame(Order = unique_Order,
                                   Order_color = colorlist_Order)
create_color_legend(Order_color_map, save_path = "planttree/all_tree_order_legend.pdf", shape = "square", 
                                layout = "col", n = 1, title = "Order", 
                                label_size = 4, title_size = 5, ifsave = TRUE)
create_color_legend(Order_color_map, save_path = "planttree/all_tree_order_legend2.pdf", shape = "circle", 
                                layout = "col", n = 2, title = "Order", 
                                label_size = 4, title_size = 5, ifsave = TRUE)
unit_Order <- create_unit(data = merge(annotation[ ,c("Genus","Order")],Order_color_map,by="Order")[ ,c("Genus","Order","Order_color")],
                                key = "Order",
                                type = "DATASET_COLORSTRIP",
                                tree = tree)
unit_Order@common_themes$legend$horizontal = NULL #调整图例为竖向
# unit_Order@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Order
#添加richness箱线图
unit_richness_16s <- create_unit(data = long_richness_16s,
                       key = "richness_16s_log10",
                       type = "DATASET_BOXPLOT",
                       tree = tree)
unit_richness_16s@specific_themes$basic_plot$dataset_scale = c(0,1,2,3,4)
unit_richness_16s@specific_themes$basic_plot$width = "300"
unit_richness_16s@profile$color = "#355070"
  
hub <- hub + unit_richness_16s
unit_richness_its <- create_unit(data = long_richness_its,
                                 key = "richness_its_log10",
                                 type = "DATASET_BOXPLOT",
                                 tree = tree)
unit_richness_its@specific_themes$basic_plot$dataset_scale = c(0,1,2,3)
unit_richness_its@specific_themes$basic_plot$width = "300"
unit_richness_its@profile$color = "#E0A458"
hub <- hub + unit_richness_its

#添加
unit_lable <- create_unit(data = annotation[ ,c("Genus","lable")],
                          key = "LABELS",
                          type = "LABELS",
                          tree = tree)
hub <- hub + unit_lable

write_hub(hub, "planttree/all_tree")

```









# Data standardization and host species-level integration
# Explanation: We have developed a new merging algorithm, which helps to enhance the representativeness of the data and reduce the interference caused by differences in sample sizes.
```{r}
rm(list = ls())
library(tibble)
library(data.table)
library(tidyr)
library(dplyr)
source("Tools.R")
dir.create("standard")
dir.create("result")
sample_taxid_abundance_16s = fread("raw/sample_taxid_abundance_16s.txt")
sample_taxid_abundance_16s = as.data.frame(column_to_rownames(sample_taxid_abundance_16s,"taxid"))

sample_taxid_abundance_its = fread("raw/sample_taxid_abundance_its.txt")
sample_taxid_abundance_its = as.data.frame(column_to_rownames(sample_taxid_abundance_its,"taxid"))

sample_taxid_01_16s = standard_01(sample_taxid_abundance_16s)
sample_taxid_01_16s = rownames_to_column(sample_taxid_01_16s,var = "taxid")
write.table(sample_taxid_01_16s,file = "standard/sample_taxid_01_16s.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")

sample_taxid_01_its = standard_01(sample_taxid_abundance_its)
sample_taxid_01_its = rownames_to_column(sample_taxid_01_its,var = "taxid")
write.table(sample_taxid_01_its,file = "standard/sample_taxid_01_its.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")

#数据合并
metadata_16s = read.delim("raw/metadata_16s.txt")
metadata_its = read.delim("raw/metadata_its.txt")


# This is a function for merging species tables, which extracts representative species from a classification.
# Merging principle: Firstly, count the occurrence frequency of each species in this group of samples, that is, the prevalence rate.
# If the sample size is greater than or equal to 10 and the number of species in all samples follows a normal distribution, or if the sample size is less than 10, then use the average number of species in all samples as the representative species count T, and extract the species with prevalence up to T as the representative species.
# If the sample size is greater than or equal to 10 and the number of species in all samples does not follow a normal distribution, then use the median number of species in all samples as the richness sampling number T, and extract the species with prevalence up to T as the representative species.
Extract_representative_species = function(indata,group_df,Sample_col = "SampleID",group_level){
  indata = check_and_set_rownames(indata)
  colnames(group_df)[colnames(group_df)==Sample_col] = "SampleID"
  colnames(group_df)[colnames(group_df)==group_level] = "levels"
  group_list = unique(group_df$levels)
  outdata = data.frame(taxid = rownames(indata))
  library(nortest)
  for (i in group_list){
    message(paste0("Currently processing:",i))
    tempdf=indata[ ,c(group_df[group_df$levels==i, ]$SampleID),drop = FALSE]
    prevalence_df = data.frame(taxid = rownames(tempdf),prevalence = rowSums(tempdf))
    sample_num = c(colSums(tempdf))
    print(paste0("The sample size is:",length(sample_num)))
    names(sample_num) = NULL
    if (length(sample_num)>=10){
      ifnormal = ad.test(sample_num)$p.value
      if (ifnormal>0.05){
        print("The number of species is normally distributed, using the average value.")
        sample_mean = as.integer(mean(sample_num))
      }else{
        print("The number of species is not normally distributed, and the median is used.")
        sample_mean = as.integer(median(sample_num, na.rm = FALSE))
      }
    }else{
      print("If the sample size is less than 10, use the average value.")
      sample_mean = as.integer(mean(sample_num))
    }
    result <- data.frame(
      taxid = rownames(tempdf),
      Presence = 0
    )
    high_taxid <- prevalence_df %>%
      arrange(desc(prevalence)) %>% 
      slice_head(n = sample_mean) %>% 
      pull(taxid) %>% 
      as.character()  
    result$Presence[result$taxid%in%high_taxid]=1
    colnames(result)[2] <- i
    outdata = merge(outdata,result,by= "taxid")
  }
  return(outdata)
}
PS_taxid_01_16s = Extract_representative_species(sample_taxid_01_16s,metadata_16s,"SampleID","Species")
write.table(PS_taxid_01_16s,file = "standard/PS_taxid_01_16s.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
PS_taxid_01_its = Extract_representative_species(sample_taxid_01_its,metadata_its,"SampleID","Species")
write.table(PS_taxid_01_its,file = "standard/PS_taxid_01_its.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")

```


#Add popularity rate, generalist-specialist classification and function prediction results to the tax table
```{r}
rm(list = ls())
library(data.table)
library(tibble)
library(tidyr)
library(dplyr)
library(vegan)
library(ggplot2)
source("Tools.R")
PS_taxid_01_16s = read.delim("standard/PS_taxid_01_16s.txt")
prevalence_Species_16s = data.frame(taxid = PS_taxid_01_16s$taxid,
                                    prevalence_s = rowSums(PS_taxid_01_16s[ ,-1]))
taxonomy_16s = read.delim("raw/taxonomy_16s.txt")
taxonomy_16s_extended = merge(taxonomy_16s,prevalence_Species_16s,by = "taxid",all.x = TRUE)


PS_taxid_01_its = read.delim("standard/PS_taxid_01_its.txt")
prevalence_Species_its = data.frame(taxid = PS_taxid_01_its$taxid,
                                    prevalence_s = rowSums(PS_taxid_01_its[ ,-1]))

taxonomy_its = read.delim("raw/taxonomy_its.txt")
taxonomy_its_extended = merge(taxonomy_its,prevalence_Species_its,by = "taxid",all.x = TRUE)




PS_taxid_01_16s = column_to_rownames(PS_taxid_01_16s,"taxid")
PS_taxid_01_16s = PS_taxid_01_16s[rowSums(PS_taxid_01_16s)!=0,]
PS_taxid_01_16s = PS_taxid_01_16s[rownames(PS_taxid_01_16s)!="Unassigned",]


PS_taxid_01_its = column_to_rownames(PS_taxid_01_its,"taxid")
PS_taxid_01_its = PS_taxid_01_its[rowSums(PS_taxid_01_its)!=0,]
PS_taxid_01_its = PS_taxid_01_its[rownames(PS_taxid_01_its)!="Unassigned",]


GOS_result_16s_s = classify_gos(PS_taxid_01_16s,10000)
GOS_result_its_s = classify_gos(PS_taxid_01_its,10000)

GOS_type_16s_s = GOS_result_16s_s[[1]]
colnames(GOS_type_16s_s)[2] = "GOS_s"

GOS_type_its_s = GOS_result_its_s[[1]]
colnames(GOS_type_its_s)[2] = "GOS_s"

taxonomy_16s_extended = merge(taxonomy_16s_extended, GOS_type_16s_s, by = "taxid",all.x = TRUE)

taxonomy_16s_extended$GOS_s[is.na(taxonomy_16s_extended$GOS_s)] = "comm"
taxonomy_its_extended = merge(taxonomy_its_extended, GOS_type_its_s, by = "taxid",all.x = TRUE)
taxonomy_its_extended$GOS_s[is.na(taxonomy_its_extended$GOS_s)] = "comm"
dir.create("result/GOS")
dist_plot = function(dist_df,out_dir){
  #去除双0行
  dist_df = dist_df[dist_df$real_num != 0 | dist_df$pre_num != 0, ]
  dist_df = dist_df[dist_df$real_num>=10,]
  df_long <- pivot_longer(dist_df, cols = c(real_num, pre_num), names_to = "variable", values_to = "value")
  df_long$Frequency = factor(df_long$Frequency)
  # 绘制簇状柱形图
  p = ggplot(df_long, aes(x = Frequency, y = value, fill = variable)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    labs(x = "Prevalence", y = "Number") +
    scale_fill_manual(values = c("real_num" = "#f7e09b", "pre_num" = "#517d7e"), # 自定义颜色
                      labels = c("Predicted", "Realistic")) + # 自定义图例标签
    theme_classic() + 
    theme(
      legend.position = "top" ,
      panel.background = element_blank(), 
      axis.text.x = element_text(size = 14, color = "black"),
      axis.title.x = element_text(size = 14, color = "black", face = "bold"),
      axis.text.y = element_text(size = 14, color = "black", angle = 0), 
      axis.title.y = element_text(size = 14, color = "black", face = "bold"), 
      legend.title = element_text(size = 14, color = "black"), 
      legend.text = element_text(size = 14, color = "black"), 
    )
  ggsave(out_dir,p,width = 10,height = 6)
  return(p)
}
dist_plot(GOS_result_16s_s[[2]],"result/GOS/real_pre_16s_s.pdf")
dist_plot(GOS_result_its_s[[2]],"result/GOS/real_pre_its_s.pdf")


### 汇总功能预测数据到tax表=====================================================
funguild_result = read.delim("raw/funguild_result.txt")
taxonomy_its_extended = merge(taxonomy_its_extended,funguild_result,by.x = "taxid",by.y = "OTU_ID",all.x = TRUE)
taxonomy_its_extended_cleaned = replace_nas(taxonomy_its_extended)
write.table(taxonomy_its_extended_cleaned,file = "standard/taxonomy_its_extended.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")

ko_data_genome = fread("raw/ko_data_genome.txt")
ko_data_genome = column_to_rownames(ko_data_genome,"taxid")
ko_data_genome_01 = ko_data_genome
ko_data_genome_01[ko_data_genome_01 != 0] = 1
ko_sum_genome = data.frame(taxid = rownames(ko_data_genome),
                            ko_sum_genome = rowSums(ko_data_genome),
                            ko_richness_genome = rowSums(ko_data_genome_01))
taxonomy_16s_extended = merge(taxonomy_16s_extended,ko_sum_genome,by = "taxid",all.x = TRUE)

bugbase_result= read.delim("raw/bugbase_result.txt")
taxonomy_16s_extended = merge(taxonomy_16s_extended,bugbase_result,by = "taxid", all.x = TRUE)
taxonomy_16s_extended = taxonomy_16s_extended %>%
  mutate(Aerobic_type = ifelse(Aerobic=="TRUE","Aerobic",ifelse(Anaerobic=="TRUE","Anaerobic",ifelse(Facultatively_Anaerobic=="TRUE","Facultatively_Anaerobic","Unknown")))) %>%
  mutate(Gram_type = ifelse(Gram_Positive=="TRUE","Positive",ifelse(Gram_Negative=="TRUE","Negative","Unknown")))
taxonomy_16s_extended_cleaned <- replace_nas(taxonomy_16s_extended)
write.table(taxonomy_16s_extended_cleaned,file = "standard/taxonomy_16s_extended.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")

table(taxonomy_16s_extended_cleaned$GOS_s)
```
# Calculate the fusion prevalence rate of microbial genera
```{r}
rm(list = ls())
library(tidyr)
library(dplyr)
library(ggplot2)
source("Tools.R")
merged_preavalence_plot = function(indata,xlable,out_dir){
  indata = indata[!rownames(indata)%in%c("Unassigned","Incertae Sedis"),]
  indata = indata[c(1:30),]
  rownames(indata)[which(rownames(indata) == "Burkholderia-Caballeronia-Paraburkholderia")] <- "Burkholderia"
  colnames(indata)[2] = "sum"
  ymin = min(indata$sum)-5
  ymax = max(indata$sum)+3
  p = ggplot(indata, aes(x = reorder(rownames(indata), -sum), y = sum)) +
    geom_col(aes(fill = sum)) + 
    scale_fill_gradient(low = "#93CEDF", high = "#E4A644") + 
    geom_text(aes(label = sum), vjust = -0.5, size = 4) +  
    labs(x = xlable, y = "Merged Prevalence",fill = "Merged\nPrevalence") +
    theme_classic() + 
    coord_cartesian(ylim = c(ymin, ymax))+
    theme(
      panel.background = element_blank(),
      panel.border = element_rect(colour = "black", fill = NA, size = 1.2), 
      axis.text.x = element_text(size = 14, color = "black", angle = -45, hjust = 0, vjust = 1),
      axis.title.x = element_text(size = 14, color = "black", face = "bold"),
      axis.text.y = element_text(size = 14, color = "black", angle = 0), 
      axis.title.y = element_text(size = 14, color = "black", face = "bold"),
      legend.title = element_text(size = 14, color = "black"), 
      legend.text = element_text(size = 12, color = "black"), 
      plot.margin = margin(t = 10,r = 20,b = 10,l = 10,unit = "pt")
    )
  ggsave(out_dir,p,width = 12,height = 5)
  return(p)
}
dir.create("result/merge_prevalence")
metadata_16s = read.delim("raw/metadata_16s.txt")
metadata_its = read.delim("raw/metadata_its.txt")
taxonomy_16s = read.delim("raw/taxonomy_16s.txt")
taxonomy_its = read.delim("raw/taxonomy_its.txt")
PS_taxid_01_16s = read.delim("standard/PS_taxid_01_16s.txt")
PS_taxid_01_its = read.delim("standard/PS_taxid_01_its.txt")

# 计算融合流行率
PS_MG_01_16s = merge_datafame(datafame = PS_taxid_01_16s,
                              row_group = taxonomy_16s[ ,c("taxid","Genus")],
                              row_method = "sum")
PS_MG_01_16s[PS_MG_01_16s != 0] = 1


merged_preavalence_PS_16S = data.frame(Genus = rownames(PS_MG_01_16s),
                                       prevalence_s = rowSums(PS_MG_01_16s))
merged_preavalence_PS_16S = merged_preavalence_PS_16S[order(merged_preavalence_PS_16S$prevalence_s,decreasing = TRUE),]
merged_preavalence_plot(merged_preavalence_PS_16S,"Bacterial Species","result/merge_prevalence/merged_preavalence_PS_16S.pdf")
write.table(merged_preavalence_PS_16S,file = "result/merge_prevalence/merged_preavalence_PS_16S.txt",
            quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")


PS_MG_01_its = merge_datafame(datafame = PS_taxid_01_its,
                              row_group = taxonomy_its[ ,c("taxid","Genus")],
                              row_method = "sum")
PS_MG_01_its[PS_MG_01_its != 0] = 1


merged_preavalence_PS_its = data.frame(Genus = rownames(PS_MG_01_its),
                                       prevalence_s = rowSums(PS_MG_01_its))
merged_preavalence_PS_its = merged_preavalence_PS_its[order(merged_preavalence_PS_its$prevalence_s,decreasing = TRUE),]
merged_preavalence_plot(merged_preavalence_PS_its,"Fungi Species","result/merge_prevalence/merged_preavalence_PS_its.pdf")
write.table(merged_preavalence_PS_its,file = "result/merge_prevalence/merged_preavalence_PS_its.txt",
            quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")


#看看融合流行率属与通才的关系
taxonomy_16s_extended = read.delim("standard/taxonomy_16s_extended.txt")
taxonomy_its_extended = read.delim("standard/taxonomy_its_extended.txt")
topn = 20
merge_top_16s = merged_preavalence_PS_16S[!rownames(merged_preavalence_PS_16S)%in%c("Unassigned","Incertae Sedis"),]
merge_top_16s = merge_top_16s[c(1:topn),]
rownames(merge_top_16s)[which(rownames(merge_top_16s) == "Burkholderia-Caballeronia-Paraburkholderia")] <- "Burkholderia"
merge_top_its = merged_preavalence_PS_its[!rownames(merged_preavalence_PS_its)%in%c("Unassigned","Incertae Sedis"),]
merge_top_its = merge_top_its[c(1:topn),]
g_gensu_16s = unique(taxonomy_16s_extended$Genus[taxonomy_16s_extended$GOS_s=="generalist"])
g_gensu_its = unique(taxonomy_its_extended$Genus[taxonomy_its_extended$GOS_s=="generalist"])
length(intersect(g_gensu_16s,merge_top_16s$Genus))
length(intersect(g_gensu_its,merge_top_its$Genus))

```

# Construction of the evolutionary tree and annotation file for generalists and specialists of microorganisms

# Note: For the 16s sequences, we use the server to build the tree, and the processed tree files are placed in the "raw" folder
# The following code also provides the Linux command for building the evolutionary tree. It is necessary to install software such as muscle, trimal, mafft, and FastTree in advance and then run it in the Linux environment
# The fungal tree was constructed based on the species classification of microorganisms, containing only the topological structure of species classification and not involving the calculation of genetic distances.

```{r}
rm(list = ls())
source("Tools.R")
dir.create("result/GOS/tree_g_16s")
dir.create("result/GOS/tree_s_16s")
dir.create("result/GOS/tree_g_its")
dir.create("result/GOS/tree_s_its")


taxonomy_16s_extended = read.delim("standard/taxonomy_16s_extended.txt")
taxonomy_its_extended = read.delim("standard/taxonomy_its_extended.txt")
annotation_g_16s = taxonomy_16s_extended[taxonomy_16s_extended$GOS_s=="generalist",]
annotation_s_16s = taxonomy_16s_extended[taxonomy_16s_extended$GOS_s=="specialist",]
annotation_g_its = taxonomy_its_extended[taxonomy_its_extended$GOS_s=="generalist",]
annotation_s_its = taxonomy_its_extended[taxonomy_its_extended$GOS_s=="specialist",]

write.table(annotation_g_16s,file = "result/GOS/tree_g_16s/annotation_g_16s.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
write.table(annotation_s_16s,file = "result/GOS/tree_s_16s/annotation_s_16s.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
write.table(annotation_g_its,file = "result/GOS/tree_g_its/annotation_g_its.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
write.table(annotation_s_its,file = "result/GOS/tree_s_its/annotation_s_its.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")

taxid_g_16s = annotation_g_16s$taxid
taxid_s_16s = annotation_s_16s$taxid

library(Biostrings)
silva_sequences <- readDNAStringSet("db/SILVA_138.2_DNA.fa")
get_target_seq(taxid_g_16s,silva_sequences,"result/GOS/tree_g_16s/tree_g_16s.fa")
get_target_seq(taxid_s_16s,silva_sequences,"result/GOS/tree_s_16s/tree_s_16s.fa")
############################################################
# # Use high-performance servers for tree building
# muscle -align tree_g_16s.fa -output tree_g_16s_muscle.fa
# trimal -in tree_g_16s_muscle.fa \
#   -out tree_g_16s_muscle_trimmed.fa \
#   -automated1
# mkdir -p iqtree_g
# iqtree3 \
#   -s tree_g_16s_muscle_trimmed.fa \
#   -m MFP \
#   -B 1000 \
#   -T 8 \
#   --prefix iqtree_g/tree_g_16s
# 
# mafft --auto --thread 100 tree_s_16s.fa > tree_s_16s_mafft.fa
# trimal -in tree_s_16s_mafft.fa \
#   -out tree_s_16s_mafft_trimmed.fa \
#   -automated1
# FastTreeMP -nt -gtr -gamma tree_s_16s_mafft_trimmed.fa > tree_s_16s_FastTree.tree
##########################################################

# Construct a classification tree for fungal species
class_tree(annotation_g_its,"result/GOS/tree_g_its/tree_g_its.nwk","taxid")
class_tree(annotation_s_its,"result/GOS/tree_s_its/tree_s_its.nwk","taxid")

```

# Create the iTol annotation file
# Preparation of bacterial generalist tree annotation file
```{r}
rm(list = ls())
source("Tools.R")
library(itol.toolkit)
library(dplyr)
library(data.table)
library(ape)
library(stringr)
library(tidyr)
#开始构建树注释文件(通才树)
annotation = read.delim("result/GOS/tree_g_16s/annotation_g_16s.txt")
tree <- read.tree("raw/tree_g_16s_iqtree_itol.nwk")
tree_label = tree$tip.label
name_df = data.frame(index = c(1:length(tree_label)),
                     taxid = tree_label)
annotation = merge(name_df,annotation,by="taxid")
annotation = annotation %>% arrange(index)


colorlist1 = c("#99C2C8","#DA7289","#9EB395","#EBD7A5","#C1D2E2",
               "#099396","#016074","#E9BDA0","#9D2227","#EE9B01",
               "#3E4069","#A23F25","#517D7E","#CA820A","#808DAA",
               "#E5A984","#552379","#B24057","#A49BC6","#F7E09B",
               "#93CEDF","#D5ABBF","#846992","#D98267","#73BEA0",
               "#E4A644","#BC9F77","#CB3F48","#A94B32","#8A926B",
               "#9B443B","#7B6D4A","#4B5B74","#C6803A","#CB4047")

colorlist2 = c("#FCB366","#FE816F","#90D7C7","#93BCD0","#AF3F4F",
               "#E5C28C","#CA8BB6","#1A6C92","#EAB46E","#8CC3BD",
               "#8F9EC5","#6C77B5","#138A8E","#4B9D77","#8E60A8",
               "#DC6D73","#B2B896","#E97DAE","#FB7A64","#EDCE35",
               "#99C2C8","#DA7289","#B24057","#EBD7A5","#92D2C1",
               "#4E56AB","#0FBFBC","#E9A564","#ADD565","#ADADAD",
               "#F3BB34","#F8AD9A","#5163B5","#613767","#56B394")

hub <- create_hub(tree = tree)
#制作Phylum色环
#配置颜色
unique_Phylum = unique(annotation$Phylum)
colorlist_Phylum = colorlist2[1:length(unique_Phylum)]
Phylum_color_map = data.frame(Phylum = unique_Phylum,
                             Phylum_color = colorlist_Phylum)
create_color_legend(Phylum_color_map, save_path = "result/GOS/tree_g_16s/Phylum_legend.pdf", shape = "square", 
                                layout = "col", n = 1, title = "Phylum", 
                                label_size = 4, title_size = 6, ifsave = TRUE)
Phylum_color_map$Phylum_color[Phylum_color_map$Phylum=="Other"] = "#eeeeee"
unit_Phylum <- create_unit(data = merge(annotation[ ,c("taxid","Phylum")],Phylum_color_map,by="Phylum")[ ,c("taxid","Phylum","Phylum_color")],
                          key = "Phylum",
                          type = "DATASET_COLORSTRIP",
                          tree = tree)
unit_Phylum@common_themes$legend$horizontal = NULL #调整图例为竖向
# unit_Phylum@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Phylum


#制作Class色环
#配置颜色
unique_Class = unique(annotation$Class)
colorlist_Class = colorlist1[1:length(unique_Class)]
Class_color_map = data.frame(Class = unique_Class,
                             Class_color = colorlist_Class)
create_color_legend(Class_color_map, save_path = "result/GOS/tree_g_16s/Class_legend.pdf", shape = "square", 
                                layout = "row", n = 3, title = "Class", 
                                label_size = 4, title_size = 6, ifsave = TRUE)
Class_color_map$Class_color[Class_color_map$Class=="Other"] = "#eeeeee"
unit_Class <- create_unit(data = merge(annotation[ ,c("taxid","Class")],Class_color_map,by="Class")[ ,c("taxid","Class","Class_color")],
                          key = "Class",
                          type = "DATASET_COLORSTRIP",
                          tree = tree)
unit_Class@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Class@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Class


#制作好氧/厌氧色环
unique_Aerobic_type = sort(unique(annotation$Aerobic_type))

colorlist_Aerobic_type = c("#99C2C8","#D4693F","#E4A644","#eeeeee")
Aerobic_type_color_map = data.frame(Aerobic_type = unique_Aerobic_type,
                                    Aerobic_type_color = colorlist_Aerobic_type)
unit_Aerobic <- create_unit(data = merge(annotation[ ,c("taxid","Aerobic_type")],Aerobic_type_color_map,by="Aerobic_type")[ ,c("taxid","Aerobic_type","Aerobic_type_color")],
                            key = "Aerobic_type", 
                            type = "DATASET_COLORSTRIP", 
                            tree = tree)
unit_Aerobic@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Aerobic@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Aerobic
#制作生物膜色环
unique_Forms_Biofilms = sort(unique(annotation$Forms_Biofilms))
colorlist_Forms_Biofilms = c("#E8C2D9","#846992","#eeeeee")
Forms_Biofilms_color_map = data.frame(Forms_Biofilms = unique_Forms_Biofilms,
                                      Forms_Biofilms_color = colorlist_Forms_Biofilms)
unit_Forms_Biofilms <- create_unit(data = merge(annotation[ ,c("taxid","Forms_Biofilms")],Forms_Biofilms_color_map,by="Forms_Biofilms")[ ,c("taxid","Forms_Biofilms","Forms_Biofilms_color")],
                                   key = "Forms_Biofilms", 
                                   type = "DATASET_COLORSTRIP",
                                   tree = tree)
unit_Forms_Biofilms@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Forms_Biofilms@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Forms_Biofilms
#制作格兰仕阴阳性色环
unique_Gram_type = unique(annotation$Gram_type)
colorlist_Gram_type = c("#A1B1CB","#4B5B74","#eeeeee")
Gram_type_color_map = data.frame(Gram_type = unique_Gram_type,
                                 Gram_type_color = colorlist_Gram_type)
unit_Gram_type <- create_unit(data = merge(annotation[ ,c("taxid","Gram_type")],Gram_type_color_map,by="Gram_type")[ ,c("taxid","Gram_type","Gram_type_color")],
                              key = "Gram_type", 
                              type = "DATASET_COLORSTRIP", 
                              tree = tree)
unit_Gram_type@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Gram_type@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Gram_type

#制作流行度柱状图
prevalence_s = annotation[ ,c("taxid","prevalence_s")]
prevalence_s$prevalence_s[prevalence_s$prevalence_s=="Unknown"] = 0
prevalence_s$prevalence_s = as.numeric(prevalence_s$prevalence_s)
unit_prevalence_s <- create_unit(data = prevalence_s,
                       key = "Prevalence",
                       type = "DATASET_SIMPLEBAR",
                       tree = tree)
unit_prevalence_s@profile$color = "#1A6C92"
unit_prevalence_s@specific_themes$basic_plot$width = 250
unit_prevalence_s@specific_themes$basic_plot$dataset_scale = c(10,20,30,40,50)
hub <- hub + unit_prevalence_s
dir.create("result/GOS/tree_g_16s/itol_annotation")
write_hub(hub, "result/GOS/tree_g_16s/itol_annotation")

# 常见属的小柱状图
library(ggplot2)
library(dplyr)
freq_pre_df = as.data.frame(table(annotation$Genus))
freq_pre_df$pre = freq_pre_df$Freq/sum(freq_pre_df$Freq)*100
freq_pre_df = freq_pre_df[freq_pre_df$Var1!="Incertae Sedis",]
colnames(freq_pre_df)[1] = "Genus"
top10_df <- freq_pre_df %>%
  arrange(desc(pre)) %>% 
  slice_head(n = 10) %>% 
  mutate(Genus = factor(Genus, levels = rev(unique(Genus))))

ggplot(top10_df, aes(x = Genus, y = pre, fill = pre)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = paste0(round(pre, 2), "%")), 
            hjust = -0.1, size = 4, color = "black") +
  scale_fill_gradient(low = "#E0F2FF", high = "#1F4A7A") +
  ggtitle("Top 10 Genus") +
  coord_flip() +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "none",  
    axis.title = element_blank(),  
    axis.line.y = element_blank(),
    axis.ticks.y = element_blank(), 
    axis.text.y = element_text(size = 11),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  expand_limits(y = max(top10_df$pre) * 1.15)
ggsave("result/GOS/tree_g_16s/top10_genus.pdf",width = 3,height = 3)
```

# Preparation of bacterial specialist tree annotation file
```{r}
rm(list = ls())
source("Tools.R")
library(itol.toolkit)
library(dplyr)
library(data.table)
library(ape)
library(stringr)
library(tidyr)
#开始构建树注释文件(通才树)
annotation = read.delim("result/GOS/tree_s_16s/annotation_s_16s.txt")
annotation <- annotation %>%
  group_by(Phylum) %>%
  mutate(Phylum_count = n()) %>%
  ungroup() %>%
  arrange(desc(Phylum_count)) %>%
  mutate(Phylum_rank = dense_rank(desc(Phylum_count))) %>%
  mutate(Phylum_simp = ifelse(Phylum_rank > 20, "Other", Phylum)) %>%
  select(-Phylum_count, -Phylum_rank)
tree <- read.tree("raw/tree_s_16s_fasttree_itol.nwk")
tree_label = tree$tip.label
name_df = data.frame(index = c(1:length(tree_label)),
                     taxid = tree_label)
annotation = merge(name_df,annotation,by="taxid")
annotation = annotation %>% arrange(index)


colorlist2 = c("#FCB366","#FE816F","#90D7C7","#93BCD0","#AF3F4F",
               "#E5C28C","#CA8BB6","#1A6C92","#EAB46E","#8CC3BD",
               "#8F9EC5","#6C77B5","#138A8E","#4B9D77","#8E60A8",
               "#DC6D73","#B2B896","#E97DAE","#FB7A64","#EDCE35",
               "#99C2C8","#DA7289","#B24057","#EBD7A5","#92D2C1",
               "#4E56AB","#0FBFBC","#E9A564","#ADD565","#ADADAD",
               "#F3BB34","#F8AD9A","#5163B5","#613767","#56B394")

hub <- create_hub(tree = tree)
#制作Phylum色环
#配置颜色
unique_Phylum_simp = unique(annotation$Phylum_simp)
unique_Phylum_simp = c(unique_Phylum_simp[unique_Phylum_simp!="Other"],"Other")
colorlist_Phylum_simp = colorlist2[1:length(unique_Phylum_simp)]
Phylum_simp_color_map = data.frame(Phylum_simp = unique_Phylum_simp,
                             Phylum_simp_color = colorlist_Phylum_simp)
Phylum_simp_color_map$Phylum_simp_color[Phylum_simp_color_map$Phylum_simp=="Other"] = "#eeeeee"

create_color_legend(Phylum_simp_color_map, save_path = "result/GOS/tree_s_16s/Phylum_simp_legend.pdf", shape = "square", 
                                layout = "row", n = 3, title = "Phylum_simp", 
                                label_size = 4, title_size = 6, ifsave = TRUE)

unit_Phylum_simp <- create_unit(data = merge(annotation[ ,c("taxid","Phylum_simp")],Phylum_simp_color_map,by="Phylum_simp")[ ,c("taxid","Phylum_simp","Phylum_simp_color")],
                          key = "Phylum",
                          type = "DATASET_COLORSTRIP",
                          tree = tree)
unit_Phylum_simp@common_themes$legend$horizontal = NULL #调整图例为竖向
# unit_Phylum_simp@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Phylum_simp


#制作好氧/厌氧色环
unique_Aerobic_type = sort(unique(annotation$Aerobic_type))

colorlist_Aerobic_type = c("#99C2C8","#D4693F","#E4A644","#eeeeee")
Aerobic_type_color_map = data.frame(Aerobic_type = unique_Aerobic_type,
                                    Aerobic_type_color = colorlist_Aerobic_type)
unit_Aerobic <- create_unit(data = merge(annotation[ ,c("taxid","Aerobic_type")],Aerobic_type_color_map,by="Aerobic_type")[ ,c("taxid","Aerobic_type","Aerobic_type_color")],
                            key = "Aerobic_type", 
                            type = "DATASET_COLORSTRIP", 
                            tree = tree)
unit_Aerobic@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Aerobic@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Aerobic
#制作生物膜色环
unique_Forms_Biofilms = sort(unique(annotation$Forms_Biofilms))
colorlist_Forms_Biofilms = c("#E8C2D9","#846992","#eeeeee")
Forms_Biofilms_color_map = data.frame(Forms_Biofilms = unique_Forms_Biofilms,
                                      Forms_Biofilms_color = colorlist_Forms_Biofilms)
unit_Forms_Biofilms <- create_unit(data = merge(annotation[ ,c("taxid","Forms_Biofilms")],Forms_Biofilms_color_map,by="Forms_Biofilms")[ ,c("taxid","Forms_Biofilms","Forms_Biofilms_color")],
                                   key = "Forms_Biofilms", 
                                   type = "DATASET_COLORSTRIP",
                                   tree = tree)
unit_Forms_Biofilms@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Forms_Biofilms@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Forms_Biofilms
#制作格兰仕阴阳性色环
unique_Gram_type = unique(annotation$Gram_type)
colorlist_Gram_type = c("#4B5B74","#eeeeee","#A1B1CB")
Gram_type_color_map = data.frame(Gram_type = unique_Gram_type,
                                 Gram_type_color = colorlist_Gram_type)
unit_Gram_type <- create_unit(data = merge(annotation[ ,c("taxid","Gram_type")],Gram_type_color_map,by="Gram_type")[ ,c("taxid","Gram_type","Gram_type_color")],
                              key = "Gram_type", 
                              type = "DATASET_COLORSTRIP", 
                              tree = tree)
unit_Gram_type@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Gram_type@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Gram_type

dir.create("result/GOS/tree_s_16s/itol_annotation")
write_hub(hub, "result/GOS/tree_s_16s/itol_annotation")
```

```{r}
# Preparation of fungal generalist tree annotation file
rm(list = ls())
source("Tools.R")
library(itol.toolkit)
library(dplyr)
library(data.table)
library(ape)
library(stringr)
library(tidyr)
#开始构建树注释文件(通才树)
annotation = read.delim("result/GOS/tree_g_its/annotation_g_its.txt")
tree <- read.tree("raw/tree_g_its_itol.nwk")
tree_label = tree$tip.label
name_df = data.frame(index = c(1:length(tree_label)),
                     taxid = tree_label)
annotation = merge(name_df,annotation,by="taxid")
annotation = annotation %>% 
  arrange(index)%>%
  mutate(Saprotrophs = ifelse(general_saprotrophs == "TRUE"|specialized_saprotrophs == "TRUE","TRUE",
                              ifelse(general_saprotrophs == "FALSE" & specialized_saprotrophs == "FALSE","FALSE","Unknown")))

colorlist1 = c("#99C2C8","#DA7289","#9EB395","#EBD7A5","#C1D2E2",
               "#099396","#016074","#E9BDA0","#9D2227","#EE9B01",
               "#3E4069","#A23F25","#517D7E","#CA820A","#808DAA",
               "#E5A984","#552379","#B24057","#A49BC6","#F7E09B",
               "#93CEDF","#D5ABBF","#846992","#D98267","#73BEA0",
               "#E4A644","#BC9F77","#CB3F48","#A94B32","#8A926B",
               "#9B443B","#7B6D4A","#4B5B74","#C6803A","#CB4047")
colorlist2 = c("#FCB366","#FE816F","#90D7C7","#93BCD0","#AF3F4F",
               "#E5C28C","#CA8BB6","#1A6C92","#EAB46E","#8CC3BD",
               "#8F9EC5","#6C77B5","#138A8E","#4B9D77","#8E60A8",
               "#DC6D73","#B2B896","#E97DAE","#FB7A64","#EDCE35",
               "#99C2C8","#DA7289","#B24057","#EBD7A5","#92D2C1",
               "#4E56AB","#0FBFBC","#E9A564","#ADD565","#ADADAD",
               "#F3BB34","#F8AD9A","#5163B5","#613767","#56B394")

hub <- create_hub(tree = tree)

#制作Phylum色环
#配置颜色
unique_Phylum = unique(annotation$Phylum)
unique_Phylum = c(unique_Phylum[unique_Phylum!="Unassigned"],"Unassigned")
colorlist_Phylum = colorlist1[1:length(unique_Phylum)]
Phylum_color_map = data.frame(Phylum = unique_Phylum,
                             Phylum_color = colorlist_Phylum)
Phylum_color_map$Phylum_color[Phylum_color_map$Phylum=="Unassigned"] = "#eeeeee"
create_color_legend(Phylum_color_map, save_path = "result/GOS/tree_g_its/Phylum_legend.pdf", shape = "square", 
                                layout = "row", n = 1, title = "Phylum", 
                                label_size = 4, title_size = 6, ifsave = TRUE)

unit_Phylum <- create_unit(data = merge(annotation[ ,c("taxid","Phylum")],Phylum_color_map,by="Phylum")[ ,c("taxid","Phylum","Phylum_color")],
                          key = "Phylum",
                          type = "DATASET_COLORSTRIP",
                          tree = tree)
unit_Phylum@common_themes$legend$horizontal = NULL #调整图例为竖向
# unit_Phylum@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Phylum


#制作Class色环
#配置颜色
unique_Class = unique(annotation$Class)
unique_Class = c(unique_Class[unique_Class!="Unassigned"],"Unassigned")
colorlist_Class = colorlist2[1:length(unique_Class)]
Class_color_map = data.frame(Class = unique_Class,
                             Class_color = colorlist_Class)
Class_color_map$Class_color[Class_color_map$Class=="Unassigned"] = "#eeeeee"
create_color_legend(Class_color_map, save_path = "result/GOS/tree_g_its/Class_legend.pdf", shape = "square", 
                                layout = "row", n = 2, title = "Class", 
                                label_size = 4, title_size = 6, ifsave = TRUE)

unit_Class <- create_unit(data = merge(annotation[ ,c("taxid","Class")],Class_color_map,by="Class")[ ,c("taxid","Class","Class_color")],
                          key = "Class",
                          type = "DATASET_COLORSTRIP",
                          tree = tree)
unit_Class@common_themes$legend$horizontal = NULL #调整图例为竖向
# unit_Class@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Class
#给颜色加alpha
add_alpha <- function(color, alpha) {
  # 将16进制颜色转为RGB
  rgb_col <- col2rgb(color)
  
  # 返回带alpha的颜色
  return(rgb(rgb_col[1,], rgb_col[2,], rgb_col[3,], 
             alpha = alpha * 255, maxColorValue = 255))
}

#制作mycorrhizal色环
unique_mycorrhizal = rev(sort(unique(annotation$mycorrhizal)))
colorlist_mycorrhizal = c("#eeeeee",add_alpha("#016074",0.6))
mycorrhizal_color_map = data.frame(mycorrhizal = unique_mycorrhizal,
                                 mycorrhizal_color = colorlist_mycorrhizal)
unit_mycorrhizal <- create_unit(data = merge(annotation[ ,c("taxid","mycorrhizal")],mycorrhizal_color_map,by="mycorrhizal")[ ,c("taxid","mycorrhizal","mycorrhizal_color")],
                              key = "Mycorrhizal", 
                              type = "DATASET_COLORSTRIP", 
                              tree = tree)
unit_mycorrhizal@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_mycorrhizal@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_mycorrhizal

#制作plant_associates色环
unique_plant_associates = rev(sort(unique(annotation$plant_associates)))
colorlist_plant_associates = c("#eeeeee","#3E4069",add_alpha("#3E4069",0.4))
plant_associates_color_map = data.frame(plant_associates = unique_plant_associates,
                                   plant_associates_color = colorlist_plant_associates)
unit_plant_associates <- create_unit(data = merge(annotation[ ,c("taxid","plant_associates")],plant_associates_color_map,by="plant_associates")[ ,c("taxid","plant_associates","plant_associates_color")],
                                key = "Plant_associates", 
                                type = "DATASET_COLORSTRIP", 
                                tree = tree)
unit_plant_associates@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_plant_associates@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_plant_associates

#制作Symbiotroph色环
unique_Symbiotroph = rev(sort(unique(annotation$Symbiotroph)))
colorlist_Symbiotroph = c("#eeeeee","#099396",add_alpha("#099396",0.4))
Symbiotroph_color_map = data.frame(Symbiotroph = unique_Symbiotroph,
                                   Symbiotroph_color = colorlist_Symbiotroph)
unit_Symbiotroph <- create_unit(data = merge(annotation[ ,c("taxid","Symbiotroph")],Symbiotroph_color_map,by="Symbiotroph")[ ,c("taxid","Symbiotroph","Symbiotroph_color")],
                                key = "Symbiotroph", 
                                type = "DATASET_COLORSTRIP", 
                                tree = tree)
unit_Symbiotroph@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Symbiotroph@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Symbiotroph

#制作plant_pathogens色环
unique_plant_pathogens = rev(sort(unique(annotation$plant_pathogens)))
colorlist_plant_pathogens = c("#eeeeee","#9D2227",add_alpha("#9D2227",0.4))
plant_pathogens_color_map = data.frame(plant_pathogens = unique_plant_pathogens,
                                 plant_pathogens_color = colorlist_plant_pathogens)
unit_plant_pathogens <- create_unit(data = merge(annotation[ ,c("taxid","plant_pathogens")],plant_pathogens_color_map,by="plant_pathogens")[ ,c("taxid","plant_pathogens","plant_pathogens_color")],
                              key = "Plant_pathogens", 
                              type = "DATASET_COLORSTRIP", 
                              tree = tree)
unit_plant_pathogens@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_plant_pathogens@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_plant_pathogens


#制作Pathotroph色环
unique_Pathotroph = rev(sort(unique(annotation$Pathotroph)))
colorlist_Pathotroph = c("#eeeeee","#BF6074",add_alpha("#BF6074",0.4))
Pathotroph_color_map = data.frame(Pathotroph = unique_Pathotroph,
                                  Pathotroph_color = colorlist_Pathotroph)
unit_Pathotroph <- create_unit(data = merge(annotation[ ,c("taxid","Pathotroph")],Pathotroph_color_map,by="Pathotroph")[ ,c("taxid","Pathotroph","Pathotroph_color")],
                               key = "Pathotroph", 
                               type = "DATASET_COLORSTRIP", 
                               tree = tree)
unit_Pathotroph@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Pathotroph@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Pathotroph

#制作Saprotrophs色环

unique_Saprotrophs = rev(sort(unique(annotation$Saprotrophs)))
colorlist_Saprotrophs = c("#eeeeee","#F3BB34",add_alpha("#F3BB34",0.4))
Saprotrophs_color_map = data.frame(Saprotrophs = unique_Saprotrophs,
                                   Saprotrophs_color = colorlist_Saprotrophs)
unit_Saprotrophs <- create_unit(data = merge(annotation[ ,c("taxid","Saprotrophs")],Saprotrophs_color_map,by="Saprotrophs")[ ,c("taxid","Saprotrophs","Saprotrophs_color")],
                                key = "Saprotrophs", 
                                type = "DATASET_COLORSTRIP", 
                                tree = tree)
unit_Saprotrophs@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Saprotrophs@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Saprotrophs


#制作Saprotroph色环
unique_Saprotroph = rev(sort(unique(annotation$Saprotroph)))
colorlist_Saprotroph = c("#eeeeee","#E9A564",add_alpha("#E9A564",0.4))
Saprotroph_color_map = data.frame(Saprotroph = unique_Saprotroph,
                                  Saprotroph_color = colorlist_Saprotroph)
unit_Saprotroph <- create_unit(data = merge(annotation[ ,c("taxid","Saprotroph")],Saprotroph_color_map,by="Saprotroph")[ ,c("taxid","Saprotroph","Saprotroph_color")],
                               key = "Saprotroph", 
                               type = "DATASET_COLORSTRIP", 
                               tree = tree)
unit_Saprotroph@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Saprotroph@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Saprotroph

#制作流行度柱状图
prevalence_s = annotation[ ,c("taxid","prevalence_s")]
prevalence_s$prevalence_s[prevalence_s$prevalence_s=="Unknown"] = 0
prevalence_s$prevalence_s = as.numeric(prevalence_s$prevalence_s)
unit_prevalence_s <- create_unit(data = prevalence_s,
                       key = "Prevalence",
                       type = "DATASET_SIMPLEBAR",
                       tree = tree)
unit_prevalence_s@profile$color = "#1A6C92"
unit_prevalence_s@specific_themes$basic_plot$width = 250
unit_prevalence_s@specific_themes$basic_plot$dataset_scale = c(10,20,40,60,80)
hub <- hub + unit_prevalence_s

unit_lable <- create_unit(data = annotation[ ,c("taxid","Genus")],
                          key = "LABELS",
                          type = "LABELS",
                          tree = tree)
hub <- hub + unit_lable

dir.create("result/GOS/tree_g_its/itol_annotation")
write_hub(hub, "result/GOS/tree_g_its/itol_annotation")

# 常见属的小柱状图
library(ggplot2)
library(dplyr)
freq_pre_df = as.data.frame(table(annotation$Genus))
freq_pre_df$pre = freq_pre_df$Freq/sum(freq_pre_df$Freq)*100
freq_pre_df = freq_pre_df[freq_pre_df$Var1!="Unassigned",]
colnames(freq_pre_df)[1] = "Genus"
top10_df <- freq_pre_df %>%
  arrange(desc(pre)) %>% 
  slice_head(n = 10) %>% 
  mutate(Genus = factor(Genus, levels = rev(unique(Genus))))

ggplot(top10_df, aes(x = Genus, y = pre, fill = pre)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = paste0(round(pre, 2), "%")), 
            hjust = -0.1, size = 4, color = "black") +
  scale_fill_gradient(low = "#E0F2FF", high = "#1F4A7A") +
  ggtitle("Top 10 Genus") +
  coord_flip() +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "none",  
    axis.title = element_blank(),  
    axis.line.y = element_blank(),
    axis.ticks.y = element_blank(), 
    axis.text.y = element_text(size = 11),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  expand_limits(y = max(top10_df$pre) * 1.15)
ggsave("result/GOS/tree_g_its/top10_genus.pdf",width = 3,height = 3)
```

# Preparation of fungal specialist tree annotation file
```{r}
rm(list = ls())
source("Tools.R")
library(itol.toolkit)
library(dplyr)
library(data.table)
library(ape)
library(stringr)
library(tidyr)
#开始构建树注释文件(通才树)
annotation = read.delim("result/GOS/tree_s_its/annotation_s_its.txt")
tree <- read.tree("raw/tree_s_its_itol.nwk")
tree_label = tree$tip.label
name_df = data.frame(index = c(1:length(tree_label)),
                     taxid = tree_label)
annotation = merge(name_df,annotation,by="taxid")

annotation <- annotation %>%
  group_by(Class) %>%
  mutate(Class_count = n()) %>%
  ungroup() %>%
  arrange(desc(Class_count)) %>%
  mutate(Class_rank = dense_rank(desc(Class_count))) %>%
  mutate(Class_simp = ifelse(Class_rank > 17, "Other", Class)) %>%
  select(-Class_count, -Class_rank)
annotation = annotation %>% 
  arrange(index)%>%
  mutate(Saprotrophs = ifelse(general_saprotrophs == "TRUE"|specialized_saprotrophs == "TRUE","TRUE",
                              ifelse(general_saprotrophs == "FALSE" & specialized_saprotrophs == "FALSE","FALSE","Unknown")))
annotation$Class_simp[annotation$Class_simp=="Unassigned"]="Other"
colorlist1 = c("#99C2C8","#DA7289","#9EB395","#EBD7A5","#C1D2E2",
               "#099396","#016074","#E9BDA0","#9D2227","#EE9B01",
               "#3E4069","#D98267","#517D7E","#CA820A","#808DAA",
               "#E5A984","#552379","#B24057","#A49BC6","#F7E09B",
               "#93CEDF","#D5ABBF","#846992","#A23F25","#73BEA0",
               "#E4A644","#BC9F77","#CB3F48","#A94B32","#8A926B",
               "#9B443B","#7B6D4A","#4B5B74","#C6803A","#CB4047")
colorlist2 = c("#FCB366","#FE816F","#90D7C7","#93BCD0","#AF3F4F",
               "#E5C28C","#CA8BB6","#1A6C92","#EAB46E","#8CC3BD",
               "#8F9EC5","#6C77B5","#138A8E","#4B9D77","#8E60A8",
               "#DC6D73","#B2B896","#E97DAE","#FB7A64","#EDCE35",
               "#99C2C8","#DA7289","#B24057","#EBD7A5","#92D2C1",
               "#4E56AB","#0FBFBC","#E9A564","#ADD565","#ADADAD",
               "#F3BB34","#F8AD9A","#5163B5","#613767","#56B394")

hub <- create_hub(tree = tree)

#制作Phylum色环
#配置颜色
unique_Phylum = unique(annotation$Phylum)
unique_Phylum = c(unique_Phylum[unique_Phylum!="Unassigned"],"Unassigned")
colorlist_Phylum = colorlist1[1:length(unique_Phylum)]
Phylum_color_map = data.frame(Phylum = unique_Phylum,
                             Phylum_color = colorlist_Phylum)
Phylum_color_map$Phylum_color[Phylum_color_map$Phylum=="Unassigned"] = "#eeeeee"
create_color_legend(Phylum_color_map, save_path = "result/GOS/tree_s_its/Phylum_legend.pdf", shape = "square", 
                                layout = "row", n = 2, title = "Phylum", 
                                label_size = 4, title_size = 6, ifsave = TRUE)
freq_phylum = as.data.frame(table(annotation$Phylum))
freq_phylum$pre = freq_phylum$Freq/sum(freq_phylum$Freq)
unit_Phylum <- create_unit(data = merge(annotation[ ,c("taxid","Phylum")],Phylum_color_map,by="Phylum")[ ,c("taxid","Phylum","Phylum_color")],
                          key = "Phylum",
                          type = "DATASET_COLORSTRIP",
                          tree = tree)
unit_Phylum@common_themes$legend$horizontal = NULL #调整图例为竖向
# unit_Phylum@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Phylum


#制作Class_simp色环
#配置颜色
unique_Class_simp = unique(annotation$Class_simp)
unique_Class_simp = c(unique_Class_simp[unique_Class_simp!="Other"],"Other")
colorlist_Class_simp = colorlist2[1:length(unique_Class_simp)]
Class_simp_color_map = data.frame(Class_simp = unique_Class_simp,
                             Class_simp_color = colorlist_Class_simp)
Class_simp_color_map$Class_simp_color[Class_simp_color_map$Class_simp=="Other"] = "#eeeeee"
create_color_legend(Class_simp_color_map, save_path = "result/GOS/tree_s_its/Class_simp_legend.pdf", shape = "square", 
                                layout = "row", n = 3, title = "Class_simp", 
                                label_size = 4, title_size = 6, ifsave = TRUE)
unit_Class_simp <- create_unit(data = merge(annotation[ ,c("taxid","Class_simp")],Class_simp_color_map,by="Class_simp")[ ,c("taxid","Class_simp","Class_simp_color")],
                          key = "Class",
                          type = "DATASET_COLORSTRIP",
                          tree = tree)
unit_Class_simp@common_themes$legend$horizontal = NULL #调整图例为竖向
# unit_Class_simp@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Class_simp
#给颜色加alpha
add_alpha <- function(color, alpha) {
  # 将16进制颜色转为RGB
  rgb_col <- col2rgb(color)
  
  # 返回带alpha的颜色
  return(rgb(rgb_col[1,], rgb_col[2,], rgb_col[3,], 
             alpha = alpha * 255, maxColorValue = 255))
}

#制作mycorrhizal色环
unique_mycorrhizal = rev(sort(unique(annotation$mycorrhizal)))
colorlist_mycorrhizal = c("#eeeeee","#016074",add_alpha("#016074",0.6))
mycorrhizal_color_map = data.frame(mycorrhizal = unique_mycorrhizal,
                                 mycorrhizal_color = colorlist_mycorrhizal)
unit_mycorrhizal <- create_unit(data = merge(annotation[ ,c("taxid","mycorrhizal")],mycorrhizal_color_map,by="mycorrhizal")[ ,c("taxid","mycorrhizal","mycorrhizal_color")],
                              key = "Mycorrhizal", 
                              type = "DATASET_COLORSTRIP", 
                              tree = tree)
unit_mycorrhizal@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_mycorrhizal@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_mycorrhizal

#制作plant_associates色环
unique_plant_associates = rev(sort(unique(annotation$plant_associates)))
colorlist_plant_associates = c("#eeeeee","#3E4069",add_alpha("#3E4069",0.4))
plant_associates_color_map = data.frame(plant_associates = unique_plant_associates,
                                   plant_associates_color = colorlist_plant_associates)
unit_plant_associates <- create_unit(data = merge(annotation[ ,c("taxid","plant_associates")],plant_associates_color_map,by="plant_associates")[ ,c("taxid","plant_associates","plant_associates_color")],
                                key = "Plant_associates", 
                                type = "DATASET_COLORSTRIP", 
                                tree = tree)
unit_plant_associates@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_plant_associates@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_plant_associates

#制作Symbiotroph色环
unique_Symbiotroph = rev(sort(unique(annotation$Symbiotroph)))
colorlist_Symbiotroph = c("#eeeeee","#099396",add_alpha("#099396",0.4))
Symbiotroph_color_map = data.frame(Symbiotroph = unique_Symbiotroph,
                                   Symbiotroph_color = colorlist_Symbiotroph)
unit_Symbiotroph <- create_unit(data = merge(annotation[ ,c("taxid","Symbiotroph")],Symbiotroph_color_map,by="Symbiotroph")[ ,c("taxid","Symbiotroph","Symbiotroph_color")],
                                key = "Symbiotroph", 
                                type = "DATASET_COLORSTRIP", 
                                tree = tree)
unit_Symbiotroph@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Symbiotroph@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Symbiotroph

#制作plant_pathogens色环
unique_plant_pathogens = rev(sort(unique(annotation$plant_pathogens)))
colorlist_plant_pathogens = c("#eeeeee","#9D2227",add_alpha("#9D2227",0.4))
plant_pathogens_color_map = data.frame(plant_pathogens = unique_plant_pathogens,
                                 plant_pathogens_color = colorlist_plant_pathogens)
unit_plant_pathogens <- create_unit(data = merge(annotation[ ,c("taxid","plant_pathogens")],plant_pathogens_color_map,by="plant_pathogens")[ ,c("taxid","plant_pathogens","plant_pathogens_color")],
                              key = "Plant_pathogens", 
                              type = "DATASET_COLORSTRIP", 
                              tree = tree)
unit_plant_pathogens@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_plant_pathogens@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_plant_pathogens


#制作Pathotroph色环
unique_Pathotroph = rev(sort(unique(annotation$Pathotroph)))
colorlist_Pathotroph = c("#eeeeee","#BF6074",add_alpha("#BF6074",0.4))
Pathotroph_color_map = data.frame(Pathotroph = unique_Pathotroph,
                                  Pathotroph_color = colorlist_Pathotroph)
unit_Pathotroph <- create_unit(data = merge(annotation[ ,c("taxid","Pathotroph")],Pathotroph_color_map,by="Pathotroph")[ ,c("taxid","Pathotroph","Pathotroph_color")],
                               key = "Pathotroph", 
                               type = "DATASET_COLORSTRIP", 
                               tree = tree)
unit_Pathotroph@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Pathotroph@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Pathotroph

#制作Saprotrophs色环

unique_Saprotrophs = rev(sort(unique(annotation$Saprotrophs)))
colorlist_Saprotrophs = c("#eeeeee","#F3BB34",add_alpha("#F3BB34",0.4))
Saprotrophs_color_map = data.frame(Saprotrophs = unique_Saprotrophs,
                                   Saprotrophs_color = colorlist_Saprotrophs)
unit_Saprotrophs <- create_unit(data = merge(annotation[ ,c("taxid","Saprotrophs")],Saprotrophs_color_map,by="Saprotrophs")[ ,c("taxid","Saprotrophs","Saprotrophs_color")],
                                key = "Saprotrophs", 
                                type = "DATASET_COLORSTRIP", 
                                tree = tree)
unit_Saprotrophs@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Saprotrophs@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Saprotrophs


#制作Saprotroph色环
unique_Saprotroph = rev(sort(unique(annotation$Saprotroph)))
colorlist_Saprotroph = c("#eeeeee","#E9A564",add_alpha("#E9A564",0.4))
Saprotroph_color_map = data.frame(Saprotroph = unique_Saprotroph,
                                  Saprotroph_color = colorlist_Saprotroph)
unit_Saprotroph <- create_unit(data = merge(annotation[ ,c("taxid","Saprotroph")],Saprotroph_color_map,by="Saprotroph")[ ,c("taxid","Saprotroph","Saprotroph_color")],
                               key = "Saprotroph", 
                               type = "DATASET_COLORSTRIP", 
                               tree = tree)
unit_Saprotroph@common_themes$legend$horizontal = NULL #调整图例为竖向
unit_Saprotroph@specific_themes$strip_label$color_branches = NULL #取消树枝颜色
hub <- hub + unit_Saprotroph

dir.create("result/GOS/tree_s_its/itol_annotation")
write_hub(hub, "result/GOS/tree_s_its/itol_annotation")
```



# Draw a pcoa diagram based on jaccard distance
```{r,fig.show='animate'}
rm(list = ls())
dir.create("result/host_microbiome")
# Load required packages
library(vegan)
library(parallelDist)
library(data.table)
library(ggplot2)
library(dplyr)
library(ggpubr)
library(tibble)
# Import data
sample_taxid_01_16s = fread("standard/sample_taxid_01_16s.txt")
sample_taxid_01_16s = as.data.frame(column_to_rownames(sample_taxid_01_16s,"taxid"))
sample_taxid_01_16s = sample_taxid_01_16s[rownames(sample_taxid_01_16s)!="Unassigned",]

sample_taxid_01_its = fread("standard/sample_taxid_01_its.txt")
sample_taxid_01_its = as.data.frame(column_to_rownames(sample_taxid_01_its,"taxid"))
sample_taxid_01_its = sample_taxid_01_its[rownames(sample_taxid_01_its)!="Unassigned",]

metadata_16s <- read.delim("raw/metadata_16s.txt")
metadata_its <- read.delim("raw/metadata_its.txt")

# 16S PCoA and PERMANOVA
jaccard_dist_16s <- parDist(t(sample_taxid_01_16s),
                        method = "binary",
                        threads = 8)
saveRDS(jaccard_dist_16s, "result/host_microbiome/jaccard_dist_16s.rds")
jaccard_dist_16s <- readRDS("result/host_microbiome/jaccard_dist_16s.rds")

design_16s <- column_to_rownames(metadata_16s, "SampleID")

adonis_16s2 = adonis2(jaccard_dist_16s ~ Species + extract_way + country, 
                     data = design_16s, permutations = 999,
                     method="jaccard", by="margin",parallel = 8)
saveRDS(adonis_16s2, "result/host_microbiome/adonis_16s2.rds")
adonis_16s2 <- readRDS("result/host_microbiome/adonis_16s2.rds")

pcoa_16s <- cmdscale(jaccard_dist_16s, k = 2, eig = TRUE)
pcoa_scores <- as.data.frame(pcoa_16s$points)
colnames(pcoa_scores) <- c("PCoA1", "PCoA2")
pcoa_scores$SampleID <- rownames(pcoa_scores)
pcoa_data_16s <- merge(pcoa_scores, metadata_16s, by = "SampleID", all.x = TRUE)

# Calculate variance explained by the first two axes
eigenvalues <- pcoa_16s$eig
variance_explained <- eigenvalues / sum(eigenvalues) * 100
pc1_var_16s <- round(variance_explained[1], 2)
pc2_var_16s <- round(variance_explained[2], 2)

# ITS PCoA and PERMANOVA
jaccard_dist_its <- parDist(t(sample_taxid_01_its),
                        method = "binary",
                        threads = 8)
saveRDS(jaccard_dist_its, "result/host_microbiome/jaccard_dist_its.rds")
jaccard_dist_its <- readRDS("result/host_microbiome/jaccard_dist_its.rds")

spe_its = t(sample_taxid_01_its)
design_its <- column_to_rownames(metadata_its, "SampleID")


adonis_its2 = adonis2(jaccard_dist_its ~ Species + extract_way + country,
                     data =design_its,permutations = 999,
                     method="jaccard", by="margin",parallel = 8)
saveRDS(adonis_its2, "result/host_microbiome/adonis_its2.rds")
adonis_its2 <- readRDS("result/host_microbiome/adonis_its2.rds")

pcoa_its <- cmdscale(jaccard_dist_its, k = 2, eig = TRUE)
pcoa_scores <- as.data.frame(pcoa_its$points)
colnames(pcoa_scores) <- c("PCoA1", "PCoA2")
pcoa_scores$SampleID <- rownames(pcoa_scores)
pcoa_data_its <- merge(pcoa_scores, metadata_its, by = "SampleID", all.x = TRUE)

# Calculate variance explained by the first two axes
eigenvalues <- pcoa_its$eig
variance_explained <- eigenvalues / sum(eigenvalues) * 100
pc1_var_its <- round(variance_explained[1], 2)
pc2_var_its <- round(variance_explained[2], 2)

# Create a unified color palette for plant genera

colorlist <- c(
  "#855B67" ,"#B6BCCA" ,"#C1B5BF" ,"#FDE0E4" ,
  "#78A495" ,"#8A926B" ,"#5D6C57" ,"#6E736D" ,"#E4E2C9",
  "#CFE1D3" ,"#995B70" ,"#846992" ,"#D4C0C2" ,"#F7E09B",
  "#099396" ,"#E7D6C2" ,"#547FB6" ,"#C5D8E7" ,"#BF6173",
  "#3A5A40", "#90DBF4", "#641E16", "#BFEFEE", "#D35400", 
  "#1D3557", "#F1C40F", "#5B2C6F", "#A7C957", "#E63946", 
  "#355070", "#F4A261", "#1E5959", "#CB997E", 
  "#2A9D8F", "#FFE8D6", "#6D597A", "#F9C74F", "#2E4053", 
  "#A8DADC", "#BA4A00", "#7DCEA0", "#424949", "#E5989B",
  "#E07A5F", "#F2CC8F", "#81B29A", "#3D405B", "#F4A261",
  "#E76F51", "#2A9D8F", "#264653", "#F9C74F", "#A8DADC",
  "#E63946", "#F1FAEE", "#457B9D", "#E0A458", "#6D597A",
  "#B5E8D5", "#FFBF69", "#2EC4B6", "#1D3557", "#F6BD60",
  "#8AC6C3", "#E56B6F", "#4A6FA5", "#F5E8C7", "#233D4D"
)
annotation = read.delim("planttree/all_tree/all_plant_annotation.txt")
unique_Order = unique(annotation$Order)
plant_Order_color = colorlist[1:length(unique_Order)]
names(plant_Order_color) = unique_Order

# Plot 16S PCoA
p1 <- ggplot(pcoa_data_16s, aes(x = PCoA1, y = PCoA2, color = Order)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(x = paste0("PCoA1 (", pc1_var_16s, "%)"),
       y = paste0("PCoA2 (", pc2_var_16s, "%)"),
       color = "Plant Order",
       title = "16S Samples",
       subtitle = paste("Plant species R²=", round((adonis_16s2$R2[1]), 3), "p<", adonis_16s2$`Pr(>F)`[1])) +
  scale_color_manual(values = plant_Order_color) +
  theme_classic() +
  theme(
    legend.position = "none",
    panel.background = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    plot.title = element_text(
      size = 20,
      face = "bold",
      color = "black",
      hjust = 0.5,
      margin = margin(b = 10)
    ),
    plot.subtitle = element_text(
      size = 12,
      color = "black",
      hjust = 0.5,
      lineheight = 1.2,
      margin = margin(b = 15)
    ),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y = element_text(size = 14, color = "black", angle = 0),
    axis.title.y = element_text(size = 14, color = "black", face = "bold"),
    legend.title = element_text(size = 14, color = "black", face = "bold"),
    legend.text = element_text(size = 14, color = "black"),
    plot.margin = margin(t = 10,
                         r = 30,
                         b = 10,
                         l = 10,
                         unit = "pt")
  )

# Plot ITS PCoA
p2 <- ggplot(pcoa_data_its, aes(x = PCoA1, y = PCoA2, color = Order)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(x = paste0("PCoA1 (", pc1_var_its, "%)"),
       y = paste0("PCoA2 (", pc2_var_its, "%)"),
       color = "Plant Order",
       title = "ITS Samples",
       subtitle = paste("Plant species R²=", round((adonis_its2$R2[1]), 3), "p<", adonis_its2$`Pr(>F)`[1])) +
  scale_color_manual(values = plant_Order_color) +
  theme_classic() +
  theme(
    legend.position = "none",
    panel.background = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    plot.title = element_text(
      size = 20,
      face = "bold",
      color = "black",
      hjust = 0.5,
      margin = margin(b = 10)
    ),
    plot.subtitle = element_text(
      size = 12,
      color = "black",
      hjust = 0.5,
      lineheight = 1.2,
      margin = margin(b = 15)
    ),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y = element_text(size = 14, color = "black", angle = 0),
    axis.title.y = element_text(size = 14, color = "black", face = "bold"),
    legend.title = element_text(size = 14, color = "black", face = "bold"),
    legend.text = element_text(size = 14, color = "black"),
    plot.margin = margin(t = 10,
                         r = 30,
                         b = 10,
                         l = 10,
                         unit = "pt")
  )

# Combine plots with a shared legend
combined_plot <- ggarrange(p1, p2,
                           ncol = 2,
                           common.legend = TRUE,
                           legend = "right")

combined_plot
ggsave("result/host_microbiome/16s_its_PCoA__jaccard2.pdf", plot = combined_plot, width = 13.6, height = 5.5)

```

# Use PERMANOVA and VPA analyses to test the contribution of host factors to the community results
```{r,fig.show='animate'}
rm(list = ls())
library(ggplot2)
library(dplyr)
library(vegan)
library(tidyr)

design_16s <- read.delim("raw/metadata_16s.txt",row.names = 1)
design_its <- read.delim("raw/metadata_its.txt",row.names = 1)
# vpa
jaccard_dist_16s <- readRDS("result/host_microbiome/jaccard_dist_16s.rds")
jaccard_dist_its <- readRDS("result/host_microbiome/jaccard_dist_its.rds")

vpa_result_16s <- varpart(jaccard_dist_16s, 
                      ~ Species, 
                      ~ extract_way, 
                      ~ country, 
                      data = design_16s)
pdf("result/host_microbiome/vpa_plot_16s.pdf", width = 7, height = 6) 
plot(vpa_result_16s, 
     digits = 2, 
     Xnames = c("Species", "Extract_way", "Country"), 
     bg = c("#355070", "#264653", "#81B29A"))
dev.off()


vpa_result_its <- varpart(jaccard_dist_its, 
                          ~ Species, 
                          ~ extract_way, 
                          ~ country, 
                          data = design_its)
pdf("result/host_microbiome/vpa_plot_its.pdf", width = 7, height = 6)
plot(vpa_result_its, 
     digits = 2, 
     Xnames = c("Species", "Extract_way", "Country"), 
     bg = c("#355070", "#264653", "#81B29A"))
dev.off()

# PERMANOVA

prepare_data <- function(df, label) {
  df %>%
    as.data.frame() %>%
    mutate(Factor = rownames(.)) %>%
    filter(Factor %in% c("Species", "country", "extract_way")) %>%
    mutate(Community = label) %>%
    select(Community, Factor, R2)
}


adonis_16s <- readRDS("result/host_microbiome/adonis_16s2.rds")
adonis_its <- readRDS("result/host_microbiome/adonis_its2.rds")

plot_df_16s <- prepare_data(adonis_16s, "16S (Bacteria)")
plot_df_its <- prepare_data(adonis_its, "ITS (Fungi)")

final_plot_df <- rbind(plot_df_16s, plot_df_its)
final_plot_df$R2_percent <- final_plot_df$R2 * 100

PERMANOVA = ggplot(final_plot_df, aes(x = Factor, y = R2_percent, fill = Community)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = sprintf("%.2f%%", R2_percent)), 
            position = position_dodge(width = 0.8), 
            vjust = -0.5, size = 3.5, fontface = "bold") +
  scale_fill_manual(values = c("16S (Bacteria)" = "#4E79A7", "ITS (Fungi)" = "#81B29A")) +
  labs(title = "Comparison of Community Variation Explained (R²)",
       x = "Explanatory Factors",
       y = "Variance Explained (R² %)",
       fill = "Community Type") +
  scale_y_continuous(limits = c(0, max(final_plot_df$R2_percent) + 2)) +
  theme_bw() +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.subtitle = element_text(hjust = 0.5, color = "grey30"),
    axis.text = element_text(size = 11, color = "black"),
    legend.position = "top"
  )
ggsave("result/host_microbiome/PERMANOVA.pdf",plot = PERMANOVA, width = 5, height = 5)
PERMANOVA
```

# The relationship between the genetic development distance of the host and the Jaccard distance of the seed community
```{r}
rm(list = ls())
library(ape)
library(vegan)
library(ggplot2)
library(ggpubr)
library(patchwork)

#计算平均距离的函数
mean_dist = function(indata,calssify_df){
  colnames(calssify_df) = c("SampleID","groups")
  common_list = intersect(rownames(indata),calssify_df$SampleID)
  calssify_df = calssify_df[calssify_df$SampleID%in%common_list,]
  indata = indata[common_list,common_list]
  group_list = unique(calssify_df$groups)
  group_list_num = length(group_list)
  group_matrix = matrix(data = 1,ncol = group_list_num,nrow = group_list_num)
  colnames(group_matrix) = group_list
  rownames(group_matrix) = group_list
  for (i in 1:group_list_num){
    row_group = group_list[i]
    row_id = calssify_df$SampleID[calssify_df$groups==row_group]
    for (j in 1:group_list_num){
      col_group = group_list[j]
      col_id = calssify_df$SampleID[calssify_df$groups==col_group]
      temp_dist_matrix = indata[row_id,col_id]
      mean_dist = mean(temp_dist_matrix)
      group_matrix[i,j]=mean_dist
    }
  }
  return(group_matrix)
}

tree_16s <- read.tree("planttree/plant_species_tree_16s.nwk")
tree_matrix_16s <- cophenetic(tree_16s)
#计算微生物距离
metadata_16s = read.delim("raw/metadata_16s.txt")
jaccard_dist_16s <- readRDS("result/host_microbiome/jaccard_dist_16s.rds")
jaccard_matrix_16s = as.matrix(jaccard_dist_16s)
species_matrix_16s = mean_dist(jaccard_matrix_16s,metadata_16s[ ,c("SampleID","Species")])
species_matrix_16s = species_matrix_16s[rownames(tree_matrix_16s),rownames(tree_matrix_16s)]

tree_its <- read.tree("planttree/plant_species_tree_its.nwk")
tree_matrix_its <- cophenetic(tree_its)
metadata_its = read.delim("raw/metadata_its.txt")
jaccard_dist_its <- readRDS("result/host_microbiome/jaccard_dist_its.rds")
jaccard_matrix_its = as.matrix(jaccard_dist_its)
species_matrix_its = mean_dist(jaccard_matrix_its,metadata_its[ ,c("SampleID","Species")])
species_matrix_its = species_matrix_its[rownames(tree_matrix_its),rownames(tree_matrix_its)]
# Mantel
mantel_result_16s <- mantel(tree_matrix_16s, species_matrix_16s, method = "pearson", permutations = 999)
mantel_16s_r = mantel_result_16s$statistic
mantel_16s_p = mantel_result_16s$signif
mantel_result_its <- mantel(tree_matrix_its, species_matrix_its, method = "pearson", permutations = 999)
mantel_its_r = mantel_result_its$statistic
mantel_its_p = mantel_result_its$signif
# MRM
mrm_16s <- ecodist::MRM(as.dist(species_matrix_16s) ~ as.dist(tree_matrix_16s), nperm = 999)
mrm_16s_r2 = mrm_16s$r.squared[1]
mrm_16s_r2p = mrm_16s$r.squared[2]
mrm_16s_F = mrm_16s$F.test[1]
mrm_16s_Fp = mrm_16s$F.test[2]
mrm_its <- ecodist::MRM(as.dist(species_matrix_its) ~ as.dist(tree_matrix_its), nperm = 999)
mrm_its_r2 = mrm_its$r.squared[1]
mrm_its_r2p = mrm_its$r.squared[2]
mrm_its_F = mrm_its$F.test[1]
mrm_its_Fp = mrm_its$F.test[2]

# dbRDA
pcoa_tree_16s <- cmdscale(tree_matrix_16s, k = 5, eig = TRUE)
tree_vectors_16s <- as.data.frame(pcoa_tree_16s$points)
dbrda_16s <- capscale(species_matrix_16s ~ ., data = tree_vectors_16s)
dbRDA_16s_all = anova(dbrda_16s, permutations = 999)
dbRDA_16s_terms = anova(dbrda_16s, by = "terms", permutations = 999)

pcoa_tree_its <- cmdscale(tree_matrix_its, k = 5, eig = TRUE)
tree_vectors_its <- as.data.frame(pcoa_tree_its$points)
dbrda_its <- capscale(species_matrix_its ~ ., data = tree_vectors_its)
dbRDA_its_all = anova(dbrda_its, permutations = 999)
dbRDA_its_terms = anova(dbrda_its, by = "terms", permutations = 999)

#结果可视化
col_16s <- "#4E79A7"
col_its <- "#81B29A"


# Mantel+MRM
get_df <- function(tree_mat, species_mat){
  idx <- upper.tri(tree_mat)
  data.frame(
    phylo = tree_mat[idx],
    micro = species_mat[idx]
  )
}

df_16s <- get_df(tree_matrix_16s, species_matrix_16s)
df_its <- get_df(tree_matrix_its, species_matrix_its)

format_p <- function(p){
  if(p < 0.01) return("<0.01")
  sprintf("%.3f", p)
}

label_16s <- paste0(
  "Mantel r = ", round(mantel_16s_r, 3), "\n",
  "p ", format_p(mantel_16s_p), "\n\n",
  "MRM R² = ", round(mrm_16s_r2, 3), "\n",
  "p ", format_p(mrm_16s_r2p)
)

label_its <- paste0(
  "Mantel r = ", round(mantel_its_r, 3), "\n",
  "p ", format_p(mantel_its_p), "\n\n",
  "MRM R² = ", round(mrm_its_r2, 3), "\n",
  "p ", format_p(mrm_its_r2p)
)

make_plot <- function(df, color, title, label){

  ggplot(df, aes(x = phylo, y = micro)) +
    geom_point(color = color, alpha = 0.2, size = 1.2) +
    
    # 回归线
    geom_smooth(method = "lm", color = color, se = FALSE, linewidth = 1.2) +
    
    # 主题
    theme_bw(base_size = 13) +
    theme(
      panel.grid = element_blank(),
      plot.title = element_text(face = "bold", size = 14),
      axis.title = element_text(size = 12)
    ) +
    
    labs(
      title = title,
      x = "Host phylogenetic distance",
      y = "Microbial community dissimilarity (Jaccard)"
    ) +
    
    # 添加统计信息
    annotate(
      "text",
      x = Inf, y = -Inf,
      label = label,
      hjust = 1.05, vjust = -0.2,
      size = 4
    )
}


p1 <- make_plot(
  df_16s,
  col_16s,
  "16S (Bacterial community)",
  label_16s
)

p2 <- make_plot(
  df_its,
  col_its,
  "ITS (Fungal community)",
  label_its
)


final_plot_mantel_mrm <- p1 + p2 + plot_layout(ncol = 2)
final_plot_mantel_mrm


# MRM可视化
get_r2 <- function(dbrda_obj){
  eig <- dbrda_obj$CCA$eig
  tot <- dbrda_obj$tot.chi
  sum(eig) / tot
}


get_p <- function(anova_res){
  anova_res$`Pr(>F)`[1]
}

label_16s <- paste0(
  "dbRDA R² = ", round(get_r2(dbrda_16s), 3), "\n",
  "p ", ifelse(get_p(dbRDA_16s_all) < 0.01, "<0.01", round(get_p(dbRDA_16s_all), 3))
)

label_its <- paste0(
  "dbRDA R² = ", round(get_r2(dbrda_its), 3), "\n",
  "p ", ifelse(get_p(dbRDA_its_all) < 0.01, "<0.01", round(get_p(dbRDA_its_all), 3))
)


scores_16s <- as.data.frame(scores(dbrda_16s, display = "sites"))
scores_its <- as.data.frame(scores(dbrda_its, display = "sites"))

make_dbrda_plot <- function(df, color, title, label){

  ggplot(df, aes(x = CAP1, y = CAP2)) +
    
    # 点
    geom_point(color = color, alpha = 0.7, size = 2) +
    
    # 零轴线（更专业）
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey70") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey70") +
    
    # 主题
    theme_bw(base_size = 13) +
    theme(
      panel.grid = element_blank(),
      plot.title = element_text(face = "bold", size = 14),
      axis.title = element_text(size = 12)
    ) +
    
    labs(
      title = title,
      x = "CAP1",
      y = "CAP2"
    ) +
    
    # 标注统计结果
    annotate(
      "text",
      x = Inf, y = Inf,
      label = label,
      hjust = 1.1, vjust = 1.5,
      size = 4
    )
}

p3 <- make_dbrda_plot(
  scores_16s,
  col_16s,
  "dbRDA (16S bacterial community)",
  label_16s
)

p4 <- make_dbrda_plot(
  scores_its,
  col_its,
  "dbRDA (ITS fungal community)",
  label_its
)


final_plot_dbrda <- p3 + p4 + plot_layout(ncol = 2)
final_plot_dbrda

ggsave("result/host_microbiome/Mantel_MRM.pdf",final_plot_mantel_mrm,height = 4,width = 8)
ggsave("result/host_microbiome/dbRDA.pdf",final_plot_dbrda,height = 6,width = 12)
```


#Trait shifts with prevalence
#16S samples
```{r}
rm(list = ls())
source("Tools.R")
library(dplyr)
library(ggplot2)
library(tidyr)
library(reshape2)
library(forcats)
dir.create("result/Trait_shifts")
taxonomy_16s_extended = read.delim("standard/taxonomy_16s_extended.txt")
#The relationship between the taxonomic-level species composition and prevalence_s of the 16s sample
prevalence_s_Class_16s = taxonomy_16s_extended[ ,c("prevalence_s","Class")]
prevalence_s_Class_16s[prevalence_s_Class_16s$prevalence_s >= 25,]$prevalence_s = 25

make_mutibar(datain = prevalence_s_Class_16s,
             out_dir = "result/Trait_shifts/Class_levels_change_16s.pdf",
             x_name = "Prevalence_s",
             y_name = "Proportion of Taxid",
             legend_name = "Class",
             limit_num = 6,
             width = 11.2,
             height = 5,
             other_list = c("Unassigned","Incertae Sedis"),
             colorlist = c("#FCB366", "#FE816F", "#90D7C7","#93BCD0", "#AF3F4F", 
                           "#F3BB34", "#F8AD9A", "#5163B5", "#613767", "#56B394",
                           "#DC6D73", "#B2B896", "#E97DAE","#FB7A64","#EDCE35", 
                           "#E5C28C", "#CA8BB6", "#1A6C92", "#EAB46E", "#8CC3BD", 
                           "#4E56AB","#0FBFBC", "#E9A564", "#ADD565", "#ADADAD", 
                           "#8F9EC5", "#6C77B5", "#138A8E", "#4B9D77", "#8E60A8")
             )
Class_test = Calculate_correlation(prevalence_s_Class_16s,"prevalence_s","Class")
class_prop = as.data.frame(table(prevalence_s_Class_16s$Class))
main_class = class_prop$Var1[class_prop$Freq>500]
Class_test = Class_test[Class_test$Variable%in%main_class,]
print(Class_test)
#----------------------------------------------------------------------

#The relationship between the aerobic/anaerobic conditions of the 16s samples and the prevalence_s rate
prevalence_s_Aerobic_type_16s = taxonomy_16s_extended[ ,c("prevalence_s","Aerobic_type")]
prevalence_s_Aerobic_type_16s <- prevalence_s_Aerobic_type_16s %>%
  replace(is.na(.), "Unknown")
prevalence_s_Aerobic_type_16s[prevalence_s_Aerobic_type_16s$prevalence_s >= 25,]$prevalence_s = 25
make_mutibar(datain = prevalence_s_Aerobic_type_16s,
             out_dir = "result/Trait_shifts/prevalence_s_Aerobic_type_change_16s.pdf",
             x_name = "Prevalence_s",
             y_name = "Proportion of Taxid",
             legend_name = "Aerobic_type",
             limit_num = 10,
             width = 11.2,
             height = 5,
             other_list = c("Unassigned","Incertae Sedis","Unknown"),
             colorlist = c("#99C2C8","#D4693F","#E4A644","#eeeeee")
)
Aerobic_type_test = Calculate_correlation(prevalence_s_Aerobic_type_16s,"prevalence_s","Aerobic_type")
print(Aerobic_type_test)
#--------------------------------------------------------------------------


#The relationship between the biofilm-forming potential of 16s samples and their prevalence_s rate
prevalence_s_Forms_Biofilms_16s = taxonomy_16s_extended[ ,c("prevalence_s","Forms_Biofilms")]
prevalence_s_Forms_Biofilms_16s <- prevalence_s_Forms_Biofilms_16s %>%
  replace(is.na(.), "Unknown")
prevalence_s_Forms_Biofilms_16s[prevalence_s_Forms_Biofilms_16s$prevalence_s >= 25,]$prevalence_s = 25
make_mutibar(datain = prevalence_s_Forms_Biofilms_16s,
             out_dir = "result/Trait_shifts/prevalence_s_Forms_Biofilms_change_16s.pdf",
             x_name = "Prevalence_s",
             y_name = "Proportion of Taxid",
             legend_name = "Forms_Biofilms",
             limit_num = 10,
             width = 10,
             height = 5,
             other_list = c("Unassigned","Incertae Sedis","Unknown"),
             colorlist = c("#846992","#E8C2D9","#eeeeee")
)
Forms_Biofilms_test = Calculate_correlation(prevalence_s_Forms_Biofilms_16s,"prevalence_s","Forms_Biofilms")
print(Forms_Biofilms_test)
#--------------------------------------------------------------------------------



#The relationship between the Gram-positive and Gram-negative properties of the 16s samples and their prevalence_s rates
prevalence_s_Gram_type_16s = taxonomy_16s_extended[ ,c("prevalence_s","Gram_type")]
prevalence_s_Gram_type_16s <- prevalence_s_Gram_type_16s %>%
  replace(is.na(.), "Unknown")
prevalence_s_Gram_type_16s[prevalence_s_Gram_type_16s$prevalence_s >= 25,]$prevalence_s = 25
make_mutibar(datain = prevalence_s_Gram_type_16s,
             out_dir = "result/Trait_shifts/prevalence_s_Gram_type_change_16s.pdf",
             x_name = "Prevalence_s",
             y_name = "Proportion of Taxid",
             legend_name = "Gram_type",
             limit_num = 10,
             width = 10,
             height = 5,
             other_list = c("Unassigned","Incertae Sedis","Unknown"),
             colorlist = c("#4B5B74","#A1B1CB","#eeeeee")
)
Gram_type_test = Calculate_correlation(prevalence_s_Gram_type_16s,"prevalence_s","Gram_type")
print(Gram_type_test)
```

## As popularity increases, the changes in genetic diversity occur.
```{r,fig.show='animate'}
rm(list = ls())
library(ggplot2)
library(ggpubr)
library(dplyr)
cds_data = read.delim("raw/cds_data.txt")
cds_data$Genome_size = (cds_data$Genome_size)/1000000


taxonomy_16s_extended <- read.delim("standard/taxonomy_16s_extended.txt")
taxonomy_16s_extended = merge(taxonomy_16s_extended,cds_data,by = "taxid",all.x = TRUE)
value_prevalence_maxmin = function(indata, Target_name, max_prevalence, out_dir) {
  out_dir = paste0("result/Trait_shifts/", colnames(indata)[1], "_", colnames(indata)[2], ".pdf")
  indata <- na.omit(indata)
  colnames(indata) = c("prevalence", "Target_value")
  indata$prevalence[indata$prevalence >= max_prevalence] = max_prevalence
  indata = indata[indata$Target_value != "Unknown", ]
  indata = indata[indata$Target_value != -1, ]
  indata$Target_value = as.numeric(indata$Target_value)
  
  # 数据聚合
  df_summary <- indata %>%
    group_by(prevalence) %>%
    summarise(Target_value_max = max(Target_value, na.rm = TRUE),
              Target_value_min = min(Target_value, na.rm = TRUE))
  
  # 对min进行拟合
  lm_model_min <- lm(Target_value_min ~ prevalence, data = df_summary)
  a2 <- coef(lm_model_min)[1] # 截距
  b2 <- coef(lm_model_min)[2] # 斜率
  lable_min <- paste0("Min: y = ", round(a2, 2), " + ", round(b2, 2), "x\n", "R² = ", round(summary(lm_model_min)$r.squared, 3))
  
  # 对max进行拟合
  lm_model_max <- lm(Target_value_max ~ prevalence, data = df_summary)
  a1 <- coef(lm_model_max)[1] # 截距
  b1 <- coef(lm_model_max)[2] # 斜率
  lable_max <- paste0("Max: y = ", round(a1, 2), " + ", round(b1, 2), "x\n", "R² = ", round(summary(lm_model_max)$r.squared, 3))
  
  # --- 新增：计算交点坐标 ---
  # 公式: a1 + b1*x = a2 + b2*x => x = (a2 - a1) / (b1 - b2)
  if (b1 != b2) {
    inter_x <- (a2 - a1) / (b1 - b2)
    inter_y <- a1 + b1 * inter_x
    lable_inter <- paste0("Intersection: (", round(inter_x, 2), ", ", round(inter_y, 2), ")")
  } else {
    lable_inter <- "Intersection: None (Parallel)"
  }
  # -------------------------

  # 创建完整图表
  final_plot <- ggplot() +
    geom_ribbon(data = df_summary, aes(x = prevalence, ymin = Target_value_min, ymax = Target_value_max),
                fill = "grey", alpha = 0.15) +
    geom_point(data = indata, aes(x = prevalence, y = Target_value, color = Target_value),
               alpha = 0.6, size = 1.5) + 
    scale_color_gradient(low = "#93CEDF", high = "#D4693F", name = Target_name) +
    geom_line(data = df_summary, aes(x = prevalence, y = Target_value_max), color = "#D4693F", size = 1) +
    geom_line(data = df_summary, aes(x = prevalence, y = Target_value_min), color = "#93CEDF", size = 1) +
    geom_smooth(data = df_summary, aes(x = prevalence, y = Target_value_min), 
                method = "lm", formula = y ~ x, se = FALSE, color = "#93CEDF", linetype = "dashed") +
    geom_smooth(data = df_summary, aes(x = prevalence, y = Target_value_max), 
                method = "lm", formula = y ~ x, se = FALSE, color = "#D4693F", linetype = "dashed") +
    # 修改处：在末尾添加交点坐标显示
    annotate("text", x = Inf, y = Inf, 
             label = paste(lable_max, lable_min, lable_inter, sep = "\n"),
             hjust = 1.1, vjust = 1.1, size = 5) +
    labs(x = "Prevalence", y = Target_name) +
    theme_bw() + 
    theme(
      panel.background = element_blank(),
      panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
      plot.title = element_text(size = 20, face = "bold", color = "black", hjust = 0.5, margin = margin(b = 10)),
      plot.subtitle = element_text(size = 12, color = "black", hjust = 0.5, lineheight = 1.2, margin = margin(b = 15)),
      axis.text.x = element_text(size = 14, color = "black"),
      axis.title.x = element_text(size = 14, color = "black", face = "bold"), 
      axis.text.y = element_text(size = 14, color = "black", angle = 0), 
      axis.title.y = element_text(size = 14, color = "black", face = "bold"),
      legend.title = element_text(size = 14, color = "black"),
      legend.text = element_text(size = 12, color = "black"),
      plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
    )
    
  ggsave(out_dir, plot = final_plot, width = 10, height = 5, units = "in")
  return(final_plot)
}

# value_prevalence_maxmin(taxonomy_16s_extended[ ,c("prevalence_g","ko_richness_merge")],"Unique KO richness",20)
# value_prevalence_maxmin(taxonomy_16s_extended[ ,c("prevalence_g","CDS")],"CDS",20)
value_prevalence_maxmin(taxonomy_16s_extended[ ,c("prevalence_s","ko_richness_genome")],"KO richness",25)
value_prevalence_maxmin(taxonomy_16s_extended[ ,c("prevalence_s","CDS")],"CDS",25)
value_prevalence_maxmin(taxonomy_16s_extended[ ,c("prevalence_s","Genome_size")],"Genome size (mbp)",25)
value_prevalence_maxmin(taxonomy_16s_extended[ ,c("prevalence_s","GC_content")],"GC content (%)",25)


# value_prevalence_maxmin(taxonomy_16s_extended[ ,c("prevelence_weight","ko_richness_genome")],"KO richness",25)
# value_prevalence_maxmin(taxonomy_16s_extended[ ,c("prevalence_s","CDS")],"CDS",25)
# value_prevalence_maxmin(taxonomy_16s_extended[ ,c("prevalence_s","Genome_size")],"Genome size (mbp)",25)
# value_prevalence_maxmin(taxonomy_16s_extended[ ,c("prevalence_s","GC_content")],"GC content (%)",25)




envelope_intersection <- function(indata, x_col, y_col, n = 1000) {
  # 提取列
  x <- indata[[x_col]]
  y <- indata[[y_col]]
  
  # 存储每次的交点
  inter_x <- numeric(n)
  inter_y <- numeric(n)
  
  for (i in seq_len(n)) {
    # 随机打乱 y 列
    y_shuffled <- sample(y)
    temp <- setNames(data.frame(x, y_shuffled), c(x_col, y_col))
    
    # 构建上下包络点
    form <- as.formula(paste(y_col, "~", x_col))
    upper <- aggregate(form, data = temp, FUN = max, na.rm = TRUE)
    lower <- aggregate(form, data = temp, FUN = min, na.rm = TRUE)
    
    # 只有至少两个点才能做线性回归
    if (nrow(upper) >= 2 && nrow(lower) >= 2) {
      lm_up <- lm(form, data = upper)
      a_up <- coef(lm_up)[2]
      b_up <- coef(lm_up)[1]
      
      lm_low <- lm(form, data = lower)
      a_low <- coef(lm_low)[2]
      b_low <- coef(lm_low)[1]
      
      # 如果两线不平行，计算交点
      if (abs(a_up - a_low) > 1e-12) {
        x_int <- (b_low - b_up) / (a_up - a_low)
        y_int <- a_up * x_int + b_up
        inter_x[i] <- x_int
        inter_y[i] <- y_int
      } else {
        inter_x[i] <- NA
        inter_y[i] <- NA
      }
    } else {
      inter_x[i] <- NA
      inter_y[i] <- NA
    }
  }
  
  # 计算有效交点的统计量
  valid <- !is.na(inter_x)
  n_valid <- sum(valid)
  
  if (n_valid > 0) {
    mean_x <- mean(inter_x, na.rm = TRUE)
    sd_x   <- sd(inter_x, na.rm = TRUE)
    mean_y <- mean(inter_y, na.rm = TRUE)
    sd_y   <- sd(inter_y, na.rm = TRUE)
  } else {
    mean_x <- NA
    sd_x   <- NA
    mean_y <- NA
    sd_y   <- NA
  }
  
  # 返回结果表格
  result <- data.frame(
    mean_x = mean_x,
    sd_x   = sd_x,
    mean_y = mean_y,
    sd_y   = sd_y,
    n_valid = n_valid,
    n_total = n,
    stringsAsFactors = FALSE
  )
  
  return(result)
}
#进行1000次随机包络分析然后计算平均交点位置
indata = taxonomy_16s_extended[ ,c("prevalence_s","CDS","ko_richness_genome","Genome_size","GC_content")]
indata = na.omit(indata)
colnames(indata) = c("prevalence","CDS","KO","Genome_size","GC_content")
CDS_TEST = envelope_intersection(indata,"prevalence","CDS",1000)
KO_TEST = envelope_intersection(indata,"prevalence","KO",1000)
Genome_size_TEST = envelope_intersection(indata,"prevalence","Genome_size",1000)
GC_TEST = envelope_intersection(indata,"prevalence","GC_content",1000)



#计算通才和专才分类群的平均CDS和KO_richness
g_ko_richness = taxonomy_16s_extended$ko_richness_genome[taxonomy_16s_extended$prevalence_s>=6]
g_CDS = taxonomy_16s_extended$CDS[taxonomy_16s_extended$prevalence_s>=6]
s_ko_richness = taxonomy_16s_extended$ko_richness_genome[taxonomy_16s_extended$prevalence_s==1]
s_CDS = taxonomy_16s_extended$CDS[taxonomy_16s_extended$prevalence_s==1]
g_Genome_size = taxonomy_16s_extended$Genome_size[taxonomy_16s_extended$prevalence_s>=6]
g_GC_content = taxonomy_16s_extended$GC_content[taxonomy_16s_extended$prevalence_s>=6]
s_Genome_size = taxonomy_16s_extended$Genome_size[taxonomy_16s_extended$prevalence_s==1]
s_GC_content = taxonomy_16s_extended$GC_content[taxonomy_16s_extended$prevalence_s==1]


g_ko_richness = g_ko_richness[g_ko_richness!=-1]
g_CDS = g_CDS[g_CDS!=-1]
s_ko_richness = s_ko_richness[s_ko_richness!=-1]
s_CDS = s_CDS[s_CDS!=-1]


plot_scatter_boxplot <- function(array_a, array_b, 
                                 lable_a = "lable_a", 
                                 lable_b = "lable_b", 
                                 lable_y = "lable_y",
                                 color_a = "#69b3a2", 
                                 color_b = "#404080", 
                                 alpha = 0.3) {
  
  # 1. 构造数据框
  array_a = array_a[!is.na(array_a)]
  array_b = array_b[!is.na(array_b)]
  df <- data.frame(
    value = c(array_a, array_b),
    group = factor(c(rep(lable_a, length(array_a)), rep(lable_b, length(array_b))),
                   levels = c(lable_a, lable_b))
  )
  
  # 2. 计算差异显著性 (T-test) 和 效应值 (Cohen's d)
  test_res <- t.test(array_a, array_b)
  p_val <- test_res$p.value
  d_res <- effsize::cohen.d(array_a, array_b, na.rm = TRUE)
  eff_size <- d_res$estimate
  
  # 格式化右上角的文本
  stats_text <- paste0("p-value: ", format.pval(p_val, digits = 3), 
                        "  Cohen's d: ", round(eff_size, 3))
  
  # 3. 计算每组的平均值和最大值位置
  # 用于在每组最上方显示平均值
  summary_info <- data.frame(
    group = c(lable_a, lable_b),
    mean_val = c(mean(array_a), mean(array_b)),
    max_val = c(max(array_a), max(array_b))
  )
  
  # 4. 绘图
  p <- ggplot(df, aes(x = group, y = value, fill = group, color = group)) +
    # 绘制箱线图 (隐藏离群点，因为我们会画散点)
    geom_boxplot(outlier.shape = NA, alpha = 0.4, width = 0.5) +
    # 绘制抖动散点
    geom_jitter(width = 0.2, alpha = alpha) +
    # 设置颜色
    scale_fill_manual(values = c(color_a, color_b)) +
    scale_color_manual(values = c(color_a, color_b)) +
    # 在右上角添加显著性和效应值
    annotate("text", x = Inf, y = Inf, label = stats_text, 
             hjust = 1.1, vjust = 1.5, size = 4, fontface = "italic", color = "black") +
    # 在每组上方显示平均值 (坐标使用该组最大值)
    geom_text(data = summary_info, 
              aes(x = group, y = max_val, label = paste("Mean:", round(mean_val, 2))),
              vjust = -0.5, inherit.aes = FALSE) +
    # 基础修饰
    theme_minimal() +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          panel.grid.minor = element_blank()) +
    labs(y = lable_y) +
    # 适当扩大 y 轴范围防止文字被顶栏遮挡
    scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))+
    theme_bw() + 
    theme(
      panel.background = element_blank(),
      legend.position = "none",
      panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
      plot.title = element_text(size = 20, face = "bold", color = "black", hjust = 0.5, margin = margin(b = 10)),
      plot.subtitle = element_text(size = 12, color = "black", hjust = 0.5, lineheight = 1.2, margin = margin(b = 15)),
      axis.text.x = element_text(size = 14, color = "black"),
      axis.title.x = element_text(size = 14, color = "black", face = "bold"), 
      axis.text.y = element_text(size = 14, color = "black", angle = 0), 
      axis.title.y = element_text(size = 14, color = "black", face = "bold"),
      plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
    )

  return(p)
}


library(ggpubr)
library(patchwork)

p_ko = plot_scatter_boxplot(g_ko_richness,s_ko_richness,"Generalist","Specialist","KO richness")
p_cds = plot_scatter_boxplot(g_CDS,s_CDS,"Generalist","Specialist","CDS")
P_Genome_size = plot_scatter_boxplot(g_Genome_size,s_Genome_size,"Generalist","Specialist","Genome size (mpb)")
P_GC_content = plot_scatter_boxplot(g_GC_content,s_GC_content,"Generalist","Specialist","GC content (%)")
final_plot_GS <- (p_ko | p_cds )/(P_GC_content | P_Genome_size)
ggsave("result/Trait_shifts/KO_CDS_GS.pdf",final_plot_GS,width = 10,height = 10)
final_plot_GS
```

#ITS Samples
```{r}
rm(list = ls())
source("Tools.R")
library(dplyr)
library(ggplot2)
library(forcats)
library(tidyr)
library(reshape2)
library(forcats)
taxonomy_its_extended = read.delim("standard/taxonomy_its_extended.txt")
#The relationship between the taxonomic species composition of ITS samples and their prevalence_s rates
prevalence_s_Class_its = taxonomy_its_extended[ ,c("prevalence_s","Class")]
prevalence_s_Class_its[prevalence_s_Class_its$prevalence_s >= 15,]$prevalence_s = 15
make_mutibar(datain = prevalence_s_Class_its,
             out_dir = "result/Trait_shifts/prevalence_s_Class_change_its.pdf",
             x_name = "Prevalence",
             y_name = "Proportion of Taxid",
             legend_name = "Class",
             limit_num = 5,
             width = 7,
             height = 4,
             other_list = c("Unassigned","Incertae Sedis","Unknown"),
             colorlist =  c("#FCB366", "#FE816F", "#90D7C7","#93BCD0", "#AF3F4F", 
                            "#F3BB34", "#F8AD9A", "#5163B5", "#613767", "#56B394",
                            "#DC6D73", "#B2B896", "#E97DAE","#FB7A64","#EDCE35", 
                            "#E5C28C", "#CA8BB6", "#1A6C92", "#EAB46E", "#8CC3BD", 
                            "#4E56AB","#0FBFBC", "#E9A564", "#ADD565", "#ADADAD", 
                            "#8F9EC5", "#6C77B5", "#138A8E", "#4B9D77", "#8E60A8")
)
Class_test = Calculate_correlation(prevalence_s_Class_its,"prevalence_s","Class")
class_prop = as.data.frame(table(prevalence_s_Class_its$Class))
main_class = class_prop$Var1[class_prop$Freq>100]
Class_test = Class_test[Class_test$Variable%in%main_class,]
print(Class_test)
#The relationship between the proportion of mycorrhizal fungi in ITS samples and their prevalence_s rate
prevalence_s_Mycorrhizal_its = taxonomy_its_extended[ ,c("prevalence_s","mycorrhizal")]
prevalence_s_Mycorrhizal_its <- prevalence_s_Mycorrhizal_its %>%
  replace(is.na(.), "Unknown")
prevalence_s_Mycorrhizal_its[prevalence_s_Mycorrhizal_its$prevalence_s >= 15,]$prevalence_s = 15
prevalence_s_Mycorrhizal_its[prevalence_s_Mycorrhizal_its$prevalence_s >= 15,]$prevalence_s = 15
make_mutibar(datain = prevalence_s_Mycorrhizal_its,
             out_dir = "result/Trait_shifts/prevalence_s_Mycorrhizal_change_its.pdf",
             x_name = "Prevalence",
             y_name = "Proportion of Taxid",
             legend_name = "Mycorrhizal",
             limit_num = 10,
             width = 7,
             height = 4,
             other_list = c("Unassigned","Incertae Sedis","Unknown"),
             colorlist = c("#C1D2E2","#4B5B74","#eeeeee")
)
Mycorrhizal_test = Calculate_correlation(prevalence_s_Mycorrhizal_its,"prevalence_s","mycorrhizal")
print(Mycorrhizal_test)
# 菌根真菌是什么属呢
mycorrhizal_tax = taxonomy_its_extended[taxonomy_its_extended$mycorrhizal == "TRUE",]
p1 = draw_pie_chart(mycorrhizal_tax,"Class", label = F, pre = TRUE,
                   "Mycorrhizal Class",limit_n = 8)
p1
ggsave("result/Trait_shifts/Mycorrhizal_Genus_distribution.pdf",p1,width = 6,height = 4)
# 菌根真菌都在那些植物中分布呢
# 挑选所有菌根真菌
library(data.table)
metadata_its = read.delim("raw/metadata_its.txt")
sample_taxid_01_its = fread("standard/sample_taxid_01_its.txt")
sample_taxid_01_its = as.data.frame(column_to_rownames(sample_taxid_01_its,"taxid"))
mycorrhizal_01_its = sample_taxid_01_its[mycorrhizal_tax$taxid,]
mycorrhizal_samples = colnames(mycorrhizal_01_its)[colSums(mycorrhizal_01_its)>0]
mycorrhizal_metadata = metadata_its[metadata_its$SampleID%in%mycorrhizal_samples,]
p2 = draw_pie_chart(mycorrhizal_metadata,"Genus", label = F, pre = TRUE,
                   "Mycorrhizal Samples",limit_n = 8)
p2
ggsave("result/Trait_shifts/Mycorrhizal_Samples_distribution.pdf",p2,width = 6,height = 4)
```


# Identify which genes have an increased or decreased occurrence rate as the popularity increases or decreases.
```{r,warning=FALSE}
rm(list = ls())
source("Tools.R")
library(dplyr)
library(tidyr)
library(data.table)
taxonomy_16s_extended = read.delim("standard/taxonomy_16s_extended.txt")
ko_data = fread("raw/ko_data_genome.txt")
ko_data = column_to_rownames(as.data.frame(ko_data),"taxid")

ko_data[ko_data != 0] = 1
type_data = taxonomy_16s_extended[,c("taxid","prevalence_s")]
type_data$prevalence_s[type_data$prevalence_s >= 25] = 25

indata = merge(type_data,ko_data,by.x = "taxid",by.y = "row.names")
# 合并数据并计算比例
result_df <- indata %>%
  # 将数据转换为长格式以便按流行率分组
  pivot_longer(cols = -c(taxid, prevalence_s), 
               names_to = "gene", 
               values_to = "presence") %>%
  # 按流行率和基因分组
  group_by(prevalence_s, gene) %>%
  # 计算每个组别的比例
  summarise(
    proportion = sum(presence == 1, na.rm = TRUE) / n(),
    .groups = "drop"
  ) 

result_df <- spread(result_df, key = "gene",
                    value = "proportion")

# 设置显著性水平（通常为0.05，可根据需要调整）
alpha <- 0.01

# 初始化存储显著正相关和负相关列名的字符向量
positive_vars <- character(0)
negative_vars <- character(0)

# 提取自变量（第一列）和因变量（其他列）
result_df$prevalence_s = as.numeric(result_df$prevalence_s)
independent_var <- result_df$prevalence_s
dependent_df <- result_df[, -1, drop = FALSE]  # 因变量数据框（保持数据框结构）

# 循环遍历每个因变量列
for (col_name in names(dependent_df)) {
  dependent_var <- dependent_df[[col_name]]  # 提取当前因变量列
  
  # 执行相关性检验（默认使用Pearson，可修改method参数）
  # 使用tryCatch避免因错误（如缺失值过多）中断循环
  test_result <- tryCatch({
    cor.test(independent_var, dependent_var, method = "pearson")  # 可改为"spearman"或"kendall"
  }, error = function(e) {
    return(NULL)  # 如果出错（如数据问题），返回NULL
  })
  
  # 跳过出错的列
  if (is.null(test_result)) next
  
  # 提取p值和相关系数
  p_value <- test_result$p.value
  cor_coef <- test_result$estimate
  
  # 检查是否有效（避免NA值）
  if (!is.na(p_value) && !is.na(cor_coef)) {
    # 判断显著性及相关方向
    if (p_value < alpha) {
      if (cor_coef > 0) {
        positive_vars <- c(positive_vars, col_name)  # 显著正相关
      } else if (cor_coef < 0) {
        negative_vars <- c(negative_vars, col_name)  # 显著负相关
      }
    }
  }
}

#导入kegg分类信息
ko_level_table = read.delim("db/ko_level_table.txt")
p_ko_level = ko_level_table[ko_level_table$KO%in%positive_vars,]
write.table(p_ko_level,file = "result/Trait_shifts/p_ko_level.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
n_ko_level = ko_level_table[ko_level_table$KO%in%negative_vars,]
write.table(n_ko_level,file = "result/Trait_shifts/n_ko_level.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
#对比level3下前20正相关基因和负相关基因的差异
freq_table <- table(p_ko_level$level3)
sorted_freq <- sort(freq_table, decreasing = TRUE)
p_level3_10 <- names(sorted_freq)[1:10]
freq_table <- table(n_ko_level$level3)
sorted_freq <- sort(freq_table, decreasing = TRUE)
n_level3_10 <- names(sorted_freq)[1:10]
# 提取p_level3_10独有的元素
p_level3_10_unique <- setdiff(p_level3_10, n_level3_10)
p_level3_10_unique

# 提取n_level3_10独有的元素
n_level3_10_unique <- setdiff(n_level3_10, p_level3_10)
n_level3_10_unique

# 共有的功能
common_func = intersect(n_level3_10, p_level3_10)
common_func
```
# Annotate and visualize the genes with significant differences
```{r}
rm(list = ls())
source("Tools.R")
library(dplyr)
library(tidyr)
library(ggplot2)
p_ko_level = read.delim("result/Trait_shifts/p_ko_level.txt")
length(unique(p_ko_level$KO))
n_ko_level = read.delim("result/Trait_shifts/n_ko_level.txt")
length(unique(n_ko_level$KO))
p_ko_pathway = as.data.frame(table(p_ko_level$level3))
colnames(p_ko_pathway)[1] = "Pathway"
p_ko_pathway = p_ko_pathway[order(p_ko_pathway$Freq, decreasing = TRUE), ][c(1:15),]
n_ko_pathway = as.data.frame(table(n_ko_level$level3))
colnames(n_ko_pathway)[1] = "Pathway"
n_ko_pathway = n_ko_pathway[order(n_ko_pathway$Freq, decreasing = TRUE), ][c(1:15),]

colnames(p_ko_pathway)[2] = "p_freq"
colnames(n_ko_pathway)[2] = "n_freq"

pn_ko_pathway = merge(p_ko_pathway,n_ko_pathway,by = "Pathway",all = T)
pn_ko_pathway[is.na(pn_ko_pathway)] = 0
mutal_pathway = pn_ko_pathway[pn_ko_pathway$Pathway%in%intersect(p_ko_pathway$Pathway,n_ko_pathway$Pathway),]
mutal_pathway$diff = mutal_pathway$p_freq-mutal_pathway$n_freq
mutal_pathway = mutal_pathway[order(-mutal_pathway$diff), ][ ,-4]
p_pathway = pn_ko_pathway[pn_ko_pathway$Pathway%in%setdiff(p_ko_pathway$Pathway,n_ko_pathway$Pathway),]
p_pathway <- p_pathway[order(-p_pathway$p_freq), ]
n_pathway = pn_ko_pathway[pn_ko_pathway$Pathway%in%setdiff(n_ko_pathway$Pathway,p_ko_pathway$Pathway),]
n_pathway <- n_pathway[order(n_pathway$n_freq), ]
pn_ko_pathway_order = rbind(p_pathway,mutal_pathway,n_pathway)

df_long <- pivot_longer(pn_ko_pathway_order, 
                        cols = c(p_freq, n_freq), 
                        names_to = "variable", 
                        values_to = "value")

df_long$Pathway <- factor(df_long$Pathway, levels = pn_ko_pathway_order$Pathway)
df_long$variable <- factor(df_long$variable, levels =c("p_freq","n_freq"))

# Plot
ggplot(df_long, aes(x = Pathway, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(x = "Pathway", y = "Occurrence Rate", title = "", fill = "") +
  scale_fill_manual(values = c("p_freq" = "#DE7117", "n_freq" = "#355070")) +
  theme_classic() +
  theme(
    legend.position = "top",
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    axis.text.x  = element_text(size = 14, color = "black", angle = -45, hjust = 0, vjust = 1),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y  = element_text(size = 14, color = "black"),
    axis.title.y = element_text(size = 14, color = "black", face = "bold"),
    legend.title = element_text(size = 14, color = "black"),
    legend.text  = element_text(size = 14, color = "black"),
    plot.margin  = margin(t = 10, r = 130, b = 10, l = 10, unit = "pt")
  )
ggsave("result/Trait_shifts/pn_ko_pathway.pdf",width = 15,height = 6)



pgpg_levels = read.delim("db/pgpg_levels.txt")
p_ko_pgpg = pgpg_levels[pgpg_levels$KO%in%p_ko_level$KO,]
n_ko_pgpg = pgpg_levels[pgpg_levels$KO%in%n_ko_level$KO,]

freq_table_Lv3_p = as.data.frame(table(p_ko_pgpg$Lv3))
freq_table_Lv3_n = as.data.frame(table(n_ko_pgpg$Lv3))

colnames(freq_table_Lv3_p)[1] = "PGPG_Function"
freq_table_Lv3_p = freq_table_Lv3_p[order(freq_table_Lv3_p$Freq, decreasing = TRUE), ][c(1:15),]
colnames(freq_table_Lv3_n)[1] = "PGPG_Function"
freq_table_Lv3_n = freq_table_Lv3_n[order(freq_table_Lv3_n$Freq, decreasing = TRUE), ][c(1:15),]


colnames(freq_table_Lv3_p)[2] = "p_freq"
colnames(freq_table_Lv3_n)[2] = "n_freq"

pn_ko_PGPG_Function = merge(freq_table_Lv3_p,freq_table_Lv3_n,by = "PGPG_Function",all = T)
pn_ko_PGPG_Function[is.na(pn_ko_PGPG_Function)] = 0
mutal_PGPG_Function = pn_ko_PGPG_Function[pn_ko_PGPG_Function$PGPG_Function%in%intersect(freq_table_Lv3_p$PGPG_Function,freq_table_Lv3_n$PGPG_Function),]
mutal_PGPG_Function$diff = mutal_PGPG_Function$p_freq-mutal_PGPG_Function$n_freq
mutal_PGPG_Function = mutal_PGPG_Function[order(-mutal_PGPG_Function$diff), ][ ,-4]
p_PGPG_Function = pn_ko_PGPG_Function[pn_ko_PGPG_Function$PGPG_Function%in%setdiff(freq_table_Lv3_p$PGPG_Function,freq_table_Lv3_n$PGPG_Function),]
p_PGPG_Function <- p_PGPG_Function[order(-p_PGPG_Function$p_freq), ]
n_PGPG_Function = pn_ko_PGPG_Function[pn_ko_PGPG_Function$PGPG_Function%in%setdiff(freq_table_Lv3_n$PGPG_Function,freq_table_Lv3_p$PGPG_Function),]
n_PGPG_Function <- n_PGPG_Function[order(n_PGPG_Function$n_freq), ]
pn_ko_PGPG_Function_order = rbind(p_PGPG_Function,mutal_PGPG_Function,n_PGPG_Function)

df_long <- pivot_longer(pn_ko_PGPG_Function_order, 
                        cols = c(p_freq, n_freq), 
                        names_to = "variable", 
                        values_to = "value")

df_long$PGPG_Function <- factor(df_long$PGPG_Function, levels = pn_ko_PGPG_Function_order$PGPG_Function)
df_long$variable <- factor(df_long$variable, levels =c("p_freq","n_freq"))

# Plot
ggplot(df_long, aes(x = PGPG_Function, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(x = "PGPG_Function", y = "Occurrence Rate", title = "", fill = "") +
  scale_fill_manual(values = c("p_freq" = "#DE7117", "n_freq" = "#355070")) +
  theme_classic() +
  theme(
    legend.position = "top",
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    axis.text.x  = element_text(size = 14, color = "black", angle = -45, hjust = 0, vjust = 1),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y  = element_text(size = 14, color = "black"),
    axis.title.y = element_text(size = 14, color = "black", face = "bold"),
    legend.title = element_text(size = 14, color = "black"),
    legend.text  = element_text(size = 14, color = "black"),
    plot.margin  = margin(t = 10, r = 150, b = 10, l = 10, unit = "pt")
  )
ggsave("result/Trait_shifts/pn_ko_PGPG_Function.pdf",width = 15,height = 8)

writeLines(as.character(pn_ko_pathway_order$Pathway),"result/Trait_shifts/pn_pathway.txt")
writeLines(as.character(pn_ko_PGPG_Function_order$PGPG_Function),"result/Trait_shifts/pn_PGPG.txt")
```


# Observe the storage situation of nitrogen-fixing bacteria in different plants
```{r,fig.show='animate'}
rm(list = ls())
source("Tools.R")
library(tidyr)
library(tibble)
library(tidyverse)
library(ComplexHeatmap)
library(ape)
dir.create("result/nitrogen_fixation")
nf = read.delim("db/nitrogen_genus.txt")
annotation_g_16s = read.delim("result/GOS/tree_g_16s/annotation_g_16s.txt")
metadata_16s <- read.delim("raw/metadata_16s.txt")
taxonomy_16s_extended <- read.delim("standard/taxonomy_16s_extended.txt")
PS_taxid_01_16s = read.delim("standard/PS_taxid_01_16s.txt")
# PG_taxid_01_16s = read.delim("standard/PG_taxid_01_16s.txt")
PG_MG_sum_16s = merge_datafame(PS_taxid_01_16s,row_group = taxonomy_16s_extended[ ,c("taxid","Genus")],
                               row_method = "sum",col_group = unique(metadata_16s[ ,c("Species","Genus")]),col_method = "sum")
rownames(PG_MG_sum_16s)[which(rownames(PG_MG_sum_16s) == "Burkholderia-Caballeronia-Paraburkholderia")] <- "Burkholderia"
#Identify which generalists are nitrogen-fixing bacteria?
generalist_nf_genus = intersect(unique(annotation_g_16s$Genus),nf$Genus)
print("The overlapping number of the Nitrogen-fixing Bacteria genus and the Generalist Microorganism genus is:")
length(generalist_nf_genus)


# Convert it into a matrix, which is the required data format for the pheatmap function.
PG_NF_sum_16s = PG_MG_sum_16s[nf$Genus, ]
PG_NF_sum_16s[PG_NF_sum_16s != 0] <- 1
PG_NF_sum_16s = PG_NF_sum_16s[rowSums(PG_NF_sum_16s)>0,]
# Count the number of plant genera that possess all three types of nitrogen-fixing bacteria simultaneously
three_type = merge_datafame(PG_NF_sum_16s,row_group = nf[nf$Genus%in%rownames(PG_NF_sum_16s),][ ,c(2,1)],row_method = "sum")
three_type[three_type!=0] = 1
three_type_nf_genus = colnames(three_type)[colSums(three_type)==3]
print("Count the number of plant genera that possess all three types of nitrogen-fixing bacteria simultaneously")
length(three_type_nf_genus)
# Average nitrogen-fixing bacteria holdings
mean(colSums(PG_NF_sum_16s))
# The average holdings of three types of rhizobia in the Gramineae and Leguminosae families
genus_list = c("Poales","Fabales")
type_list = unique(nf$classify)
for (pg in genus_list){
  pglist = unique(metadata_16s$Genus[metadata_16s$Order==pg])
  for (nt in type_list){
    ntgenus = nf$Genus[nf$classify==nt]
    ntgenus = intersect(ntgenus,rownames(PG_NF_sum_16s))
    temp_matrix = PG_NF_sum_16s[ntgenus,pglist]
    mean_val = mean(colSums(temp_matrix))
    print(paste0(pg," - ",nt," = ",mean_val))
  }
}


# Draw a heat map to visualize the distribution of nitrogen-fixing bacteria
colnames(PG_NF_sum_16s) = paste(colnames(PG_NF_sum_16s),colSums(PG_NF_sum_16s),sep = "_")
rownames(PG_NF_sum_16s) = paste(rownames(PG_NF_sum_16s),rowSums(PG_NF_sum_16s),sep = "_")
# Initialize the color matrix (set all values to white, corresponding to 0)
Host_order = colnames(PG_NF_sum_16s)[order(colSums(PG_NF_sum_16s), decreasing = TRUE)]
PG_NF_sum_16s = PG_NF_sum_16s[ ,Host_order]
mat <- as.matrix(t(PG_NF_sum_16s))
color_mat <- matrix("#FFFFFF", nrow = nrow(mat), ncol = ncol(mat))
# Set the color of the middle values in different column intervals to 1.
color_mat[, 1:13][mat[, 1:13] == 1]   <- "#FE816F"
color_mat[, 14:20][mat[, 14:20] == 1] <- "#90D7C7"
color_mat[, 21:34][mat[, 21:34] == 1] <- "#FCB366"
Genus_order_raw = data.frame(Genus = colnames(mat),raw_index = c(1:ncol(mat)))
# Version sorted by quantity
# Sorting of host nitrogen-fixing bacteria holdings
NF_order1 = c(rownames(PG_NF_sum_16s[1:13,])[order(rowSums(PG_NF_sum_16s[1:13,]), decreasing = TRUE)],
              rownames(PG_NF_sum_16s[14:20,])[order(rowSums(PG_NF_sum_16s[14:20,]), decreasing = TRUE)],
              rownames(PG_NF_sum_16s[21:34,])[order(rowSums(PG_NF_sum_16s[21:34,]), decreasing = TRUE)])
Genus_order_1 = data.frame(Genus = NF_order1,order_index1 = c(1:length(NF_order1)))
NF_order2 = rownames(PG_NF_sum_16s)[order(rowSums(PG_NF_sum_16s), decreasing = TRUE)]
Genus_order_2 = data.frame(Genus = NF_order2,order_index2 = c(1:length(NF_order2)))
Genus_order_merge = merge(Genus_order_raw,Genus_order_1,by = "Genus")
Genus_order_merge = merge(Genus_order_merge,Genus_order_2,by = "Genus")
color_mat1 = color_mat[ ,Genus_order_merge$raw_index[order(Genus_order_merge$order_index1)]]
color_mat2 = color_mat[ ,Genus_order_merge$raw_index[order(Genus_order_merge$order_index2)]]
mat1 = mat[ ,NF_order1]
mat2 = mat[ ,NF_order2]

ha <- HeatmapAnnotation(
  Types = c(rep("Symbiotic", 13),
               rep("Free-living", 7),
               rep("Associative", 14)),
  col = list(
    Types = c(
      "Symbiotic" = "#FE816F",
      "Free-living" = "#90D7C7",
      "Associative" = "#FCB366"
    )
  ),
  show_legend = FALSE
)
# 绘制热图

pdf("result/nitrogen_fixation/nitrogen_fixation_heatmap1.pdf", width = 13, height = 18) 
Heatmap(mat1,
        rect_gp = grid::gpar(fill = color_mat1, col = "grey80", lwd = 0.5),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        top_annotation = ha,
        show_row_names = TRUE,
        show_heatmap_legend = FALSE)
dev.off()
pdf("result/nitrogen_fixation/nitrogen_fixation_heatmap2.pdf", width = 10, height = 16) 
Heatmap(mat2,
        rect_gp = grid::gpar(fill = color_mat2, col = "grey80", lwd = 0.5),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        # top_annotation = ha,
        show_row_names = TRUE,
        show_heatmap_legend = FALSE)
dev.off()
Heatmap(mat1,
        rect_gp = grid::gpar(fill = color_mat1, col = "grey80", lwd = 0.5),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        top_annotation = ha,
        show_row_names = TRUE,
        show_heatmap_legend = FALSE)



# Take a look at the holdings of nitrogen-fixing genes in generalist microorganisms
library(data.table)
library(tibble)
nf_gene = read.delim("db/nf_gene.txt")
ko_data = fread("raw/ko_data_genome.txt")
ko_data = column_to_rownames(as.data.frame(ko_data),"taxid")
annotation_g_16s = read.delim("result/GOS/tree_g_16s/annotation_g_16s.txt",row.names = 1)
annotation_g_16s$Genus[annotation_g_16s$Genus=="Burkholderia-Caballeronia-Paraburkholderia"] = "Burkholderia"
ko_data_g_nf = ko_data[rownames(annotation_g_16s),nf_gene$KO]
ko_data_g_nf = na.omit(ko_data_g_nf)
ko_data_g_nf[ko_data_g_nf!=0]=1
ko_data_g_nf_DHK = ko_data_g_nf[rowSums(ko_data_g_nf[ ,c(1:6)])>0,]
colnames(ko_data_g_nf_DHK) = nf_gene$gene
ko_data_g_nf_DHK = ko_data_g_nf_DHK[order(rowSums(ko_data_g_nf_DHK),decreasing = TRUE),]
rownames(ko_data_g_nf_DHK) = make_unique_sequential(annotation_g_16s[rownames(ko_data_g_nf_DHK),]$Genus)



mat <- as.matrix(ko_data_g_nf_DHK)
color_mat <- matrix("#FFFFFF", nrow = nrow(mat), ncol = ncol(mat))
# 设置不同列区间中值为1的颜色
color_mat[, 1:6][mat[, 1:6] == 1]   <- "#E5A984"
color_mat[, 7:21][mat[, 7:21] == 1] <- "#F7E09B"
color_mat[, 22:27][mat[, 22:27] == 1] <- "#808DAA"
ha <- HeatmapAnnotation(
  Types = c(rep("Core nif genes", 6),
               rep("Auxiliary nif genes", 15),
               rep("Nod-related genes", 6)),
  col = list(
    Types = c(
      "Core nif genes" = "#E5A984",
      "Auxiliary nif genes" = "#F7E09B",
      "Nod-related genes" = "#808DAA"
    )
  ),
  show_legend = FALSE
)
pdf("result/nitrogen_fixation/generalist_nitrogen_gene.pdf", width = 10, height = 8) 
Heatmap(mat,
        rect_gp = grid::gpar(fill = color_mat, col = "grey80", lwd = 0.5),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        top_annotation = ha,
        show_row_names = TRUE,
        show_heatmap_legend = FALSE)
dev.off()
Heatmap(mat,
        rect_gp = grid::gpar(fill = color_mat, col = "grey80", lwd = 0.5),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        top_annotation = ha,
        show_row_names = TRUE,
        show_heatmap_legend = FALSE)

ko_data_g_nf$sum = rowSums(ko_data_g_nf[ ,c(1:27)])
ko_data_g_nf$nod = rowSums(ko_data_g_nf[ ,c(22:27)])
ko_data_g_nf$nif = rowSums(ko_data_g_nf[ ,c(1:21)])
ko_data_g_nf$HDK = rowSums(ko_data_g_nf[ ,c(1:6)])

nf_g_tax = merge(annotation_g_16s[ ,c(1:10)],ko_data_g_nf[ ,c("sum","nod","nif","HDK")],by = "row.names")
#含有固氮酶的通才的数量
sum(nf_g_tax$nod>0)
sum(nf_g_tax$nif>0)
sum(nf_g_tax$HDK>0)
sum(nf_g_tax$nod>0&nf_g_tax$HDK>0)
```


# Draw a heatmap of microbial functions
```{r,fig.show='animate'}
rm(list = ls())
library(pheatmap)
library(ComplexHeatmap)
micro_function    <- read.delim("raw/micro_function.txt")
micro_function_merge <- micro_function %>%
  group_by(Genus) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  column_to_rownames(var = "Genus")
taxonomy_16s_extended <- read.delim("standard/taxonomy_16s_extended.txt")
taxonomy_its_extended <- read.delim("standard/taxonomy_its_extended.txt")

high_prevalence_16s <- unique(taxonomy_16s_extended$Genus[taxonomy_16s_extended$GOS_s=="generalist"])
high_prevalence_16s[high_prevalence_16s=="Burkholderia-Caballeronia-Paraburkholderia"] = "Burkholderia"
high_prevalence_its <- unique(taxonomy_its_extended$Genus[taxonomy_its_extended$GOS_s=="generalist"])

micro_function_merge_16s <- micro_function_merge[high_prevalence_16s, ]
micro_function_merge_16s <- na.omit(micro_function_merge_16s)
micro_function_merge_16s <- micro_function_merge_16s[rowSums(micro_function_merge_16s != 0) > 0, ]
#按照功能丰富度进行排序
micro_function_01_16s = micro_function_merge_16s
micro_function_01_16s[micro_function_01_16s!=0] = 1
micro_function_merge_16s = micro_function_merge_16s[order(rowSums(micro_function_01_16s),decreasing = TRUE), ]

micro_function_merge_its <- micro_function_merge[high_prevalence_its, ]
micro_function_merge_its <- na.omit(micro_function_merge_its)
micro_function_merge_its <- micro_function_merge_its[rowSums(micro_function_merge_its != 0) > 0, ]
#按照功能丰富度进行排序
micro_function_01_its = micro_function_merge_its
micro_function_01_its[micro_function_01_its!=0] = 1
micro_function_merge_its = micro_function_merge_its[order(rowSums(micro_function_01_its),decreasing = TRUE), ]


mat_16s = as.matrix(micro_function_merge_16s)
color_mat_16s <- matrix("#FFFFFF", nrow = nrow(mat_16s), ncol = ncol(mat_16s))

# 设置不同列区间中值为1的颜色
color_mat_16s[, 1:12][mat_16s[, 1:12] == 1]   <- "#E5F5E0"
color_mat_16s[, 1:12][mat_16s[, 1:12] == 2]   <- "#C7E9C0"
color_mat_16s[, 1:12][mat_16s[, 1:12] == 3]   <- "#A1D99B"
color_mat_16s[, 1:12][mat_16s[, 1:12] == 4]   <- "#74C476"
color_mat_16s[, 1:12][mat_16s[, 1:12] >= 5]   <- "#41AB5D"
color_mat_16s[, 13][mat_16s[, 13] == 1]   <- "#F5E5E0"
color_mat_16s[, 13][mat_16s[, 13] == 2]   <- "#E9C7C0"
color_mat_16s[, 13][mat_16s[, 13] == 3]   <- "#D9A19B"
color_mat_16s[, 13][mat_16s[, 13] == 4]   <- "#C47476"
color_mat_16s[, 13][mat_16s[, 13] >= 5]   <- "#AB415D"
dir.create("result/article_function")
pdf("result/article_function/article_function_16s.pdf", width = 10, height = 10.1) 
Heatmap(mat_16s,
        rect_gp = grid::gpar(fill = color_mat_16s, col = "grey80", lwd = 0.5),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        show_row_names = TRUE,
        show_heatmap_legend = FALSE,
        cell_fun = function(j, i, x, y, width, height, fill) {
    if (mat_16s[i, j] > 0) {
      grid.text(label = mat_16s[i, j], x = x, y = y, gp = gpar(fontsize = 8))
    }
  })
dev.off()
Heatmap(mat_16s,
        rect_gp = grid::gpar(fill = color_mat_16s, col = "grey80", lwd = 0.5),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        show_row_names = TRUE,
        show_heatmap_legend = FALSE,
        cell_fun = function(j, i, x, y, width, height, fill) {
    if (mat_16s[i, j] > 0) {
      grid.text(label = mat_16s[i, j], x = x, y = y, gp = gpar(fontsize = 8))
    }
  })

mat_its = as.matrix(micro_function_merge_its)
color_mat_its <- matrix("#FFFFFF", nrow = nrow(mat_its), ncol = ncol(mat_its))

# 设置不同列区间中值为1的颜色
color_mat_its[, 1:12][mat_its[, 1:12] == 1]   <- "#E5F5E0"
color_mat_its[, 1:12][mat_its[, 1:12] == 2]   <- "#C7E9C0"
color_mat_its[, 1:12][mat_its[, 1:12] == 3]   <- "#A1D99B"
color_mat_its[, 1:12][mat_its[, 1:12] == 4]   <- "#74C476"
color_mat_its[, 1:12][mat_its[, 1:12] >= 5]   <- "#41AB5D"
color_mat_its[, 13][mat_its[, 13] == 1]   <- "#F5E5E0"
color_mat_its[, 13][mat_its[, 13] == 2]   <- "#E9C7C0"
color_mat_its[, 13][mat_its[, 13] == 3]   <- "#D9A19B"
color_mat_its[, 13][mat_its[, 13] == 4]   <- "#C47476"
color_mat_its[, 13][mat_its[, 13] >= 5]   <- "#AB415D"

pdf("result/article_function/article_function_its.pdf", width = 8, height = 4.5) 
Heatmap(mat_its,
        rect_gp = grid::gpar(fill = color_mat_its, col = "grey80", lwd = 0.5),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        show_row_names = TRUE,
        show_heatmap_legend = FALSE,
        cell_fun = function(j, i, x, y, width, height, fill) {
    if (mat_its[i, j] > 0) {
      grid.text(label = mat_its[i, j], x = x, y = y, gp = gpar(fontsize = 8))
    }
  })
dev.off()
Heatmap(mat_its,
        rect_gp = grid::gpar(fill = color_mat_its, col = "grey80", lwd = 0.5),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        show_row_names = TRUE,
        show_heatmap_legend = FALSE,
        cell_fun = function(j, i, x, y, width, height, fill) {
    if (mat_its[i, j] > 0) {
      grid.text(label = mat_its[i, j], x = x, y = y, gp = gpar(fontsize = 8))
    }
  })
```


# The relationship between YAS strategy score and prevalence rate
```{r}

rm(list = ls())
source("Tools.R")
cds_data = read.delim("raw/cds_data.txt")
taxonomy_16s_extended <- read.delim("standard/taxonomy_16s_extended.txt")
trait_matrixatgranularity1 = read.csv("raw/micortrait_otuput/microtrait_granularity1.csv",row.names = 1)
#Calculate the YAS score
Y_score <- rowSums(trait_matrixatgranularity1[, grepl("Resource\\.Use", names(trait_matrixatgranularity1))])
A_score <- rowSums(trait_matrixatgranularity1[, grepl("Resource\\.Acquisition", names(trait_matrixatgranularity1))])
S_score <- rowSums(trait_matrixatgranularity1[, grepl("Stress\\.Tolerance", names(trait_matrixatgranularity1))])
YAS_score = data.frame(row.names = rownames(trait_matrixatgranularity1),
                       Y = Y_score,
                       A = A_score,
                       S = S_score)
YAS_score$sum = rowSums(YAS_score)
YAS_score = merge(cds_data,YAS_score,by.x = "genome_name",by.y = "row.names",all.x = TRUE)
YAS_score = na.omit(YAS_score)
YAS_score$Y_ra = YAS_score$Y/YAS_score$CDS
YAS_score$A_ra = YAS_score$A/YAS_score$CDS
YAS_score$S_ra = YAS_score$S/YAS_score$CDS

cor.test(YAS_score$Y_ra, YAS_score$CDS, method = "pearson")
cor.test(YAS_score$A_ra, YAS_score$CDS, method = "pearson")
cor.test(YAS_score$S_ra, YAS_score$CDS, method = "pearson")

cor.test(YAS_score$Y_ra, YAS_score$CDS, method = "spearman")
cor.test(YAS_score$A_ra, YAS_score$CDS, method = "spearman")
cor.test(YAS_score$S_ra, YAS_score$CDS, method = "spearman")

YAS_score_prevalence = merge(YAS_score,taxonomy_16s_extended[ ,c("taxid","prevalence_s")],by="taxid",all.x = TRUE)
YAS_score_prevalence$prevalence_s[YAS_score_prevalence$prevalence_s>25]=25

Y_result = tukey_boxplot(df = YAS_score_prevalence, group_col = "prevalence_s", value_col = "Y_ra", 
              fill_color = "#F8AD9A",point_color = "#FE816F", point_alpha = 0.2, HSD = F, 
              xlab = "Prevalence", ylab = "Yield score", tlab = "Prevalence-Yield score",
              tilted = F,force = F, MAXLAB = FALSE,rmargin=30,show_n = F,order = F)
A_result = tukey_boxplot(df = YAS_score_prevalence, group_col = "prevalence_s", value_col = "A_ra", 
              fill_color = "#EBD7A5",point_color = "#E4A644", point_alpha = 0.2, HSD = F, 
              xlab = "Prevalence", ylab = "Acquisition score", tlab = "Prevalence-Acquisition score",
              tilted = F,force = F, MAXLAB = FALSE,rmargin=30,show_n = F,order = F)
S_result = tukey_boxplot(df = YAS_score_prevalence, group_col = "prevalence_s", value_col = "S_ra", 
              fill_color = "#93BCD0",point_color = "#1A6C92", point_alpha = 0.2, HSD = F, 
              xlab = "Prevalence", ylab = "Stresstolerance score", tlab = "Prevalence-Stresstolerance score",
              tilted = F,force = F, MAXLAB = FALSE,rmargin=30,show_n = F,order = F)
Y_result[[2]]
A_result[[2]]
S_result[[2]]

cor_test <- cul_cor_df(YAS_score_prevalence,"prevalence_s")

dir.create("result/YAS")
A_plot = A_result[[2]]+
  annotate("text", x = Inf, y = Inf, 
            label = paste0("spearman p: ",ifsig(cor_test["A_ra","spearman_p"])),
            hjust = 1.1, vjust = 1.5, size = 5) 
ggsave("result/YAS/A_plot.pdf",A_plot,width = 8,height = 4)
Y_plot = Y_result[[2]]+
  annotate("text", x = Inf, y = Inf, 
            label = paste0("spearman p: ",ifsig(cor_test["Y_ra","spearman_p"])),
            hjust = 1.1, vjust = 1.5, size = 5) 
ggsave("result/YAS/Y_plot.pdf",Y_plot,width = 8,height = 4)
S_plot = S_result[[2]]+
  annotate("text", x = Inf, y = Inf, 
            label = paste0("spearman p: ",ifsig(cor_test["S_ra","spearman_p"])),
            hjust = 1.1, vjust = 1.5, size = 5) 
ggsave("result/YAS/S_plot.pdf",S_plot,width = 8,height = 4)
```

# The relationship between the occurrence frequency and prevalence of YAS-related micro-traits
```{r}
rm(list = ls())
source("Tools.R")
library(dplyr)
library(tidyr)
library(data.table)
cds_data = read.delim("raw/cds_data.txt")
trait_matrixatgranularity3 = read.csv("raw/micortrait_otuput/microtrait_granularity3.csv",row.names = 1)
taxonomy_16s_extended = read.delim("standard/taxonomy_16s_extended.txt")
taxid_trait_3 = merge(cds_data[,c(1,2)],trait_matrixatgranularity3,by.x = "genome_name",by.y = "row.names",all.x = TRUE)[,-1]
taxid_trait_3 = tibble::column_to_rownames(taxid_trait_3,"taxid")


taxid_trait_3[taxid_trait_3 != 0] = 1
type_data = taxonomy_16s_extended[,c("taxid","prevalence_s")]
type_data$prevalence_s[type_data$prevalence_s >= 25] = 25
indata = merge(type_data,taxid_trait_3,by.x = "taxid",by.y = "row.names")

result_df <- indata %>%
  pivot_longer(cols = -c(taxid, prevalence_s), 
               names_to = "gene", 
               values_to = "presence") %>%
  group_by(prevalence_s, gene) %>%
  summarise(
    proportion = sum(presence == 1, na.rm = TRUE) / n(),
    .groups = "drop"
  ) 

result_df <- spread(result_df, key = "gene",
                    value = "proportion")

# Set the significance level
alpha <- 0.01

# Initialize a character vector to store the names of the significantly positively correlated and negatively correlated columns
positive_vars <- character(0)
negative_vars <- character(0)

# Extract the independent variable (in the first column) and the dependent variable (in the other columns)
result_df$prevalence_s = as.numeric(result_df$prevalence_s)
independent_var <- result_df$prevalence_s
dependent_df <- result_df[, -1, drop = FALSE] 


for (col_name in names(dependent_df)) {
  dependent_var <- dependent_df[[col_name]]
  test_result <- tryCatch({
    cor.test(independent_var, dependent_var, method = "pearson")
  }, error = function(e) {
    return(NULL) 
  })
  if (is.null(test_result)) next
  p_value <- test_result$p.value
  cor_coef <- test_result$estimate
  if (!is.na(p_value) && !is.na(cor_coef)) {
    if (p_value < alpha) {
      if (cor_coef > 0) {
        positive_vars <- c(positive_vars, col_name)
      } else if (cor_coef < 0) {
        negative_vars <- c(negative_vars, col_name)
      }
    }
  }
}
sig_microtrait = data.frame(micro_trait = c(positive_vars,negative_vars),cor = c(rep("positive",length(positive_vars)),rep("negative",length(negative_vars))))
sig_microtrait$type = "unknown"
sig_microtrait$type[grepl("Resource\\.Use", sig_microtrait$micro_trait)] = "Yield"
sig_microtrait$type[grepl("Resource\\.Acquisition", sig_microtrait$micro_trait)] = "Resource Acquisition"
sig_microtrait$type[grepl("Stress\\.Tolerance", sig_microtrait$micro_trait)] = "Stress Tolerance"
write.table(sig_microtrait,file = "result/Trait_shifts/Table_s3_sig_microtrait.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
# The number of times each trait corresponding to the YAS strategy appears in the positively correlated micro-traits
sum(grepl("Resource\\.Use", positive_vars))
sum(grepl("Resource\\.Acquisition", positive_vars))
sum(grepl("Stress\\.Tolerance", positive_vars))
# The number of times each trait corresponding to the YAS strategy appears in the negatively correlated micro-traits
sum(grepl("Resource\\.Use", positive_vars))
sum(grepl("Resource\\.Use", negative_vars))
sum(grepl("Resource\\.Acquisition", negative_vars))
sum(grepl("Stress\\.Tolerance", negative_vars))
```

# Study the relationship between the proportion of PGPT in the genome and the changes in its prevalence
```{r}
library(tibble)
library(tidyverse)
source("Tools.R")
cds_data = read.delim("raw/cds_data.txt")
pgpg_levels = read.delim("db/pgpg_levels.txt")
taxonomy_16s_extended = read.delim("standard/taxonomy_16s_extended.txt")
gene_counts_lv4 = t(read.delim("raw/pgpg_finder_output/gene_counts_Lv4.txt",row.names = 1))
gene_counts_lv4_counts = column_to_rownames(merge(cds_data,gene_counts_lv4,by.x = "genome_name",by.y = "row.names",all.x = TRUE),"taxid")[ ,-1]
gene_counts_lv4_normalized = apply(gene_counts_lv4_counts[ , -1], 2, function(x) x / gene_counts_lv4_counts[ , 1])
gene_counts_lv4_normalized = as.data.frame(t(gene_counts_lv4_normalized))

gene_counts_lv3 = t(read.delim("raw/pgpg_finder_output/gene_counts_lv3.txt",row.names = 1))
gene_counts_lv3_counts = column_to_rownames(merge(cds_data,gene_counts_lv3,by.x = "genome_name",by.y = "row.names",all.x = TRUE),"taxid")[ ,-1]
gene_counts_lv3_normalized = apply(gene_counts_lv3_counts[ , -1], 2, function(x) x / gene_counts_lv3_counts[ , 1])
gene_counts_lv3_normalized = as.data.frame(t(gene_counts_lv3_normalized))

#聚类为lv2等级表
row_group = unique(pgpg_levels[,c("Lv3","Lv2")])
row_group = row_group[row_group$Lv3 %in% rownames(gene_counts_lv3_normalized),]
gene_counts_lv2_normalized = merge_datafame(gene_counts_lv3_normalized,row_group = row_group,row_method = "sum")

prevalence_pgpg_lv3 = merge(taxonomy_16s_extended[ ,c("taxid","prevalence_s")],t(gene_counts_lv3_normalized),by.x = "taxid",by.y = "row.names")
prevalence_pgpg_lv4 = merge(taxonomy_16s_extended[ ,c("taxid","prevalence_s")],t(gene_counts_lv4_normalized),by.x = "taxid",by.y = "row.names")
prevalence_pgpg_lv2 = merge(taxonomy_16s_extended[ ,c("taxid","prevalence_s")],t(gene_counts_lv2_normalized),by.x = "taxid",by.y = "row.names")
cor_test_lv2 = cul_cor_df(prevalence_pgpg_lv2,"prevalence_s")
cor_test_lv3 = cul_cor_df(prevalence_pgpg_lv3,"prevalence_s")
cor_test_lv4 = cul_cor_df(prevalence_pgpg_lv4,"prevalence_s")

pgpg_levels_lv3_lv2 = unique(pgpg_levels[ ,c("Lv3","Lv2")])
sig_lv3_df = cor_test_lv3[cor_test_lv3$spearman_p<0.05,]
sig_lv3_df = rownames_to_column(sig_lv3_df,"Lv3")
sig_lv3_df$sig <- sapply(sig_lv3_df$spearman_p, ifsig)
sig_lv3_df = merge(sig_lv3_df,pgpg_levels_lv3_lv2,by = "Lv3",all.x = TRUE)
write.table(sig_lv3_df,file = "result/Trait_shifts/Table_s4_pgpg_sig_lv3.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")

```
