---
title: "analysis_pipeline"
author: "Zilong Wang"
date: "2026-03-16"
output:
  pdf_document: default
  html_document: default
---

# Data preprocessing

```{r}
rm(list = ls())
library(dplyr)
library(ggplot2)
library(tibble)
library(vegan)
library(ggplot2)
library(dplyr)
library(tidyr)
dir.create("result")
# Remove the taxid rows that contain "Eukaryota" (fungi), "Chloroplast" (chloroplasts), and "Mitochondria" (mitochondria).
sample_taxid_abundance_16s = read.delim("raw/sample_taxid_abundance_16s.txt")
taxonomy_16s <- read.delim("raw/taxonomy_all16s.txt")
pattern <- "Eukaryota|Chloroplast|Mitochondria"

taxonomy_cols <- taxonomy_16s[, -1, drop = FALSE]
matches <- sapply(taxonomy_cols, function(col) {
  grepl(pattern, col, perl = TRUE) 
})

matches[is.na(matches)] <- FALSE

# Determine whether there is at least one match in each row (any column contains keywords)）
has_match <- apply(matches, 1, any) 

# Extract the otuid of the matching line and convert it into a character vector
non_bac <- as.character(taxonomy_16s$taxid[has_match])
filtered_taxonomy <- taxonomy_16s[! (taxonomy_16s$taxid %in% non_bac), ]
write.table(filtered_taxonomy,file = "raw/taxonomy_all16s.txt",sep = "\t", 
            row.names = FALSE,col.names = TRUE,quote = FALSE,fileEncoding = "UTF-8")
filtered_table <- sample_taxid_abundance_16s[! (sample_taxid_abundance_16s$taxid %in% non_bac), ]
write.table(filtered_table,file = "raw/sample_taxid_abundance_16s.txt",sep = "\t", 
            row.names = FALSE,col.names = TRUE,quote = FALSE,fileEncoding = "UTF-8")

# Remove samples with too few annotations
sample_taxid_abundance_16s = read.delim("raw/sample_taxid_abundance_16s.txt",row.names = 1)
metadata_16s <- read.csv("raw/metadata_all16s.csv")
merged_16s_classified <- sample_taxid_abundance_16s[rownames(sample_taxid_abundance_16s) != "Unassigned", ]
merged_16s_temp <- merged_16s_classified %>%
  select(where(~ is.numeric(.) && sum(., na.rm = TRUE) >= 1000))
pure_list = colnames(merged_16s_temp)
merged_16s_filter = sample_taxid_abundance_16s[ ,c(pure_list)]
merged_16s_filter = rownames_to_column(merged_16s_filter,var="taxid")
metadata_16s_filter = metadata_16s[metadata_16s$SampleID %in% c(pure_list), ]
write.table(merged_16s_filter,file = "raw/sample_taxid_abundance_16s.txt",sep = "\t", 
            row.names = FALSE,col.names = TRUE,quote = FALSE,fileEncoding = "UTF-8")
write.csv(metadata_16s_filter, "raw/metadata_all16s.csv", fileEncoding = "UTF-8", row.names = FALSE)
```

#Convert the matrix into a 01 matrix
```{r}
rm(list = ls())
library(Hmisc)
library(psych)
library(ggClusterNet)
library(metagenomeSeq)
library(tibble)
sample_taxid_abundance_16s = read.delim("raw/sample_taxid_abundance_16s.txt", row.names = 1)
sample_taxid_abundance_its = read.delim("raw/sample_taxid_abundance_its.txt", row.names = 1)
sample_taxid_01_16s <- as.data.frame(lapply(sample_taxid_abundance_16s, function(x) ifelse(x > 0, 1, 0)))
rownames(sample_taxid_01_16s) <- rownames(sample_taxid_abundance_16s)
sample_taxid_01_16s = rownames_to_column(sample_taxid_01_16s,var = "taxid")
write.table(sample_taxid_01_16s,file = "sample_taxid_01_16s.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
sample_taxid_01_its <- as.data.frame(lapply(sample_taxid_abundance_its, function(x) ifelse(x > 0, 1, 0)))
rownames(sample_taxid_01_its) <- rownames(sample_taxid_abundance_its)
sample_taxid_01_its = rownames_to_column(sample_taxid_01_its,var = "taxid")
write.table(sample_taxid_01_its,file = "sample_taxid_01_its.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
```

#The proportion of annotable reads in the visualized sequence to the total number of reads
```{r,fig.height = 5,fig.with = 12}
rm(list = ls())
dir.create("result")
library(ggplot2)
library(patchwork)
sample_taxid_abundance_16s = read.delim("raw/sample_taxid_abundance_16s.txt", row.names = 1)
sample_taxid_abundance_its = read.delim("raw/sample_taxid_abundance_its.txt", row.names = 1)
# Calculate the annotation ratio ((1 - Unassigned / Total sum of each column))
unassigned_values <- as.numeric(sample_taxid_abundance_16s[rownames(sample_taxid_abundance_16s) == "Unassigned", ])
col_totals <- colSums(sample_taxid_abundance_16s)
annotated_prop <- 1 - (unassigned_values / col_totals)
plot_data <- data.frame(annotated_proportion = annotated_prop)

# Calculate the basic statistics (mean and standard deviation)
mean_val <- mean(plot_data$annotated_proportion, na.rm = TRUE)
sd_val <- sd(plot_data$annotated_proportion, na.rm = TRUE)


hist_plot1 <- ggplot(plot_data, aes(x = annotated_proportion)) +
  geom_histogram(fill = "#DD9273", color = "black", bins = 30, alpha = 0.8) + 
  xlab("The annotated proportion") +
  ylab("Counts")+
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.1,
           label = paste0("Mean: ", round(mean_val, 3), "\nSD: ", round(sd_val, 3)),
           size = 5) +
  theme_classic() + 
  theme(
    legend.position = "top" ,
    panel.background = element_blank(), 
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2), 
    plot.title = element_text(
      size = 20,
      face = "bold",
      color = "black",
      hjust = 0.5,  
      margin = margin(b = 10)  
    ),
    plot.subtitle = element_text(
      size = 12,
      color = "black",
      hjust = 0.5,  
      lineheight = 1.2,  
      margin = margin(b = 15) 
    ),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y = element_text(size = 14, color = "black", angle = 0), 
    axis.title.y = element_text(size = 14, color = "black", face = "bold"), 
    legend.title = element_text(size = 14, color = "black"), 
    legend.text = element_text(size = 14, color = "black"), 
    plot.margin = margin(t = 10,  
                         r = 30,  
                         b = 10,  
                         l = 10,  
                         unit = "pt") 
  )

#its
unassigned_values <- as.numeric(sample_taxid_abundance_its[rownames(sample_taxid_abundance_its) == "Unassigned", ])
col_totals <- colSums(sample_taxid_abundance_its)
annotated_prop <- 1 - (unassigned_values / col_totals)
plot_data <- data.frame(annotated_proportion = annotated_prop)
mean_val <- mean(plot_data$annotated_proportion, na.rm = TRUE)
sd_val <- sd(plot_data$annotated_proportion, na.rm = TRUE)
hist_plot2 <- ggplot(plot_data, aes(x = annotated_proportion)) +
  geom_histogram(fill = "#DD9273", color = "black", bins = 30, alpha = 0.8) +
  xlab("The annotated proportion") +
  ylab("Counts")+
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.1,
           label = paste0("Mean: ", round(mean_val, 3), "\nSD: ", round(sd_val, 3)),
           size = 5) +
  theme_classic() + 
  theme(
    legend.position = "top" ,
    panel.background = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2), 
    plot.title = element_text(
      size = 20,
      face = "bold",
      color = "black",
      hjust = 0.5,
      margin = margin(b = 10) 
    ),
    plot.subtitle = element_text(
      size = 12,
      color = "black",
      hjust = 0.5,  
      lineheight = 1.2, 
      margin = margin(b = 15) 
    ),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y = element_text(size = 14, color = "black", angle = 0),
    axis.title.y = element_text(size = 14, color = "black", face = "bold"), 
    legend.title = element_text(size = 14, color = "black"),
    legend.text = element_text(size = 14, color = "black"), 
    plot.margin = margin(t = 10,  
                         r = 30,  
                         b = 10,  
                         l = 10, 
                         unit = "pt")
  )
library(ggpubr)
combined_plot <- ggarrange(hist_plot1, hist_plot2,  
                           ncol = 2 
)
combined_plot

ggsave("result/Annotated_Proportions.pdf",combined_plot,width = 12,height = 5)
```

#Visualization extraction methods and the proportion of sequencing regions
```{r,fig.show='animate'}
rm(list = ls())
source("Tools.R")
library(ggplot2)
library(dplyr)
library(scales)
metadata_16s <- read.csv("raw/metadata_all16s.csv")
metadata_its <- read.csv("raw/metadata_allits.csv")
# 16s extract_way
p1 = draw_pie_chart(metadata_16s,"extract_way", label = TRUE, pre = TRUE,"Extract Way")
p1
ggsave("result/extact_way_16s.pdf",p1,width = 6,height = 6)
# 16s Sequencing Region
p2 = draw_pie_chart(metadata_16s,"region", label = F, pre = TRUE,"Sequencing Region")
p2
ggsave("result/Sequencing_Region_16s.pdf",p2,width = 6,height = 4)
# its extract_way
p3 = draw_pie_chart(metadata_its,"extract_way", label = TRUE, pre = TRUE,"Extract Way")
p3
ggsave("result/extact_way_its.pdf",p3,width = 6,height = 4)
# its Sequencing Region
p4 = draw_pie_chart(metadata_its,"region", label = TRUE, pre = TRUE,"Region")
p4
ggsave("result/Sequencing_Region_its.pdf",p4,width = 6,height = 4)
```

# Draw a pcoa diagram based on jaccard distance
```{r,fig.show='animate'}
rm(list = ls())

# Load required packages
library(vegan)
library(ggplot2)
library(dplyr)
library(ggpubr)
library(tibble)
# Import data
sample_taxid_01_16s = read.delim("sample_taxid_01_16s.txt", row.names=1)
sample_taxid_01_its = read.delim("sample_taxid_01_its.txt", row.names=1)
metadata_16s <- read.csv("raw/metadata_all16s.csv")
metadata_its <- read.csv("raw/metadata_allits.csv")

# 16S PCoA and PERMANOVA
jaccard_dist <- vegdist(t(sample_taxid_01_16s), method = "jaccard")
spe1 = decostand(t(sample_taxid_01_16s), "hel")
design2 <- column_to_rownames(metadata_16s, "SampleID")
adonis = adonis2(spe1 ~ Genus, data = design2, permutations = 999, method = "jaccard", by = "margin")

pcoa__16s <- cmdscale(jaccard_dist, k = 2, eig = TRUE)
pcoa_scores <- as.data.frame(pcoa__16s$points)
colnames(pcoa_scores) <- c("PCoA1", "PCoA2")
pcoa_scores$SampleID <- rownames(pcoa_scores)
pcoa_data_16s <- merge(pcoa_scores, metadata_16s, by = "SampleID", all.x = TRUE)

# Calculate variance explained by the first two axes
eigenvalues <- pcoa__16s$eig
variance_explained <- eigenvalues / sum(eigenvalues) * 100
pc1_var_16s <- round(variance_explained[1], 2)
pc2_var_16s <- round(variance_explained[2], 2)

# ITS PCoA and PERMANOVA
otutab_filtered_its <- sample_taxid_01_its[rowSums(sample_taxid_01_its) > 0, ]
jaccard_dist <- vegdist(t(sample_taxid_01_its), method = "jaccard")
spe2 = decostand(t(sample_taxid_01_its), "hel")
design3 <- column_to_rownames(metadata_its, "SampleID")
adonis3 = adonis2(spe2 ~ Genus, data = design3, permutations = 999, method = "jaccard", by = "margin")

pcoa__its <- cmdscale(jaccard_dist, k = 2, eig = TRUE)
pcoa_scores <- as.data.frame(pcoa__its$points)
colnames(pcoa_scores) <- c("PCoA1", "PCoA2")
pcoa_scores$SampleID <- rownames(pcoa_scores)
pcoa_data_its <- merge(pcoa_scores, metadata_its, by = "SampleID", all.x = TRUE)

eigenvalues <- pcoa__its$eig
variance_explained <- eigenvalues / sum(eigenvalues) * 100
pc1_var_its <- round(variance_explained[1], 2)
pc2_var_its <- round(variance_explained[2], 2)

# Create a unified color palette for plant genera
colorlist <- c("#C29D59", "#9B443B", "#7B6D4A", "#4B5B74", "#C6803A", "#CB4047",
               "#CD5176", "#A08A6D", "#816148", "#CDB97E", "#A971A0", "#3E4069", "#CDA59D",
               "#B24057", "#BC9D88", "#886349", "#B07E9F", "#726F8E", "#A17D5D",
               "#875215", "#C1C19D", "#E4A644", "#E0CF5B", "#826859", "#E9BDA0", "#E5B8A1",
               "#E7D6C2", "#BDB79F", "#B99FC6", "#D69DA4", "#C1D2E2", "#6D5249",
               "#FDE0E4", "#465069", "#B37C75", "#B19E8F", "#6C7612", "#624733", "#D56847",
               "#BC9F77", "#A1997B", "#CB3F48", "#A94B32", "#A54B43", "#8A926B",
               "#427370", "#A18F81", "#4E95C5", "#D5ABBF", "#E0C8B0", "#8D6A4A", "#67361C",
               "#F7E09B", "#9C7C52", "#CB9D7B", "#B89052", "#3C2913", "#A7785B",
               "#DDD2BE", "#E8EEBC", "#F6DC54", "#B2620C", "#CA9C4F", "#509C82", "#517D7E",
               "#8F6C4B", "#CB7E3A", "#808DAA", "#C5A376", "#DE7117", "#9EB395")

plant_genus = unique(c(c(pcoa_data_16s$Genus), c(pcoa_data_its$Genus)))
plant_genus_color = character()
i = 1
for (p in plant_genus) {
  plant_genus_color[p] = colorlist[i]
  i = i + 1
}

# Plot 16S PCoA
p1 <- ggplot(pcoa_data_16s, aes(x = PCoA1, y = PCoA2, color = Genus)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(x = paste0("PCoA1 (", pc1_var_16s, "%)"),
       y = paste0("PCoA2 (", pc2_var_16s, "%)"),
       color = "Plant Genus",
       title = "16S Samples",
       subtitle = paste("R2=", round((adonis$R2[1]), 3), "p=", adonis$`Pr(>F)`[1])) +
  scale_color_manual(values = plant_genus_color) +
  theme_classic() +
  theme(
    legend.position = "none",
    panel.background = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    plot.title = element_text(
      size = 20,
      face = "bold",
      color = "black",
      hjust = 0.5,
      margin = margin(b = 10)
    ),
    plot.subtitle = element_text(
      size = 12,
      color = "black",
      hjust = 0.5,
      lineheight = 1.2,
      margin = margin(b = 15)
    ),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y = element_text(size = 14, color = "black", angle = 0),
    axis.title.y = element_text(size = 14, color = "black", face = "bold"),
    legend.title = element_text(size = 14, color = "black", face = "bold"),
    legend.text = element_text(size = 14, color = "black"),
    plot.margin = margin(t = 10,
                         r = 30,
                         b = 10,
                         l = 10,
                         unit = "pt")
  )

# Plot ITS PCoA
p2 <- ggplot(pcoa_data_its, aes(x = PCoA1, y = PCoA2, color = Genus)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(x = paste0("PCoA1 (", pc1_var_its, "%)"),
       y = paste0("PCoA2 (", pc2_var_its, "%)"),
       color = "Plant Genus",
       title = "ITS Samples",
       subtitle = paste("R2=", round((adonis3$R2[1]), 3), "p=", adonis3$`Pr(>F)`[1])) +
  scale_color_manual(values = plant_genus_color) +
  theme_classic() +
  theme(
    legend.position = "none",
    panel.background = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    plot.title = element_text(
      size = 20,
      face = "bold",
      color = "black",
      hjust = 0.5,
      margin = margin(b = 10)
    ),
    plot.subtitle = element_text(
      size = 12,
      color = "black",
      hjust = 0.5,
      lineheight = 1.2,
      margin = margin(b = 15)
    ),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y = element_text(size = 14, color = "black", angle = 0),
    axis.title.y = element_text(size = 14, color = "black", face = "bold"),
    legend.title = element_text(size = 14, color = "black", face = "bold"),
    legend.text = element_text(size = 14, color = "black"),
    plot.margin = margin(t = 10,
                         r = 30,
                         b = 10,
                         l = 10,
                         unit = "pt")
  )

# Combine plots with a shared legend
combined_plot <- ggarrange(p1, p2,
                           ncol = 2,
                           common.legend = TRUE,
                           legend = "right")

combined_plot
ggsave("result/16s_its_PCoA__jaccard.pdf", plot = combined_plot, width = 15, height = 5.5)

```

# Use PERMANOVA and VPA analyses to test the contribution of host factors to the community results
```{r,fig.show='animate'}
rm(list = ls())
library(ggplot2)
library(dplyr)
library(tidyr)
sample_taxid_01_16s = read.delim("sample_taxid_01_16s.txt", row.names=1)
sample_taxid_01_its = read.delim("sample_taxid_01_its.txt", row.names=1)
metadata_16s <- read.csv("raw/metadata_all16s.csv")
metadata_its <- read.csv("raw/metadata_allits.csv")
spe1=decostand(t(sample_taxid_01_16s),"hel")
design2 <-column_to_rownames(metadata_16s,"SampleID")
spe2=decostand(t(sample_taxid_01_its),"hel")
design3 <-column_to_rownames(metadata_its,"SampleID")
# vpa
jaccard_dist_16s <- vegdist(t(sample_taxid_01_16s), method = "jaccard")
jaccard_dist_its <- vegdist(t(sample_taxid_01_its), method = "jaccard")

vpa_result_16s <- varpart(jaccard_dist_16s, 
                      ~ Genus, 
                      ~ extract_way, 
                      ~ Country, 
                      data = design2)
pdf("result/vpa_plot_16s.pdf", width = 7, height = 6) 
plot(vpa_result_16s, 
     digits = 2, 
     Xnames = c("Genus", "Extract_way", "Country"), 
     bg = c("tomato", "skyblue", "lightgreen"))
dev.off()


vpa_result_its <- varpart(jaccard_dist_its, 
                          ~ Genus, 
                          ~ extract_way, 
                          ~ Country, 
                          data = design3)
pdf("result/vpa_plot_its.pdf", width = 7, height = 6)
plot(vpa_result_its, 
     digits = 2, 
     Xnames = c("Genus", "Extract_way", "Country"), 
     bg = c("tomato", "skyblue", "lightgreen"))
dev.off()

# PERMANOVA

prepare_data <- function(df, label) {
  df %>%
    as.data.frame() %>%
    mutate(Factor = rownames(.)) %>%
    filter(Factor %in% c("Genus", "Country", "extract_way")) %>%
    mutate(Community = label) %>%
    select(Community, Factor, R2)
}


adonis_test_16s = adonis2(spe1 ~ Genus + extract_way + Country,data =design2,permutations = 999,method="jaccard", by="margin",parallel = 8)
adonis_test_its = adonis2(spe2 ~ Genus + extract_way + Country,data =design3,permutations = 999,method="jaccard", by="margin",parallel = 8)
plot_df_16s <- prepare_data(adonis_test_16s, "16S (Bacteria)")
plot_df_its <- prepare_data(adonis_test_its, "ITS (Fungi)")


final_plot_df <- rbind(plot_df_16s, plot_df_its)
final_plot_df$R2_percent <- final_plot_df$R2 * 100


PERMANOVA = ggplot(final_plot_df, aes(x = Factor, y = R2_percent, fill = Community)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = sprintf("%.2f%%", R2_percent)), 
            position = position_dodge(width = 0.8), 
            vjust = -0.5, size = 3.5, fontface = "bold") +
  scale_fill_manual(values = c("16S (Bacteria)" = "#4E79A7", "ITS (Fungi)" = "#F28E2B")) +
  labs(title = "Comparison of Community Variation Explained (R²)",
       x = "Explanatory Factors",
       y = "Variance Explained (R² %)",
       fill = "Community Type") +
  scale_y_continuous(limits = c(0, max(final_plot_df$R2_percent) + 2)) +
  theme_bw() +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.subtitle = element_text(hjust = 0.5, color = "grey30"),
    axis.text = element_text(size = 11, color = "black"),
    legend.position = "top"
  )
ggsave("result/PERMANOVA.pdf",plot = PERMANOVA, width = 8, height = 7)
```



# Draw a scatter plot showing the relationship between the number of plant genera and the number of microbial species
```{r,fig.show='animate'}
rm(list = ls())
source("Tools.R")
library(tidyr)
library(tidyverse)
sample_taxid_01_16s = read.delim("sample_taxid_01_16s.txt", row.names=1)
sample_taxid_01_its = read.delim("sample_taxid_01_its.txt", row.names=1)
metadata_16s <- read.csv("raw/metadata_all16s.csv")
metadata_its <- read.csv("raw/metadata_allits.csv")

column_sums <- colSums(sample_taxid_01_16s)
result_df <- data.frame(
  SampleID = names(column_sums),
  sum = column_sums
)
rownames(result_df) <- NULL
genus_tax_count_16s = merge(metadata_16s, result_df, by = "SampleID")
genus_tax_count_16s$log10 = log10(genus_tax_count_16s$sum + 1)

# Calculate mean richness for each plant genus
mean_richness_16s = genus_tax_count_16s %>%
  group_by(Genus) %>%
  summarise(mean_richness = mean(sum, na.rm = TRUE)) %>%
  arrange(desc(mean_richness))

column_sums <- colSums(sample_taxid_01_its)
result_df <- data.frame(
  SampleID = names(column_sums),
  sum = column_sums
)
rownames(result_df) <- NULL
genus_tax_count_its = merge(metadata_its, result_df, by = "SampleID")
genus_tax_count_its$log10 = log10(genus_tax_count_its$sum + 1)

# Calculate mean richness for each plant genus
mean_richness_its = genus_tax_count_its %>%
  group_by(Genus) %>%
  summarise(mean_richness = mean(sum, na.rm = TRUE)) %>%
  arrange(desc(mean_richness))

# Generate boxplots for 16S and ITS richness
genus_count_16s = tukey_boxplot(genus_tax_count_16s, "Genus", "log10", "#427370", "#4B5B74", 0.2, FALSE, "Plant Genus", "Richness (log10)", "The number of microbial species in the seeds of each genus of plants(16S Samples)")
genus_count_16s[[2]]
ggsave("result/genus_tax_count_16s.pdf", plot = genus_count_16s[[2]], width = 16, height = 6)

genus_count_its = tukey_boxplot(genus_tax_count_its, "Genus", "log10", "#427370", "#4B5B74", 0.2, FALSE, "Plant Genus", "Richness (log10)", "The number of microbial species in the seeds of each genus of plants(ITS Samples)")
genus_count_its[[2]]
ggsave("result/genus_tax_count_its.pdf", plot = genus_count_its[[2]], width = 16, height = 6)
```

#Merge the functional prediction results with the taxonomy table

```{r}
rm(list = ls())
taxonomy_16s <- read.delim("raw/taxonomy_all16s.txt")
taxonomy_its <- read.delim("raw/taxonomy_allits.txt")
picrust2_result = read.delim("raw/picrust2_ko_richness.txt")
bugbase_result= read.delim("raw/bugbase_result.txt")
funguild_result = read.delim("raw/funguild_result.txt")
taxonomy_16s_final = merge(taxonomy_16s,picrust2_result,by = "taxid", all.x = TRUE)
taxonomy_16s_final = merge(taxonomy_16s_final,bugbase_result,by = "taxid", all.x = TRUE)
taxonomy_16s_final = taxonomy_16s_final %>%
  mutate(Aerobic_type = ifelse(Aerobic=="TRUE","Aerobic",ifelse(Anaerobic=="TRUE","Anaerobic",ifelse(Facultatively_Anaerobic=="TRUE","Facultatively_Anaerobic","Unknown")))) %>%
  mutate(Gram_type = ifelse(Gram_Positive=="TRUE","Positive",ifelse(Gram_Negative=="TRUE","Negative","Unknown")))%>%
  replace(is.na(.), "Unknown")

write.table(taxonomy_16s_final,file = "taxonomy_16s_final.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
taxonomy_its_final = merge(taxonomy_its,funguild_result,by = "taxid", all.x = TRUE)%>%replace(is.na(.), "Unknown")
taxonomy_its_final = taxonomy_its_final%>%replace(is.na(.), "Unknown")
write.table(taxonomy_its_final,file = "taxonomy_its_final.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
```

# Identify General Microbiome
## Create the "PG–TAXID" table
```{r}
rm(list = ls())
source("Tools.R")
library(dplyr)
library(tibble)
library(vegan)
# import data
sample_taxid_01_16s = read.delim("sample_taxid_01_16s.txt", row.names=1)
sample_taxid_01_its = read.delim("sample_taxid_01_its.txt", row.names=1)
metadata_16s <- read.csv("raw/metadata_all16s.csv")
metadata_its <- read.csv("raw/metadata_allits.csv")
taxonomy_16s <- read.delim("raw/taxonomy_all16s.txt")
taxonomy_its <- read.delim("raw/taxonomy_allits.txt")

Genus_list = unique(metadata_16s$Genus)
PG_taxid_16s = taxonomy_16s[ ,c("taxid"), drop = FALSE]
for (i in Genus_list){
  message(paste0("processing",i))
  tempdf=sample_taxid_01_16s[ ,metadata_16s[metadata_16s$Genus==i, ]$SampleID,drop = FALSE]
  if (ncol(tempdf)>=10){
    result = classify_generalist(tempdf,i,1000)
  }else if(ncol(tempdf)==1){
    result <- rownames_to_column(tempdf,"taxid")
    colnames(result)[2] <- i
  }else{
    result <- data.frame(
      taxid = rownames(tempdf),
      Presence = as.integer(apply(tempdf, 1, function(row) any(row != 0)))
    )
    colnames(result)[2] <- i
  }
  PG_taxid_16s = merge(PG_taxid_16s,result,by= "taxid")
}
write.table(PG_taxid_16s,file = "PG_taxid_16s.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")

Genus_list = unique(metadata_its$Genus)
PG_taxid_its = taxonomy_its[ ,c("taxid"), drop = FALSE]
for (i in Genus_list){
  tempdf=sample_taxid_01_its[ ,metadata_its[metadata_its$Genus==i, ]$SampleID,drop = FALSE]
  if (ncol(tempdf)>=10){
    result = classify_generalist(tempdf,i,1000)
  }else if(ncol(tempdf)==1){
    result <- rownames_to_column(tempdf,"taxid")
    colnames(result)[2] <- i
  }else{
    result <- data.frame(
      taxid = rownames(tempdf),
      Presence = as.integer(apply(tempdf, 1, function(row) any(row != 0)))
    )
    colnames(result)[2] <- i
  }
  PG_taxid_its = merge(PG_taxid_its,result,by= "taxid")
}
write.table(PG_taxid_its,file = "PG_taxid_its.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
```

## Calculate the host corresponding to the specialist
```{r}
rm(list = ls())
library(tidyr)
library(tibble)
PG_taxid_16s = read.delim("PG_taxid_16s.txt")
PG_taxid_its = read.delim("PG_taxid_its.txt")
taxonomy_16s_final = read.delim("taxonomy_16s_final.txt")
taxonomy_its_final = read.delim("taxonomy_its_final.txt")

taxonomy_16s_final = merge(
  taxonomy_16s_final,
  data.frame(taxid = PG_taxid_16s$taxid, prevalence = rowSums(PG_taxid_16s[, -1])),
  by = "taxid"
)
taxonomy_its_final = merge(
  taxonomy_its_final,
  data.frame(taxid = PG_taxid_its$taxid, prevalence = rowSums(PG_taxid_its[, -1])),
  by = "taxid"
)

# Identify taxa that occur in only one sample and record the corresponding sample name
occurrence_count <- rowSums(PG_taxid_16s[, -1])
specialist_indices <- which(occurrence_count == 1)
specialist_samples <- rep(NA, nrow(PG_taxid_16s))

for (i in specialist_indices) {
  row_data <- PG_taxid_16s[i, -1]
  sample_index <- which(row_data == 1)
  specialist_samples[i] <- names(PG_taxid_16s[, -1])[sample_index]
}

specialist_plant <- data.frame(
  taxid = PG_taxid_16s[, 1],
  specialist_plant = specialist_samples,
  stringsAsFactors = FALSE
)

taxonomy_16s_final = merge(taxonomy_16s_final, specialist_plant, by.x = "taxid", by.y = "taxid")

# Identify taxa that occur in only one sample and record the corresponding sample name
occurrence_count <- rowSums(PG_taxid_its[, -1])
specialist_indices <- which(occurrence_count == 1)
specialist_samples <- rep(NA, nrow(PG_taxid_its))

for (i in specialist_indices) {
  row_data <- PG_taxid_its[i, -1]
  sample_index <- which(row_data == 1)
  specialist_samples[i] <- names(PG_taxid_its[, -1])[sample_index]
}

specialist_plant <- data.frame(
  taxid = PG_taxid_its[, 1],
  specialist_plant = specialist_samples,
  stringsAsFactors = FALSE
)

taxonomy_its_final = merge(taxonomy_its_final, specialist_plant, by.x = "taxid", by.y = "taxid")

write.table(taxonomy_16s_final, file = "taxonomy_16s_final.txt", quote = FALSE, sep = "\t", row.names = FALSE, fileEncoding = "UTF-8")
write.table(taxonomy_its_final, file = "taxonomy_its_final.txt", quote = FALSE, sep = "\t", row.names = FALSE, fileEncoding = "UTF-8")
```

##Create the PO_MG table
```{r}
rm(list = ls())
library(tibble)
library(tidyr)
PG_taxid_16s = read.delim("PG_taxid_16s.txt")
PG_taxid_its = read.delim("PG_taxid_its.txt")
taxonomy_16s_final = read.delim("taxonomy_16s_final.txt")
taxonomy_its_final = read.delim("taxonomy_its_final.txt")
metadata_16s <- read.csv("raw/metadata_all16s.csv")
metadata_its <- read.csv("raw/metadata_allits.csv")
#16s
PG_MG_01_16s = merge(taxonomy_16s_final[ ,c("taxid","Genus")],PG_taxid_16s,by = "taxid")[ ,-1] %>%
  group_by(Genus) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  column_to_rownames(var = "Genus")
PG_MG_01_16s[PG_MG_01_16s != 0] <- 1
PG_MG_01_16s = as.data.frame(t(PG_MG_01_16s))

metadata_16s_classify = unique(metadata_16s[ ,c("Genus","Phylum","Class","Order","Family","pathway")])
freq_table <- table(metadata_16s$Order) 
sorted_freq <- as.data.frame(sort(freq_table, decreasing = TRUE) )
desort_colname <- as.character(sorted_freq[sorted_freq$Freq >= 10, ]$Var1) 

PO_MG_01_16s  = merge(metadata_16s_classify[ ,c("Genus","Order")],PG_MG_01_16s,by.x = "Genus",by.y = "row.names")[ ,-1] %>%
  group_by(Order) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  column_to_rownames(var = "Order")
PO_MG_01_16s [PO_MG_01_16s  != 0] <- 1
PO_MG_01_16s  = PO_MG_01_16s [desort_colname, ]
write.table(rownames_to_column(PO_MG_01_16s,"Order"),file = "PO_MG_01_16s.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")

#its
PG_MG_01_its = merge(taxonomy_its_final[ ,c("taxid","Genus")],PG_taxid_its,by = "taxid")[ ,-1] %>%
  group_by(Genus) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  column_to_rownames(var = "Genus")
PG_MG_01_its[PG_MG_01_its != 0] <- 1
PG_MG_01_its = as.data.frame(t(PG_MG_01_its))

metadata_its_classify = unique(metadata_its[ ,c("Genus","Phylum","Class","Order","Family","pathway")])
freq_table <- table(metadata_its$Order) 
sorted_freq <- as.data.frame(sort(freq_table, decreasing = TRUE) )
desort_colname <- as.character(sorted_freq[sorted_freq$Freq >= 10, ]$Var1) 

PO_MG_01_its  = merge(metadata_its_classify[ ,c("Genus","Order")],PG_MG_01_its,by.x = "Genus",by.y = "row.names")[ ,-1] %>%
  group_by(Order) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  column_to_rownames(var = "Order")
PO_MG_01_its [PO_MG_01_its  != 0] <- 1
PO_MG_01_its  = PO_MG_01_its [desort_colname, ]
write.table(rownames_to_column(PO_MG_01_its,"Order"),file = "PO_MG_01_its.txt",quote = FALSE,sep = "\t",row.names = FALSE,fileEncoding = "UTF-8")
```

##Draw the petal diagram and search for General Microbiome
```{r,fig.show='animate'}
rm(list = ls())
library(tibble)
library(tidyr)
source("Tools.R")
PO_MG_01_16s = read.delim("PO_MG_01_16s.txt",row.names = 1)
PO_MG_01_its = read.delim("PO_MG_01_its.txt",row.names = 1)
# Based on data, calculate the unique and coexisting quantities
n_env <- nrow(PO_MG_01_16s)
env_names <- rownames(PO_MG_01_16s)
microbe_names <- colnames(PO_MG_01_16s)
# Calculate the total number of coexisting microorganisms (those present in all environments, meaning all columns are filled with 1s)
core_microbes_16s <- microbe_names[colSums(PO_MG_01_16s) == n_env]
core_microbes_16s <- core_microbes_16s[core_microbes_16s != "Unassigned"]
core_microbes_16s <- core_microbes_16s[core_microbes_16s != "Incertae Sedis"]
writeLines(core_microbes_16s, "core_microbes_16s.txt")
core_microbes_16s <- readLines("core_microbes_16s.txt")
core <- length(core_microbes_16s)
# Calculate the number of unique microorganisms for each environment (those that exist only in that environment, and are all zero elsewhere)
num <- numeric(n_env)
for (i in 1:n_env) {
  other_sums <- colSums(PO_MG_01_16s[-i, , drop = FALSE])
  unique_to_i <- sum(PO_MG_01_16s[i, ] == 1 & other_sums == 0)
  num[i] <- unique_to_i
}

sample <- env_names
n_colors <- length(sample)
ellipse_col <- c("#A23F25","#3E4069","#517D7E","#CA820A","#C1D2E2","#F7E09B","#808DAA","#E5A984","#552379","#B24057","#A49BC6")

flower_plot(
  sample = sample,
  num = num,
  core = core,
  ellipse_col = ellipse_col,
  alpha = 0.5,  
  start = 90,
  a = 0.6, 
  b = 2,   
  r = 1,  
  nameSize = 1.2,
  numSize = 1.2,
  coreSize = 1.2
)



n_env <- nrow(PO_MG_01_its )
env_names <- rownames(PO_MG_01_its )
microbe_names <- colnames(PO_MG_01_its )
core_microbes_its <- microbe_names[colSums(PO_MG_01_its ) == n_env]
core_microbes_its <- core_microbes_its[core_microbes_its != "Unassigned"]
core_microbes_its <- core_microbes_its[core_microbes_its != "Incertae Sedis"]
writeLines(core_microbes_its, "core_microbes_its.txt")
core_microbes_its <- readLines("core_microbes_its.txt")
core <- length(core_microbes_its)
num <- numeric(n_env)
for (i in 1:n_env) {
  other_sums <- colSums(PO_MG_01_its [-i, , drop = FALSE])
  unique_to_i <- sum(PO_MG_01_its [i, ] == 1 & other_sums == 0)
  num[i] <- unique_to_i
}
sample <- env_names
n_colors <- length(sample)
ellipse_col <- c("#A23F25","#3E4069","#517D7E","#CA820A","#C1D2E2","#F7E09B","#808DAA","#E5A984")

flower_plot(
  sample = sample,
  num = num,
  core = core,
  ellipse_col = ellipse_col,
  alpha = 0.5,
  start = 90,
  a = 0.6, 
  b = 2,
  r = 1, 
  nameSize = 0.9,
  numSize = 0.8,
  coreSize = 1.2
)

```

## Draw a bar chart showing the integrated prevalence of the core microbial genera
```{r,fig.show='animate'}
rm(list = ls())
library(tibble)
library(tidyr)
PG_taxid_16s = read.delim("PG_taxid_16s.txt")
PG_taxid_its = read.delim("PG_taxid_its.txt")
taxonomy_16s_final = read.delim("taxonomy_16s_final.txt")
taxonomy_its_final = read.delim("taxonomy_its_final.txt")
core_microbes_16s <- readLines("core_microbes_16s.txt")
core_microbes_its <- readLines("core_microbes_its.txt")


PG_MG_sum_16s = merge(taxonomy_16s_final[ ,c("taxid","Genus")],PG_taxid_16s,by = "taxid")[ ,-1] %>%
  group_by(Genus) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  column_to_rownames(var = "Genus")
PG_MG_sum_16s[PG_MG_sum_16s != 0] <- 1
PG_MG_sum_16s$sum = rowSums(PG_MG_sum_16s)
write.table(rownames_to_column(PG_MG_sum_16s,"PG"),file = "PG_MG_sum_16s.txt",sep = "\t", 
            row.names = FALSE,col.names = TRUE,quote = FALSE,fileEncoding = "UTF-8")
PG_MG_sum_16s <- PG_MG_sum_16s[, c("sum", setdiff(names(PG_MG_sum_16s), "sum"))]
df_sorted_16s <- PG_MG_sum_16s[order(PG_MG_sum_16s$sum, decreasing = TRUE), ]
df_sorted_16s = df_sorted_16s[core_microbes_16s,]
rownames(df_sorted_16s)[which(rownames(df_sorted_16s) == "Burkholderia-Caballeronia-Paraburkholderia")] <- "Burkholderia"
# 绘制柱状图
library(ggplot2)
ggplot(df_sorted_16s, aes(x = reorder(rownames(df_sorted_16s), -sum), y = sum)) +
  geom_col(aes(fill = sum)) + 
  scale_fill_gradient(low = "#93CEDF", high = "#E4A644") + 
  geom_text(aes(label = sum), vjust = -0.5, size = 4) +  
  labs(x = "Core Bacterial Genus", y = "Merged Prevalence",fill = "Merged\nPrevalence") +
  theme_classic() + 
  coord_cartesian(ylim = c(20, NA))+
  theme(
    panel.background = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2), 
    plot.title = element_text(
      size = 20,
      face = "bold",
      color = "black",
      hjust = 0.5,  
      margin = margin(b = 10)  
    ),
    plot.subtitle = element_text(
      size = 12,
      color = "black",
      hjust = 0.5, 
      lineheight = 1.2,  
      margin = margin(b = 15)  
    ),
    axis.text.x = element_text(size = 14, color = "black", angle = -45, hjust = 0, vjust = 1),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y = element_text(size = 14, color = "black", angle = 0), 
    axis.title.y = element_text(size = 14, color = "black", face = "bold"),
    legend.title = element_text(size = 14, color = "black"), 
    legend.text = element_text(size = 12, color = "black"), 
    plot.margin = margin(t = 10,
                         r = 10,  
                         b = 10,  
                         l = 10,  
                         unit = "pt")
  )
ggsave("result/Merged_Prevalence_16s.pdf",width = 10,height = 7.5)


PG_MG_sum_its = merge(taxonomy_its_final[ ,c("taxid","Genus")],PG_taxid_its,by = "taxid")[ ,-1] %>%
  group_by(Genus) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  column_to_rownames(var = "Genus")
PG_MG_sum_its[PG_MG_sum_its != 0] <- 1
PG_MG_sum_its$sum = rowSums(PG_MG_sum_its)
write.table(rownames_to_column(PG_MG_sum_its,"PG"),file = "PG_MG_sum_its.txt",sep = "\t", 
            row.names = FALSE,col.names = TRUE,quote = FALSE,fileEncoding = "UTF-8")
PG_MG_sum_its <- PG_MG_sum_its[, c("sum", setdiff(names(PG_MG_sum_its), "sum"))]
df_sorted_its <- PG_MG_sum_its[order(PG_MG_sum_its$sum, decreasing = TRUE), ]
df_sorted_its = df_sorted_its[core_microbes_its,]

ggplot(df_sorted_its, aes(x = reorder(rownames(df_sorted_its), -sum), y = sum)) +
  geom_col(aes(fill = sum)) +  
  scale_fill_gradient(low = "#93CEDF", high = "#E4A644") +  
  geom_text(aes(label = sum), vjust = -0.5, size = 4) +  
  labs(x = "TOP 40 Bacterial Genus", y = "Merged Prevalence",fill = "Merged\nPrevalence") +
  theme_classic() + 
  coord_cartesian(ylim = c(20, NA))+
  theme(
    panel.background = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2), 
    plot.title = element_text(
      size = 20,
      face = "bold",
      color = "black",
      hjust = 0.5,  
      margin = margin(b = 10)  
    ),
    plot.subtitle = element_text(
      size = 12,
      color = "black",
      hjust = 0.5, 
      lineheight = 1.2,  
      margin = margin(b = 15)  
    ),
    axis.text.x = element_text(size = 14, color = "black", angle = -45, hjust = 0, vjust = 1),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y = element_text(size = 14, color = "black", angle = 0), 
    axis.title.y = element_text(size = 14, color = "black", face = "bold"),
    legend.title = element_text(size = 14, color = "black"), 
    legend.text = element_text(size = 12, color = "black"), 
    plot.margin = margin(t = 10,
                         r = 10,  
                         b = 10,  
                         l = 10,  
                         unit = "pt")
  )
ggsave("result/Merged_Prevalence_its.pdf",width = 8,height = 7.5)
```




## Draw the General Genus microbial evolutionary tree
### The detailed process of constructing the evolutionary tree and preparing the tree file has been omitted here.
### All the files for constructing the evolutionary tree are located in the "General_tree_16s" and "General_tree_its" folders.
### Use these files to draw the evolutionary tree on the itol website.



#Trait shifts with prevalence
#16S samples
```{r}
rm(list = ls())
source("Tools.R")
library(dplyr)
library(ggplot2)
library(forcats)
library(tidyr)
library(reshape2)
library(forcats)
taxonomy_16s_final = read.delim("taxonomy_16s_final.txt")
#The relationship between the taxonomic-level species composition and prevalence of the 16s sample
prevalence_Class_16s = taxonomy_16s_final[ ,c("prevalence","Class")]
prevalence_Class_16s[prevalence_Class_16s$prevalence >= 20,]$prevalence = 20

make_mutibar(datain = prevalence_Class_16s,
             out_dir = "result/Class_levels_change_16s.pdf",
             x_name = "prevalence",
             y_name = "Proportion of Taxid",
             legend_name = "Class",
             limit_num = 10,
             width = 15,
             height = 7,
             other_list = c("Unassigned","Incertae Sedis"),
             colorlist = c("#FCB366", "#FE816F", "#90D7C7","#93BCD0", "#AF3F4F", 
                           "#F3BB34", "#F8AD9A", "#5163B5", "#613767", "#56B394",
                           "#DC6D73", "#B2B896", "#E97DAE","#FB7A64","#EDCE35", 
                           "#E5C28C", "#CA8BB6", "#1A6C92", "#EAB46E", "#8CC3BD", 
                           "#4E56AB","#0FBFBC", "#E9A564", "#ADD565", "#ADADAD", 
                           "#8F9EC5", "#6C77B5", "#138A8E", "#4B9D77", "#8E60A8")
             )
Class_test = Calculate_correlation(prevalence_Class_16s,"prevalence","Class")
class_prop = as.data.frame(table(prevalence_Class_16s$Class))
main_class = class_prop$Var1[class_prop$Freq>200]
Class_test = Class_test[Class_test$Variable%in%main_class,]
print(Class_test)
#----------------------------------------------------------------------

#The relationship between the aerobic/anaerobic conditions of the 16s samples and the prevalence rate
prevalence_Aerobic_type_16s = taxonomy_16s_final[ ,c("prevalence","Aerobic_type")]
prevalence_Aerobic_type_16s <- prevalence_Aerobic_type_16s %>%
  replace(is.na(.), "Unknown")
prevalence_Aerobic_type_16s[prevalence_Aerobic_type_16s$prevalence >= 20,]$prevalence = 20
make_mutibar(datain = prevalence_Aerobic_type_16s,
             out_dir = "result/prevalence_Aerobic_type_change_16s.pdf",
             x_name = "prevalence",
             y_name = "Proportion of Taxid",
             legend_name = "Aerobic_type",
             limit_num = 10,
             width = 10,
             height = 4,
             other_list = c("Unassigned","Incertae Sedis","Unknown"),
             colorlist = c("#99C2C8","#D4693F","#E4A644","#eeeeee")
)
Aerobic_type_test = Calculate_correlation(prevalence_Aerobic_type_16s,"prevalence","Aerobic_type")
print(Aerobic_type_test)
#--------------------------------------------------------------------------


#The relationship between the biofilm-forming potential of 16s samples and their prevalence rate
prevalence_Forms_Biofilms_16s = taxonomy_16s_final[ ,c("prevalence","Forms_Biofilms")]
prevalence_Forms_Biofilms_16s <- prevalence_Forms_Biofilms_16s %>%
  replace(is.na(.), "Unknown")
prevalence_Forms_Biofilms_16s[prevalence_Forms_Biofilms_16s$prevalence >= 20,]$prevalence = 20
make_mutibar(datain = prevalence_Forms_Biofilms_16s,
             out_dir = "result/prevalence_Forms_Biofilms_change_16s.pdf",
             x_name = "prevalence",
             y_name = "Proportion of Taxid",
             legend_name = "Forms_Biofilms",
             limit_num = 10,
             width = 10,
             height = 4,
             other_list = c("Unassigned","Incertae Sedis","Unknown"),
             colorlist = c("#846992","#E8C2D9","#eeeeee")
)
Forms_Biofilms_test = Calculate_correlation(prevalence_Forms_Biofilms_16s,"prevalence","Forms_Biofilms")
print(Forms_Biofilms_test)
#--------------------------------------------------------------------------------



#The relationship between the Gram-positive and Gram-negative properties of the 16s samples and their prevalence rates
prevalence_Gram_type_16s = taxonomy_16s_final[ ,c("prevalence","Gram_type")]
prevalence_Gram_type_16s <- prevalence_Gram_type_16s %>%
  replace(is.na(.), "Unknown")
prevalence_Gram_type_16s[prevalence_Gram_type_16s$prevalence >= 20,]$prevalence = 20
make_mutibar(datain = prevalence_Gram_type_16s,
             out_dir = "result/prevalence_Gram_type_change_16s.pdf",
             x_name = "prevalence",
             y_name = "Proportion of Taxid",
             legend_name = "Gram_type",
             limit_num = 10,
             width = 10,
             height = 4,
             other_list = c("Unassigned","Incertae Sedis","Unknown"),
             colorlist = c("#4B5B74","#A1B1CB","#eeeeee")
)
Gram_type_test = Calculate_correlation(prevalence_Gram_type_16s,"prevalence","Gram_type")
print(Gram_type_test)
```

## As popularity increases, the changes in genetic diversity occur.
```{r,fig.show='animate'}
rm(list = ls())

library(ggplot2)
library(ggpubr)
library(dplyr)

taxonomy_16s_final <- read.delim("taxonomy_16s_final.txt")

df_ko_prevalance <- taxonomy_16s_final[, c("prevalence", "KO_richness")]
df_ko_prevalance <- df_ko_prevalance[df_ko_prevalance$KO_richness != "Unknown", ]
df_ko_prevalance$KO_richness <- as.numeric(df_ko_prevalance$KO_richness)

df_summary <- df_ko_prevalance %>%
  group_by(prevalence) %>%
  summarise(
    KO_richness_max = max(KO_richness, na.rm = TRUE),
    KO_richness_min = min(KO_richness, na.rm = TRUE)
  )

lm_model <- lm(KO_richness_min ~ prevalence, data = df_summary)

# Extract equation and statistics for annotation
lm_eq     <- paste0("y = ", round(coef(lm_model)[1], 2), " + ", round(coef(lm_model)[2], 2), "x")
r_squared <- paste0("R² = ", round(summary(lm_model)$r.squared, 4))
p_value   <- summary(lm_model)$coefficients["prevalence", "Pr(>|t|)"]
p_value_text <- paste0("P = ", signif(p_value, 3))

final_plot <- ggplot() +
  # Range ribbon (min–max per prevalence bin)
  geom_ribbon(data = df_summary, 
              aes(x = prevalence, ymin = KO_richness_min, ymax = KO_richness_max),
              fill = "grey", alpha = 0.15) +
  
  # Individual data points with color gradient
  geom_point(data = df_ko_prevalance, 
             aes(x = prevalence, y = KO_richness, color = KO_richness),
             alpha = 0.6, size = 1.5) +
  
  scale_color_gradient(low = "#93CEDF", high = "#D4693F", name = "KO Richness") +
  
  geom_line(data = df_summary, aes(x = prevalence, y = KO_richness_max), 
            color = "#D4693F", size = 1) +
  geom_line(data = df_summary, aes(x = prevalence, y = KO_richness_min), 
            color = "#93CEDF", size = 1) +
  
  geom_smooth(data = df_summary, aes(x = prevalence, y = KO_richness_min),
              method = "lm", formula = y ~ x, se = FALSE, 
              color = "black", linetype = "dashed") +
  
  annotate("text", x = Inf, y = Inf,
           label = paste(lm_eq, r_squared, sep = "\n"),
           hjust = 1.1, vjust = 1.1, size = 5) +
  
  labs(x = "Prevalence", y = "KO_Richness") +
  theme_bw() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    axis.text.x  = element_text(size = 14, color = "black"),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y  = element_text(size = 14, color = "black"),
    axis.title.y = element_text(size = 14, color = "black", face = "bold"),
    legend.title = element_text(size = 14),
    legend.text  = element_text(size = 12),
    plot.margin  = margin(10, 10, 10, 10, "pt")
  )

print(final_plot)

ggsave("result/prevalence_KO_richness_change_16s.pdf",
       plot = final_plot, width = 10, height = 6, units = "in")
```

#ITS Samples
```{r}
rm(list = ls())
source("Tools.R")
library(dplyr)
library(ggplot2)
library(forcats)
library(tidyr)
library(reshape2)
library(forcats)
taxonomy_its_final = read.delim("taxonomy_its_final.txt")
#The relationship between the taxonomic species composition of ITS samples and their prevalence rates
prevalence_Class_its = taxonomy_its_final[ ,c("prevalence","Class")]
prevalence_Class_its[prevalence_Class_its$prevalence >= 15,]$prevalence = 15
make_mutibar(datain = prevalence_Class_its,
             out_dir = "result/prevalence_Class_change_its.pdf",
             x_name = "prevalence",
             y_name = "Proportion of Taxid",
             legend_name = "Class",
             limit_num = 10,
             width = 15,
             height = 7,
             other_list = c("Unassigned","Incertae Sedis","Unknown"),
             colorlist =  c("#FCB366", "#FE816F", "#90D7C7","#93BCD0", "#AF3F4F", 
                            "#F3BB34", "#F8AD9A", "#5163B5", "#613767", "#56B394",
                            "#DC6D73", "#B2B896", "#E97DAE","#FB7A64","#EDCE35", 
                            "#E5C28C", "#CA8BB6", "#1A6C92", "#EAB46E", "#8CC3BD", 
                            "#4E56AB","#0FBFBC", "#E9A564", "#ADD565", "#ADADAD", 
                            "#8F9EC5", "#6C77B5", "#138A8E", "#4B9D77", "#8E60A8")
)
Class_test = Calculate_correlation(prevalence_Class_its,"prevalence","Class")
class_prop = as.data.frame(table(prevalence_Class_its$Class))
main_class = class_prop$Var1[class_prop$Freq>50]
Class_test = Class_test[Class_test$Variable%in%main_class,]
print(Class_test)
#The relationship between the proportion of mycorrhizal fungi in ITS samples and their prevalence rate
prevalence_Mycorrhizal_its = taxonomy_its_final[ ,c("prevalence","Mycorrhizal")]
prevalence_Mycorrhizal_its <- prevalence_Mycorrhizal_its %>%
  replace(is.na(.), "Unknown")
prevalence_Mycorrhizal_its[prevalence_Mycorrhizal_its$prevalence >= 15,]$prevalence = 15
prevalence_Mycorrhizal_its[prevalence_Mycorrhizal_its$prevalence >= 15,]$prevalence = 15
make_mutibar(datain = prevalence_Mycorrhizal_its,
             out_dir = "result/prevalence_Mycorrhizal_change_its.pdf",
             x_name = "prevalence",
             y_name = "Proportion of Taxid",
             legend_name = "Mycorrhizal",
             limit_num = 10,
             width = 10,
             height = 4,
             other_list = c("Unassigned","Incertae Sedis","Unknown"),
             colorlist = c("#C1D2E2","#4B5B74","#eeeeee")
)
Mycorrhizal_test = Calculate_correlation(prevalence_Mycorrhizal_its,"prevalence","Mycorrhizal")
print(Mycorrhizal_test)
```


#Identify which genes have an increased or decreased occurrence rate as the popularity increases or decreases.
```{r,warning=FALSE}
rm(list = ls())
source("Tools.R")

library(dplyr)
library(tidyr)

taxonomy_16s_final <- read.delim("taxonomy_16s_final.txt")
ko_data <- read.table("raw/ko_data.tsv", header = TRUE, row.names = 1, sep = "\t")

colnames(ko_data) <- gsub("ko\\.", "", colnames(ko_data))
ko_data[ko_data != 0] <- 1

type_data <- taxonomy_16s_final[, c("taxid", "prevalence")]
type_data$prevalence[type_data$prevalence >= 20] <- 20

indata <- merge(type_data, ko_data, by.x = "taxid", by.y = "row.names")

# Convert to long format and calculate proportion of presence per prevalence group
result_df <- indata %>%
  pivot_longer(
    cols = -c(taxid, prevalence),
    names_to = "gene",
    values_to = "presence"
  ) %>%
  group_by(prevalence, gene) %>%
  summarise(
    proportion = sum(presence == 1, na.rm = TRUE) / n(),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = gene,
    values_from = proportion
  )

alpha <- 0.001

positive_vars <- character(0)
negative_vars <- character(0)

result_df$prevalence <- as.numeric(result_df$prevalence)
independent_var <- result_df$prevalence
dependent_df <- result_df[, -1, drop = FALSE]

for (col_name in names(dependent_df)) {
  dependent_var <- dependent_df[[col_name]]
  
  test_result <- tryCatch({
    cor.test(independent_var, dependent_var, method = "pearson")
  }, error = function(e) NULL)
  
  if (is.null(test_result)) next
  
  p_value <- test_result$p.value
  cor_coef <- test_result$estimate
  
  if (!is.na(p_value) && !is.na(cor_coef)) {
    if (p_value < alpha) {
      if (cor_coef > 0) {
        positive_vars <- c(positive_vars, col_name)
      } else if (cor_coef < 0) {
        negative_vars <- c(negative_vars, col_name)
      }
    }
  }
}

positive_vars <- data.frame(p_ko = positive_vars)
negative_vars <- data.frame(n_ko = negative_vars)

# Load KEGG classification
ko_level_table <- read.delim("raw/ko_level_table.txt")

p_ko_level <- merge(positive_vars, ko_level_table, by.x = "p_ko", by.y = "KO")
n_ko_level <- merge(negative_vars, ko_level_table, by.x = "n_ko", by.y = "KO")

# Top 10 level3 categories
p_level3_10 <- names(sort(table(p_ko_level$level3), decreasing = TRUE)[1:10])
n_level3_10 <- names(sort(table(n_ko_level$level3), decreasing = TRUE)[1:10])

# Unique and shared functions
p_level3_10_unique <- setdiff(p_level3_10, n_level3_10)
n_level3_10_unique <- setdiff(n_level3_10, p_level3_10)
common_func <- intersect(p_level3_10, n_level3_10)

print(p_level3_10_unique)
print(n_level3_10_unique)
print(common_func)
```

#Use a pie chart to observe the species composition and characteristics in the groups with low popularity and low genetic diversity.
```{r,fig.show='animate'}
rm(list = ls())
source("Tools.R")
taxonomy_16s_final = read.delim("taxonomy_16s_final.txt")
taxonomy_16s_final = taxonomy_16s_final[taxonomy_16s_final$KO_richness != "Unknown",]
taxonomy_16s_final$KO_richness = as.numeric(taxonomy_16s_final$KO_richness)
taxonomy_16s_final_ko800 = taxonomy_16s_final[taxonomy_16s_final$KO_richness <= 800 & taxonomy_16s_final$prevalence <=5,]
p1 = draw_pie_chart(taxonomy_16s_final_ko800,"Class", label = F, pre = TRUE,"Class",limit_n = 10)
p1
ggsave("result/ko800_Class.pdf",p1,width = 7,height = 6)

p2 = draw_pie_chart(taxonomy_16s_final_ko800,"Aerobic_type", label = F, pre = TRUE,"Aerobic_type",color_list = c("#D4693F","#eeeeee","#E4A644","#99C2C8"))
p2
ggsave("result/ko800_Aerobic_type.pdf",p2,width = 7,height = 6)

p3 = draw_pie_chart(taxonomy_16s_final_ko800,"Forms_Biofilms", label = F, pre = TRUE,"Forms_Biofilms",color_list = c("#E8C2D9","#846992","#eeeeee"))
p3
ggsave("result/ko800_Forms_Biofilms.pdf",p3,width = 7,height = 6)



#Observe the composition of mycorrhizal fungi and the main plants they are distributed among.
taxonomy_its_final = read.delim("taxonomy_its_final.txt")
Mycorrhizal_TAX = taxonomy_its_final[taxonomy_its_final$Mycorrhizal=="TRUE", ]

p4 = draw_pie_chart(Mycorrhizal_TAX,"Order", label = F, pre = TRUE,"Mycorrhizal_Order",limit_n = 15)
p4
ggsave("result/Mycorrhizal_Order_pie.pdf", plot = p4,width = 6,height = 5,units = "in")
Mycorrhizal_TAX[is.na(Mycorrhizal_TAX)]="Unknown"
p5 = draw_pie_chart(Mycorrhizal_TAX[Mycorrhizal_TAX$specialist_plant != "Unknown", ],"specialist_plant", label = F, pre = TRUE,"Mycorrhizal_specialist_plant",limit_n = 15)
p5
ggsave("result/Mycorrhizal_specialist_plant_pie.pdf", plot = p5,width = 6,height = 5,units = "in")
```

# Search for microorganisms that occur at a frequency difference of more than 30% in C3 and C4 plants.
```{r,fig.show='animate'}
rm(list = ls())

library(tidyr)
library(tibble)
library(tidyverse)

# Load data
metadata_16s       <- read.csv("raw/metadata_all16s.csv")
taxonomy_16s_final <- read.delim("taxonomy_16s_final.txt")
PG_taxid_16s       <- read.delim("PG_taxid_16s.txt", row.names = 1)
sample_taxid_01_16s <- read.delim("sample_taxid_01_16s.txt", row.names = 1)

# Split metadata by photosynthetic pathway
c4_metadata_16s <- metadata_16s[metadata_16s$pathway == "C4", ]
C3_metadata_16s <- metadata_16s[metadata_16s$pathway == "C3", ]

# Genus-level presence/absence table
Sid_BG_01_16s <- merge(taxonomy_16s_final[, c("taxid", "Genus")], 
                       sample_taxid_01_16s, 
                       by.x = "taxid", by.y = "row.names")[ ,-1] %>%
  group_by(Genus) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  column_to_rownames(var = "Genus")

Sid_BG_01_16s[Sid_BG_01_16s != 0] <- 1

# Subset by group
c4_01_16s_BG <- Sid_BG_01_16s[, c4_metadata_16s$SampleID]
c4_01_16s_BG$rate <- rowSums(c4_01_16s_BG) / ncol(c4_01_16s_BG)

C3_01_16s_BG <- Sid_BG_01_16s[, C3_metadata_16s$SampleID]
C3_01_16s_BG$rate <- rowSums(C3_01_16s_BG) / ncol(C3_01_16s_BG)

# Prepare comparison (difference ≥ 30%)
all_genus <- rownames(c4_01_16s_BG)
all_genus <- all_genus[!all_genus %in% c("Unassigned", "Incertae Sedis")]

C4_VS_C3_allrate <- data.frame(
  Genus   = all_genus,
  C4_rate = c4_01_16s_BG[all_genus, ]$rate,
  C3_rate = C3_01_16s_BG[all_genus, ]$rate
)

C4_VS_C3_allrate$differ <- C4_VS_C3_allrate$C4_rate - C4_VS_C3_allrate$C3_rate

# Select genera with strong difference
C4_VS_C3_differ_p <- C4_VS_C3_allrate[C4_VS_C3_allrate$differ >= 0.3, ]
C4_VS_C3_differ_p <- C4_VS_C3_differ_p[order(C4_VS_C3_differ_p$C4_rate, decreasing = TRUE), -4]

C4_VS_C3_differ_n <- C4_VS_C3_allrate[C4_VS_C3_allrate$differ <= -0.3, ]
C4_VS_C3_differ_n <- C4_VS_C3_differ_n[order(C4_VS_C3_differ_n$C3_rate, decreasing = FALSE), -4]

C4_VS_C3_differ <- rbind(C4_VS_C3_differ_p, C4_VS_C3_differ_n)

# Prepare data for plotting
df_long <- pivot_longer(C4_VS_C3_differ, 
                        cols = c(C4_rate, C3_rate), 
                        names_to = "variable", 
                        values_to = "value")

df_long$Genus <- factor(df_long$Genus, levels = C4_VS_C3_differ$Genus)

# Plot
ggplot(df_long, aes(x = Genus, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(x = "Genus", y = "Occurrence Rate", title = "", fill = "") +
  scale_fill_manual(values = c("C4_rate" = "#f7e09b", "C3_rate" = "#517d7e")) +
  theme_classic() +
  theme(
    legend.position = "top",
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    axis.text.x  = element_text(size = 14, color = "black", angle = -45, hjust = 0, vjust = 1),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y  = element_text(size = 14, color = "black"),
    axis.title.y = element_text(size = 14, color = "black", face = "bold"),
    legend.title = element_text(size = 14, color = "black"),
    legend.text  = element_text(size = 14, color = "black"),
    plot.margin  = margin(t = 10, r = 30, b = 10, l = 10, unit = "pt")
  )

ggsave("result/C4_VS_C3_differ.pdf", width = 12, height = 5)
```

# Observe the storage situation of nitrogen-fixing bacteria in different plants
```{r,fig.show='animate'}
rm(list = ls())
source("Tools.R")
library(tidyr)
library(tibble)
library(tidyverse)
library(pheatmap)
library(ape)
nf = read.delim("raw/nitrogen_fixation.txt")
metadata_16s = read.csv("raw/metadata_all16s.csv")
taxonomy_16s_final = read.delim("taxonomy_16s_final.txt")
PG_taxid_16s = read.delim("PG_taxid_16s.txt")

PG_MG_sum_16s = merge(taxonomy_16s_final[ ,c("taxid","Genus")],PG_taxid_16s,by = "taxid")[ ,-1] %>%
  group_by(Genus) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  column_to_rownames(var = "Genus")
rownames(PG_MG_sum_16s)[which(rownames(PG_MG_sum_16s) == "Burkholderia-Caballeronia-Paraburkholderia")] <- "Burkholderia"


#Create the plant classification tree corresponding to the 16s sample (please visualize it using itol)
plant_taxonomy = unique(metadata_16s[ ,c(5:9)])
plant_taxonomy$Kingdom = "Plantae"
write.table(plant_taxonomy,file = "result/plant_taxonomy.txt",sep = "\t", 
            row.names = FALSE,col.names = TRUE,quote = FALSE,fileEncoding = "UTF-8")
class_tree("result/plant_taxonomy.txt","result/plant_tree_16s.newick","Genus")
#Arrange in the order specified in ITOL
plant_order = c("Azolla","Achnatherum","Zea","Urochloa","Triticum","Thinopyrum","Sorghum",
                "Setaria","Saccharum","Panicum","Oryza","Lolium","Hordeum","Brachypodium",
                "Aegilops","Cannabis","Anemone","Solanum","Scabiosa","Parnassia","Manihot",
                "Citrus","Citrullus","Nothofagus","Carya","Panax","Astrantia","Theobroma",
                "Gossypium","Cistanche","Phelipanche","Euphrasia","Saussurea","Ligularia",
                "Helianthus","Coffea","Halenia","Gentianella","Gentiana","Heliosperma",
                "Chenopodium","Stenocereus","Ferocactus","Echinocactus","Thlaspi","Raphanus",
                "Descurainia","Brassica","Arabidopsis","Sophora","Pisum","Phaseolus","Lens","Glycine")

# Convert it into a matrix, which is the required data format for the pheatmap function.


PG_NF_sum_16s = PG_MG_sum_16s[nf$Genus, ]
PG_NF_sum_16s[PG_NF_sum_16s != 0] <- 1
PG_NF_sum_16s = PG_NF_sum_16s[ ,plant_order]
data_matrix <- as.matrix(t(PG_NF_sum_16s))
my_colors <- c("white", "#A23F25")

pheatmap(data_matrix,
               color = my_colors,        
               cluster_rows = F,      
               cluster_cols = F,     
               clustering_distance_cols = "euclidean", 
               clustering_method = "complete", 
               show_rownames = TRUE,     
               show_colnames = TRUE,     
               main = "The Distribution of Symbiotic Nitrogen-fixing Bacteria" 
)


PG_MG_sum_16s[PG_MG_sum_16s != 0] <- 1
PG_MG_sum_16s_t = t(PG_MG_sum_16s)
PG_MG_sum_16s$sum = rowSums(PG_MG_sum_16s)

PG_NF_sum_16s = merge(nf,PG_MG_sum_16s,by.x ="Genus",by.y = "row.names")
df_sorted <- PG_NF_sum_16s[order(PG_NF_sum_16s$sum, decreasing = TRUE), ]
df_sorted <- na.omit(df_sorted)

df_sorted$color <- ifelse(df_sorted$classify == "Endophytic", "#FCB366",
                          ifelse(df_sorted$classify == "Rhizobial", "#FE816F",
                                 ifelse(df_sorted$classify == "Free_living", "#90D7C7", NA)))
ggplot(df_sorted, aes(x = reorder(df_sorted$Genus, -sum), y = sum, fill = Genus)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = setNames(df_sorted$color, df_sorted$Genus)) +
  geom_text(aes(label = sum), vjust = -0.5, size = 4)+
  theme_minimal() +
  labs(x = "Nitrogen-fixing Bacteria Genus", y = "Merged Prevalence")+
  theme_bw() + 
  theme(
    panel.background = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2), 
    plot.title = element_text(
      size = 20,
      face = "bold",
      color = "black",
      hjust = 0.5, 
      margin = margin(b = 10)  
    ),
    plot.subtitle = element_text(
      size = 12,
      color = "black",
      hjust = 0.5,  
      lineheight = 1.2, 
      margin = margin(b = 15)
    ),
    axis.text.x = element_text(size = 14, color = "black", angle = -45, hjust = 0, vjust = 1),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y = element_text(size = 14, color = "black", angle = 0), 
    axis.title.y = element_text(size = 14, color = "black", face = "bold"), 
    legend.title = element_text(size = 14, color = "black"), 
    legend.text = element_text(size = 12, color = "black"), 
    plot.margin = margin(t = 10,  
                         r = 70,  
                         b = 10,  
                         l = 10,  
                         unit = "pt")  
  )
ggsave("result/NF_BAR_SUM.pdf",width = 15,height = 9)



PG_MG_sum_16s_t_nf =as.data.frame(PG_MG_sum_16s_t[ ,nf$Genus]) 
PG_MG_sum_16s_t_nf$sum = rowSums(PG_MG_sum_16s_t_nf)
df_sorted <- PG_MG_sum_16s_t_nf[order(PG_MG_sum_16s_t_nf$sum, decreasing = TRUE), ]
df_sorted <- na.omit(df_sorted)
ggplot(df_sorted, aes(x = reorder(rownames(df_sorted), -sum), y = sum)) +
  geom_col(aes(fill = sum)) + 
  scale_fill_gradient(low = "#93CEDF", high = "#E4A644") + 
  geom_text(aes(label = sum), vjust = -0.5, size = 4) + 
  labs(x = "Plant Genus", y = "Nitrogen-fixing Bacteria Richness",fill = "Nitrogen-fixing\nBacteria Richness\n(Max= 41)") +
  theme_bw() + 
  theme(
    panel.background = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, size = 1.2), 
    plot.title = element_text(
      size = 20,
      face = "bold",
      color = "black",
      hjust = 0.5, 
      margin = margin(b = 10)  
    ),
    plot.subtitle = element_text(
      size = 12,
      color = "black",
      hjust = 0.5,  
      lineheight = 1.2, 
      margin = margin(b = 15)
    ),
    axis.text.x = element_text(size = 14, color = "black", angle = -45, hjust = 0, vjust = 1),
    axis.title.x = element_text(size = 14, color = "black", face = "bold"),
    axis.text.y = element_text(size = 14, color = "black", angle = 0), 
    axis.title.y = element_text(size = 14, color = "black", face = "bold"), 
    legend.title = element_text(size = 14, color = "black"), 
    legend.text = element_text(size = 12, color = "black"), 
    plot.margin = margin(t = 10,  
                         r = 70,  
                         b = 10,  
                         l = 10,  
                         unit = "pt")  
  )
ggsave("result/Plant_NF_Richness.pdf",width = 15,height = 9)


General16s_genus = read.delim("General_tree_16s/annotation_16s.txt")$Genus
General_nf_genus = intersect(General16s_genus, nf$Genus)
print(General_nf_genus)
```

# Draw a heatmap of microbial functions
```{r,fig.show='animate'}
rm(list = ls())
library(pheatmap)

micro_function    <- read.delim("raw/micro_function.txt")
taxonomy_16s_final <- read.delim("taxonomy_16s_final.txt")
PG_MG_sum_16s     <- read.delim("PG_MG_sum_16s.txt", row.names = 1)
PG_MG_sum_its     <- read.delim("PG_MG_sum_its.txt", row.names = 1)

micro_function_merge <- micro_function %>%
  group_by(Genus) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  column_to_rownames(var = "Genus")

high_prevalence_16s <- rownames(PG_MG_sum_16s[PG_MG_sum_16s$sum >= 28, ] %>% arrange(desc(sum)))
high_prevalence_its <- rownames(PG_MG_sum_its[PG_MG_sum_its$sum >= 15, ] %>% arrange(desc(sum)))

micro_function_merge_16s <- micro_function_merge[high_prevalence_16s, ]
micro_function_merge_16s <- na.omit(micro_function_merge_16s)
micro_function_merge_16s <- micro_function_merge_16s[rowSums(micro_function_merge_16s != 0) > 0, ]

micro_function_merge_its <- micro_function_merge[high_prevalence_its, ]
micro_function_merge_its <- na.omit(micro_function_merge_its)
micro_function_merge_its <- micro_function_merge_its[rowSums(micro_function_merge_its != 0) > 0, ]

# Custom green gradient: 0–5 light-to-dark green, >5 stays darkest green
create_green_gradient <- function(n_colors = 32) {
  greens <- colorRampPalette(c("#ffffff", "#E5F5E0", "#C7E9C0",
                               "#A1D99B", "#74C476", "#41AB5D"))(6)
  
  colors <- character(n_colors)
  for (i in 0:(n_colors-1)) {
    if (i <= 5) {
      colors[i+1] <- greens[i+1]
    } else {
      colors[i+1] <- "#238B45"
    }
  }
  return(colors)
}

# Method to show numbers only when value > 0
create_number_matrix <- function(data_matrix) {
  num_matrix <- matrix(as.character(data_matrix),
                       nrow = nrow(data_matrix),
                       ncol = ncol(data_matrix))
  num_matrix[data_matrix == 0] <- ""
  return(num_matrix)
}

# ─── 16S version ───────────────────────────────────────
inputmatrix_16s <- as.matrix(micro_function_merge_16s) + 1
number_matrix_16s <- create_number_matrix(as.matrix(micro_function_merge_16s))

pheatmap(inputmatrix_16s,
         color = create_green_gradient(33),
         breaks = seq(0, 33, length.out = 33),
         display_numbers = number_matrix_16s,
         number_color = "black",
         fontsize_number = 10,
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         show_rownames = TRUE,
         show_colnames = TRUE,
         main = "",
         border_color = "gray80",
         cellwidth = 40,
         cellheight = 15)

# ─── ITS version ───────────────────────────────────────
inputmatrix_its <- as.matrix(micro_function_merge_its) + 1
number_matrix_its <- create_number_matrix(as.matrix(micro_function_merge_its))

pheatmap(inputmatrix_its,
         color = create_green_gradient(22),
         breaks = seq(0, 22, length.out = 22),
         display_numbers = number_matrix_its,
         number_color = "black",
         fontsize_number = 10,
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         show_rownames = TRUE,
         show_colnames = TRUE,
         main = "",
         border_color = "gray80",
         cellwidth = 40,
         cellheight = 15)
```




